### 2022-03-08 01-processing-raw-data_reviews.Rmd
first commit

---
title: "Processing of raw data"
author: "Ulrik Lyngs"
date: '2022-03-01'
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(rvest)
library(janitor)
library(jsonlite)
```

# Chrome Web store
## Extract reviews from page sources
In this case, we've store .RDS files that contain the page sources.

We'll first create the helper functions we need.

```{r}
get_reviews_from_full_source <- function(source) {
  source %>% 
    read_html() %>% 
    html_nodes(".ba-fb > div > .ba-bc-Xb")
}

extract_info_from_review_div <- function(review_div) {
  user_name <- review_div %>% 
    html_node(".comment-thread-displayname") %>% 
    html_text()

  date_modified <- review_div %>% 
    html_node(".ba-Eb-Nf") %>% 
    html_text()

  numstars <- review_div %>% 
    html_nodes(".rsw-starred") %>% 
    length()

  text <- review_div %>% 
    html_node(".ba-Eb-ba") %>% 
    html_text()

  result_tibble <- tibble(type = "review",
                          user_name,
                          reply_to = NA,
                          date_modified,
                          numstars,
                          text)

  replies <- review_div %>% 
    html_nodes(".Fg-b-ob-fb .ba-bc-Xb-K")

  if (length(replies) > 0) {
    for (i in seq_along(replies)) {
      name_replier <- replies[[i]] %>% 
        html_node(".comment-thread-displayname") %>% 
        html_text()

      date_modified <- replies[[i]] %>% 
        html_node(".ba-Eb-Nf") %>% 
        html_text()

      text <- replies[[i]] %>% 
        html_node(".ba-Eb-ba") %>% 
        html_text()

      result_tibble <- bind_rows(result_tibble,
                                 tibble(type = "reply",
                                        reply_to = user_name,
                                        user_name = name_replier,
                                        date_modified,
                                        numstars = NA,
                                        text))
    }
  }

  return(result_tibble)

}

get_review_content_from_list <- function(list_with_review_divs) {
  map_dfr(list_with_review_divs, extract_info_from_review_div)
}

return_reviews_from_source <- function(tibble_with_page_sources) {
  tibble_with_page_sources %>% 
    mutate(review_divs = map(unlist(value), get_reviews_from_full_source),
           review_content = map(review_divs, get_review_content_from_list)) %>% 
    select(-value, -review_divs) %>% 
    rename(review_page = name) %>% 
    unnest(cols = c(review_content)) %>% 
    filter(!is.na(text))
}
```

We'll do things in two steps, to avoid running into memory issues.

We first just read in the RDS files one at a time, extract the reviews, and store them in a CSV file (note: we're not doing this in a very efficient way, so it might take dozens of seconds for extensions with many reviews, e.g. 'Block Site'):

It takes a while, so we'll set it to not evaluate by default.

```{r eval=FALSE}
#dir.create(here::here("data", "processed", "chrome", "2019-02-08-reviews_extracted"))

read_process_and_store <- function(filepath, filename) {
  savename <- filename %>% str_remove("\\.rds") %>% make_clean_names()

  print(str_c("doing ", filename))

  read_rds(filepath) %>% 
    return_reviews_from_source() %>% 
    write_csv(str_c(here::here("data", "processed", "chrome", "2019-02-08-reviews_extracted/"), savename, ".csv"))
}

rds_files_w_reviews <- tibble(
  filepath = list.files(path = here::here("data", "raw", "chrome", "2019-02-08-page_sources"), pattern = "rds$", full.names = TRUE),
  filename = basename(filepath)
)

rds_files_w_reviews %>% 
  pwalk(read_process_and_store)
```

## Clean up and save out

```{r}
review_CSVs <- list.files(path = here::here("data", "processed", "chrome", "2019-02-08-reviews_extracted"), pattern = "csv$", full.names = TRUE)

chrome_reviews <- map_dfr(review_CSVs, read_csv, 
                          col_types = cols(
                            numstars = col_integer(),
                            reply_to = col_character())
                          )

chrome_reviews %>% 
  # add date
  mutate(date_modified = str_remove(date_modified, "Modified "),
         date_modified = case_when(
          str_detect(date_modified, "hours ago") ~ "Feb 8, 2019",
          str_detect(date_modified, "1 day ago") ~ "Feb 7, 2019",
          str_detect(date_modified, "2 days ago") ~ "Feb 6, 2019",
          str_detect(date_modified, "3 days ago") ~ "Feb 5, 2019",
          str_detect(date_modified, "4 days ago") ~ "Feb 4, 2019",
          str_detect(date_modified, "5 days ago") ~ "Feb 3, 2019",
          str_detect(date_modified, "6 days ago") ~ "Feb 2, 2019",
          TRUE ~ date_modified
        ),
    date_modified = lubridate::mdy(date_modified)) %>% 
  rename(rating = numstars) %>% 
  distinct(across(-review_page)) %>% 
  write_csv(here::here("data", "processed", "chrome", "2019-02-08-chrome_reviews_unique.csv"))

```

# Google Play
Here we have all the reviews in one big json file

```{r}
play_reviews_json <- read_json(here::here("data", "raw", "play", "2019_03_19_play_all_reviews.json"))

# have a look at the first app
# play_reviews_json[[1]] %>% 
#   View()

# put the data in a tibble
play_reviews <- tibble(
  app_id = map_chr(play_reviews_json, "packageName"),
  num_reviews = map_int(play_reviews_json, "numberOfReviews"),
  reviews = map(play_reviews_json, "reviews")
) %>% 
  arrange(desc(num_reviews))

# let's just see how many we have
play_reviews %>% 
  summarise(num_apps = n(),
    num_reviews = sum(num_reviews))

```

Now we grab the information we want from the reviews and put it in a tibble

```{r}
# let's make a function to extract the information we want from the reviews
extract_play_review_info <- function(play_store_review) {
  tibble(
    user_name = map_chr(play_store_review, "userName"),
    date = map_chr(play_store_review, "date"),
    rating = map_int(play_store_review, "score"),
    text = map_chr(play_store_review, "text")
  )
}

# put 'em all in one big happy unnested tibble
play_reviews_tibble <- play_reviews %>% 
  mutate(review_info = map(reviews, extract_play_review_info)) %>% 
  select(-reviews, -num_reviews) %>% 
  unnest(cols = c(review_info))

```

## Save out unique reviews
Then we remove duplicates and save them out

```{r}
play_reviews_tibble %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  distinct() %>% 
  write_csv(here::here("data", "processed", "play", "2019_03_19-play_reviews_unique.csv"))

```

# Apple App Store
Here we have things in separate files.

## JSON files from the GitHub scraper
The scraper from https://github.com/facundoolano/app-store-scraper has given us JSON files

```{r}
# have a look at one
(example_apple_json <- read_json(here::here("data", "raw", "apple", "gh_scraper_2019-03-19_individual_reviews", "club.donutdog.ios.json")))

# grab filepaths
json_files <- list.files(path  = here::here("data", "raw", "apple", "gh_scraper_2019-03-19_individual_reviews"), pattern = "json$", full.names = TRUE)

apple_reviews_json <- json_files %>% 
  map(read_json)

apple_reviews <- tibble(
  app_id = map_chr(apple_reviews_json, "app"),
  reviews = map(apple_reviews_json, "reviews"),
  num_reviews = map_int(reviews, length)
)

# let's just see how many we have
apple_reviews %>% 
  summarise(num_apps = n(),
            num_reviews = sum(num_reviews))

```

Let's actually grab the information we want and put in a useful format.

NOTE: Apple reviews have a heading for the reviews - we'll merge this into the text.

```{r}
extract_apple_review_info <- function(apple_review) {
  tibble(
    user_name = map_chr(apple_review, "userName"),
    rating = map_int(apple_review, "score"),
    review_title = map_chr(apple_review, "title"),
    review_text = map_chr(apple_review, "text")
  )
}

# put 'em all in one big happy unnested tibble
apple_review_tibble <- apple_reviews %>% 
  mutate(review_info = map(reviews, extract_apple_review_info)) %>% 
  select(-reviews) %>% 
  unnest()

```

We noticed on manual inspection that these apps had reviews on their store pages that the github scraper missed:
- pabloweb.net.SelfControl
- com.imobiapp.screentimer
- com.getcluster.Compose
- club.donutdog.ios
- com.getfeedless.feedless
- com.riko.suyasaso
- com.bellostudios.hooked
- com.jordan-carney.Liberate
- org.aliyilmaz.notelr
- ca.genoe.Refrain
- vn.wehelp.SelfControlLite
- com.yeowjin.simpletext
- com.usedopamine.app.space
- com.erichuju.pomodoro
- com.bluecocoa.time2

We scraped the page sources for these apps ourselves. Let's add them in.

## Extract reviews from page sources
```{r}
apps_w_missing_reviews <- c('pabloweb.net.SelfControl','com.imobiapp.screentimer','com.getcluster.Compose','club.donutdog.ios','com.getfeedless.feedless','com.riko.suyasaso','com.bellostudios.hooked','com.jordan-carney.Liberate','org.aliyilmaz.notelr','ca.genoe.Refrain','vn.wehelp.SelfControlLite','com.yeowjin.simpletext','com.usedopamine.app.space','com.erichuju.pomodoro','com.bluecocoa.time2')

length(apps_w_missing_reviews)
```

Let's make the helper functions:

```{r}
# for extracting developer responses
extract_dev_response_text <- function(review_div){
  html_nodes(review_div, '.we-customer-review__body') %>% .[[2]] %>%  html_text()
}
safe_extract_dev_response_text <- possibly(extract_dev_response_text, NA_character_)

grab_reviews_from_single_div <- function(single_review_div){
  tibble(
    user_name = html_node(single_review_div, '.we-customer-review__user') %>% html_text(),
    date = html_node(single_review_div, 'time') %>% html_text(),
    rating = html_node(single_review_div, '.we-customer-review__rating') %>% html_attrs() %>% enframe() %>% filter(name == 'aria-label') %>% pull(value),
    review_title = html_node(single_review_div, '.we-customer-review__title') %>% html_text(),
    review_text = html_node(single_review_div, '.we-customer-review__body') %>% html_text(),
    dev_response_title = html_node(single_review_div, '.we-customer-review__header--response .we-customer-review__title') %>% html_text(),
    dev_response_date = html_node(single_review_div, '.we-customer-review__header--response time') %>% html_text(),
    dev_response_text = safe_extract_dev_response_text(single_review_div)
  )

}

iterate_over_review_divs <- function(div_with_reviews){
  map_dfr(div_with_reviews, grab_reviews_from_single_div)
}

```

Then let's do it

```{r}
reviews_visible_on_store_pages <- tibble(
  app_id = apps_w_missing_reviews,
  html = map(apps_w_missing_reviews, ~ read_html(paste0(here::here("data", "raw", "apple", "own_scraper_2019-03-20_page_sources/"), .x, ".xml"))),
  review_divs = map(html, ~ html_nodes(.x, ".l-row .we-customer-review")),
  reviews = map(review_divs, iterate_over_review_divs)
) %>% 
  select(app_id, reviews) %>% 
  unnest(cols = c(reviews))

```

## Merge reviews w/ github scraper
Finally, we take reviews from what was scraped with the ordinary scraper from github, and merge with those we scraped ourselves.

```{r}
apple_reviews_all <- reviews_visible_on_store_pages %>% 
  mutate(rating = str_extract(rating, "\\d") %>% as.integer()) %>% 
  bind_rows(apple_review_tibble %>% select(-num_reviews))

```

## Clean up and save out unique reviews

```{r}
apple_reviews_all %>% 
  # merge the title of the apple reviews into the text, then remove line breaks
  mutate(user_name = str_remove(user_name, "\n"),
         review_text = str_c(review_title, review_text, sep = ". ") %>% str_replace_all("\n", "")) %>% 
  select(-review_title) %>% 
  # turn the rating into an integer
  mutate(rating = str_replace(rating, "out of 5", "") %>% as.integer()) %>% 
  # change date format
  mutate(date = lubridate::dmy(date)) %>% 
  distinct() %>% 
  write_csv(here::here("data", "processed", "apple", "2019-03-21_apple_reviews_unique.csv"))
```

### 2022-03-08 02-wrangling_reviews.Rmd
first commit

---
title: "Wrangling - combine reviews"
author: "Ulrik Lyngs"
date: '2022-03-04'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
```

# Combine the reviews

```{r, message=FALSE}
# read in
chrome_reviews <- read_csv(here::here("data", "processed", "chrome", "2019-02-08-chrome_reviews_unique.csv")) %>% 
  mutate(store = "chrome")
play_reviews <- read_csv(here::here("data", "processed", "play", "2019_03_19-play_reviews_unique.csv")) %>% 
  mutate(store = "play")
apple_reviews <- read_csv(here::here("data", "processed", "apple", "2019-03-21_apple_reviews_unique.csv")) %>% 
  mutate(store = "apple")

# counts
chrome_review_num <- chrome_reviews %>% filter(type == "review") %>% nrow()
chrome_reply_num <- chrome_reviews %>% filter(type == "reply") %>% nrow()

apple_reply_num <- apple_reviews %>% filter(!is.na(dev_response_date)) %>% nrow()

total_reviews_and_replies <- nrow(chrome_reviews) + nrow(play_reviews) + nrow(apple_reviews) + apple_reply_num

```

Our own scrapers also collected replies to the reviews (at the time of scraping, developers could reply to a review on the Apple App Store; anyone could reply to a review on the Chrome Web Store).
In total, we collected `r total_reviews_and_replies` reviews and replies; `r nrow(chrome_reviews)` from the Chrome Web store (`r chrome_review_num` reviews and `r chrome_reply_num` replies), `r nrow(apple_reviews) + apple_reply_num` from the Apple App store (`r nrow(apple_reviews)` reviews and `r apple_reply_num` developer replies), and `r nrow(play_reviews)` reviews from the Google Play store.

```{r}
# join up
all_reviews <- chrome_reviews %>% 
  filter(type == "review") %>% 
  select(tool_id = extension_url, user_name, date = date_modified, rating, review_text = text, store) %>% 
  bind_rows(play_reviews %>% rename(review_text = text, tool_id = app_id)) %>% 
  bind_rows(apple_reviews %>% select(tool_id = app_id, user_name, date, rating, review_text, store))

# get rid of reviews that only differ in their posting date -- i.e. someone might have posted the same review on different dates
all_reviews_distinct <- all_reviews %>% 
  distinct(across(-date))

review_count_by_store <- all_reviews_distinct %>% count(store)

review_posting_dates <- all_reviews %>% 
  summarise(first = min(date, na.rm = TRUE),
            last = max(date, na.rm = TRUE))
```

For our analyses, we excluded review replies, as well as reviews that differed only in their posting date (a small number of reviews were posted multiple times by the same reviewer), leaving us with a final dataset of `r nrow(all_reviews_distinct)` reviews 
(Chrome Web: `r review_count_by_store %>% filter(store == "chrome") %>% pull(n)`, 
Play: `r review_count_by_store %>% filter(store == "play") %>% pull(n)`,
Apple: `r review_count_by_store %>% filter(store == "apple") %>% pull(n)`), 
posted between `r review_posting_dates$first` and `r review_posting_dates$last`.

# Simplify tool id for the chrome web store and save our
```{r}
# join it up with the reviews and save out
all_reviews_distinct %>%
  # simplify toolid for chrome web extensions
  mutate(tool_id = if_else(store == "chrome", 
                           str_remove(tool_id, "\\?hl=.+") %>%  str_extract("/[^/]+$") %>% str_remove("/"),
                           tool_id)) %>% 
  relocate(tool_id, store) %>% 
  write_csv(here::here("data", "processed", "distinct_reviews.csv"))

```

### 2022-03-08 03-wrangling_metadata.Rmd
first commit

---
title: "Wrangling - meta data"
author: "Ulrik Lyngs"
date: '2022-03-04'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(jsonlite)
```

# Combine the meta data
## Read in
### Chrome & Apple
These are straight-forward for us to read, we simply do it:

```{r, message=FALSE}
# chrome
chrome_tools <- read_csv(here::here("data", "processed", "chrome", "2019-03-21-chrome_meta_data.csv")) %>% 
  mutate(store = "chrome")

apple_tools <- read_csv(here::here("data", "processed", "apple", "2019-03-20-app_info_extracted_from_page_sources.csv")) %>% 
  mutate(store = "apple")
```

### Play
From the below, the apps that were unavailable will have 'null' as their title. 
We'll exclude them, then afterwards we'll unlist the list columns.

```{r}
# play store
play_tools_raw <- read_json(here::here("data", "raw", "play", "2019-03-19_play_app_info.json"))

play_tools_w_nulls <- tibble(
  app = map_chr(play_tools_raw, "app"),
  app_header_title = map(play_tools_raw, "results") %>% map("title"), #note, is null if couldn't scrape
  app_header_subtitle = NA,
  price = map(play_tools_raw, "results") %>% map("price"),
  company = map(play_tools_raw, "results") %>% map("developer"),
  rating_and_count = NA,
  summary = map(play_tools_raw, "results") %>% map("summary"),
  description = map(play_tools_raw, "results") %>% map("description"),
  version = map(play_tools_raw, "results") %>% map("version"),
  average_rating = map(play_tools_raw, "results") %>% map("score"),
  num_ratings = map(play_tools_raw, "results") %>% map("ratings"),
  num_reviews = map(play_tools_raw, "results") %>% map("reviews"),
  num_5_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("5"),
  num_4_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("4"),
  num_3_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("3"),
  num_2_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("2"),
  num_1_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("1"),
  min_installs = map(play_tools_raw, "results") %>% map("minInstalls")
)

play_tools <- play_tools_w_nulls %>% 
  filter(!map_lgl(app_header_title, is.null)) %>%  # because app_header_title is a list column, we need to use map
  mutate_if(is.list, unlist) %>% 
  mutate(store = "play")
```

## Combine
```{r}
apple_merge <- apple_tools %>% 
  mutate(num_reviews = NA_integer_,
         min_installs = NA_integer_,
         num_users = NA_integer_) %>% 
  select(store, tool_id = app, title = app_header_title, company, average_rating, num_ratings, num_reviews, min_installs, num_users)

play_merge <- play_tools %>% 
  mutate(num_users = NA_integer_) %>% 
  select(store, tool_id = app, title = app_header_title, company, average_rating, num_ratings, num_reviews, min_installs, num_users)

chrome_merge <- chrome_tools %>% 
  mutate(company = str_replace(company, "offered by", "")) %>% 
  # use the unique identifier in the URL as tool id
  mutate(tool_id = str_extract(extension_url, "/[^/]+$") %>% str_remove("/")) %>% 
  mutate(num_reviews = NA_integer_,
         min_installs = NA_integer_) %>% 
  select(store, tool_id, title = extension_header_title, company, average_rating, num_ratings, num_reviews, min_installs, num_users, company)

tool_info <- apple_merge %>% 
  bind_rows(play_merge) %>% 
  bind_rows(chrome_merge)

```

## Add Lyngs et al.'s feature categorisation
```{r}
# get the original feature coding from Lyngs et al, and create a column with its main types of design patterns in alphabetical order
original_tools_w_alphabetical_categories <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv")) %>% 
  select(tool_id, url, `main category`:`main category 4`) %>% 
  unite(categories, c(`main category`,`main category 2`,`main category 3`,`main category 4`), sep = " + ") %>% 
  mutate(categories = categories %>% str_replace_all(" \\+ NA", "")) %>%   #unite doesn't have a drop.na option
  mutate(categories_alphabetically = str_split(categories, " \\+ ")) %>% 
  rowwise() %>%
  mutate(num_categories = length(categories_alphabetically)) %>%
  mutate(categories_alphabetically = categories_alphabetically %>% str_sort() %>% str_c(collapse = " + ")) %>%
  ungroup() %>% 
  select(-categories)

tool_info %>% 
  left_join(original_tools_w_alphabetical_categories) %>% 
  relocate(title, tool_id, store) %>% 
  write_csv(here::here("data", "processed", "meta_data_all_w_design_pattern_types.csv"))
```

### 2022-03-08 figures/figures.Rmd
first commit

---
title: "Figure code"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(ggtextures)
```

```{r read_data}
meta_data <- read_csv(here::here("data", "processed", "meta_data_all_w_design_pattern_types.csv"))
original_tools <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv"))
```

```{r}
### try and draw the apple one again with color
tools_w_categories_and_image <- meta_data %>% 
  mutate(image_single = case_when(
    categories_alphabetically == "Block/removal" ~ "1_block",
    categories_alphabetically == "Goal-advancement" ~ "1_goal",
    categories_alphabetically == "Self-tracking" ~ "1_track",
    categories_alphabetically == "Reward/punish" ~ "1_reward",
    categories_alphabetically == "Block/removal + Goal-advancement" ~ "2_block_goal",
    categories_alphabetically == "Block/removal + Reward/punish" ~ "2_block_reward",
    categories_alphabetically == "Block/removal + Self-tracking" ~ "2_block_track",
    categories_alphabetically == "Goal-advancement + Reward/punish" ~ "2_goal_reward",
    categories_alphabetically == "Goal-advancement + Self-tracking" ~ "2_track_goal",
    categories_alphabetically == "Reward/punish + Self-tracking" ~ "2_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Self-tracking" ~ "3_block_track_goal",
    categories_alphabetically == "Block/removal + Reward/punish + Self-tracking" ~ "3_block_track_reward",
    categories_alphabetically == "Goal-advancement + Reward/punish + Self-tracking" ~ "3_goal_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Reward/punish + Self-tracking" ~ "4_all",
  )) %>% 
  mutate(image_single = str_c(here::here("figures", "plot_colours/"), image_single, ".png")) %>% 
  select(tool_id, title, num_ratings, min_installs, num_users, store, image_single)

```

# Plot of tools by user numbers
generate apple plot
```{r}
apple_data <- tools_w_categories_and_image %>% 
  filter(store == "apple") %>% 
  filter(num_ratings > 0)

images <- apple_data$image_single %>% 
  set_names(apple_data$title)

images_fill = c(
  chrome =  here::here("figures", "textures", "fillpatterns3.gif"), #"https://www.pyngl.ucar.edu/Graphics/Images/fillpatterns3.gif",
  play = here::here("figures", "textures", "siding1-256.jpg"),  #"http://www.hypergridbusiness.com/wp-content/uploads/2012/12/siding1-256.jpg",
  apple = here::here("figures", "textures", "concrete1-256.jpg") # "http://www.hypergridbusiness.com/wp-content/uploads/2012/12/concrete1-256.jpg"
)

apple_features <- apple_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_ratings), y = -1000, image = title), stat = "identity", width = 0.8) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    #scale_y_log10() +
    guides(image = "none") +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(0.6, 0, 0.8, 0, "cm"))

apple_bar_chart <- apple_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_ratings), y = num_ratings, image = store), stat = "identity", width = 0.8) +
    scale_image_manual(values = images_fill) +
    labs(x = "", y = "Number of ratings (log scale)") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    scale_y_log10(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = FALSE) +
    theme_minimal() +
    theme(plot.margin = margin(0.6, 0, 0, 0, "cm")) +
    coord_flip()

apple_combined <- plot_grid(apple_bar_chart, apple_features, rel_widths = c(4,1), labels = c("Apple App", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1)

save_plot(here::here("figures", "apple_ratings_and_features.png"), test_apple,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

generate google play plot

```{r}
num_tools_play_chrome <- 25

##### create the google play plot
play_data <- tools_w_categories_and_image %>% 
  filter(store == "play") %>% 
  arrange(desc(min_installs)) %>%
  slice(1:num_tools_play_chrome)

##### create the google play plot
images <- play_data$image_single %>% 
  set_names(play_data$title)

play_feature_plot <- play_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, min_installs), y = -1000000, image = title), stat = "identity", width = 0.8) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = FALSE) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(0.6, 0, 0.8, 0, "cm"))

play_install_bar <- play_data %>%
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, min_installs), y = min_installs, image = store), stat = "identity", width = 0.8) +
    scale_image_manual(values = images_fill) +
    labs(x = "", y = "Minimum number of installs (log scale)") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    scale_y_log10(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = "none") +
    theme_minimal() +
    theme(plot.margin = margin(0.6, 0, 0, 0, "cm")) +
    coord_flip()

(play_combined <- plot_grid(play_install_bar, play_feature_plot, rel_widths = c(4,1), labels = c("Google Play", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1))

save_plot(here::here("figures", "play_installs_and_functionality.png"), play_combined,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

generate chrome web plot

```{r}
##### create the chrome web plot
chrome_data <- tools_w_categories_and_image %>% 
  filter(store == "chrome") %>% 
  arrange(desc(num_users)) %>% 
  slice(1:num_tools_play_chrome)

##### create the chrome extensions plot
images <- chrome_data$image_single %>% 
  set_names(chrome_data$title)

chrome_feature_plot <- chrome_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_users), y = -200000, image = title), stat = "identity", width = 0.8) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = FALSE) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(0.6, 0, 0.8, 0, "cm"))

# IF WE WANT THE FILL TO BE THE SAME AS FOR RATINGS
chrome_user_bar <- chrome_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_users), y = num_users, image = store), stat = "identity", width = 0.8) +
    scale_image_manual(values = images_fill) +
    labs(x = "", y = "Number of users (log scale)") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    scale_y_log10(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = "none") +
    theme_minimal() +
    theme(plot.margin = margin(0.6, 0, 0, 0, "cm")) +
    coord_flip()

chrome_combined <- plot_grid(chrome_user_bar, chrome_feature_plot, rel_widths = c(4,1), labels = c("Chrome Web", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1)

save_plot(here::here("figures", "chrome_users_and_functionality.png"), chrome_combined,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

put them all together

```{r}
apple <- ggdraw() + draw_image(here::here("figures", "apple_ratings_and_features.png")) + theme(plot.margin = unit(c(0.2,0,0,0.2), "cm"))
play <- ggdraw() + draw_image(here::here("figures", "play_installs_and_functionality.png")) + theme(plot.margin = unit(c(0.2,0,0,0.2), "cm"))
chrome <- ggdraw() + draw_image(here::here("figures", "chrome_users_and_functionality.png")) + theme(plot.margin = unit(c(0.2,0,0,0.2), "cm"))
legend <- ggdraw() + draw_image(here::here("figures", "legend_horisontal.png"), scale = 0.8) + theme(plot.margin = unit(c(-8,0,0,0.5), "cm"))

user_numbers_combined_plot <- plot_grid(chrome, play, apple, legend, ncol=1)

save_plot(here::here("figures", "user_numbers_combined_plot.png"), user_numbers_combined_plot,
          ncol = 1, # we're saving a grid plot of 2 columns
          nrow = 4, # and 2 rows
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

# Plot of average ratings by design pattern types

```{r}
ratings_more_than <- 14

ave_ratings_table <- meta_data %>%
  filter(num_ratings > ratings_more_than) %>%
  group_by(categories_alphabetically) %>%
  summarise(num_tools = n(),
            median_rating = median(average_rating, na.rm = TRUE) %>% round(2),
            #mean_rating = mean(average_rating, na.rm = TRUE),
            min_rating = min(average_rating, na.rm = TRUE) %>% round(2),
            max_rating = max(average_rating, na.rm = TRUE) %>% round(2)) %>%
  arrange(desc(median_rating)) %>% 
  left_join(meta_data %>% distinct(categories_alphabetically, num_categories)) %>% 
  arrange(num_categories, desc(median_rating)) %>% 
  select(-num_categories) %>% 
  filter(!is.na(median_rating)) %>% 
  mutate(image_single = case_when(
    categories_alphabetically == "Block/removal" ~ "1_block",
    categories_alphabetically == "Goal-advancement" ~ "1_goal",
    categories_alphabetically == "Self-tracking" ~ "1_track",
    categories_alphabetically == "Reward/punish" ~ "1_reward",
    categories_alphabetically == "Block/removal + Goal-advancement" ~ "2_block_goal",
    categories_alphabetically == "Block/removal + Reward/punish" ~ "2_block_reward",
    categories_alphabetically == "Block/removal + Self-tracking" ~ "2_block_track",
    categories_alphabetically == "Goal-advancement + Reward/punish" ~ "2_goal_reward",
    categories_alphabetically == "Goal-advancement + Self-tracking" ~ "2_track_goal",
    categories_alphabetically == "Reward/punish + Self-tracking" ~ "2_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Self-tracking" ~ "3_block_track_goal",
    categories_alphabetically == "Block/removal + Reward/punish + Self-tracking" ~ "3_block_track_reward",
    categories_alphabetically == "Goal-advancement + Reward/punish + Self-tracking" ~ "3_goal_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Reward/punish + Self-tracking" ~ "4_all",
  )) %>% 
  mutate(image_single = str_c(here::here("figures", "plot_colours/"), image_single, ".png"))

ave_ratings_levels <- ave_ratings_table$categories_alphabetically

# now create the feature plot
images <- ave_ratings_table$image_single %>% 
  set_names(ave_ratings_table$categories_alphabetically)

feature_plot_compare_types <- ave_ratings_table %>% 
  ungroup() %>% 
  mutate(categories_alphabetically = factor(categories_alphabetically, levels = rev(ave_ratings_levels), ordered = TRUE)) %>% 
  ggplot() +
    geom_textured_bar(aes(x = categories_alphabetically, y = -0.9, image = categories_alphabetically), stat = "identity", width = 0.75) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    guides(image = FALSE) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(1.5, 0, 0.82, 0, "cm"))

# create the boxplot
# with store
boxplot_ratings <- meta_data %>%
  filter(num_ratings > ratings_more_than) %>%
  mutate(categories_alphabetically = factor(categories_alphabetically, levels = rev(ave_ratings_levels), ordered = TRUE)) %>%
  ggplot(aes(x = categories_alphabetically, y = average_rating)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.3, height = 0.1, size = 1.1, alpha = 0.7) +
    theme_minimal() +
    labs(x = "", y = "") +
    coord_flip() +
    theme(axis.text.y=element_blank()) +
    theme(plot.margin = margin(1.5, 0, 0, 0, "cm")) +
    scale_y_continuous(limits = c(1,5))

ratings_by_type <- plot_grid(boxplot_ratings, feature_plot_compare_types, rel_widths = c(3,1.2), labels = c("Average rating", "Features"), hjust = -0.14)

save_plot(here::here("figures", "ratings_by_type.png"), ratings_by_type,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )

ggdraw() + 
  draw_image(here::here("figures", "ratings_by_type.png")) +
  annotate(
    "text", label = "Block/\nremoval",
    x = 0.765, y = 0.84, size = 2.8, colour = "black"
  ) +
  annotate(
    "text", label = "Self-\ntracking",
    x = 0.83, y = 0.84, size = 2.8, colour = "black"
  ) +
  annotate(
    "text", label = "Goal-\nadvance",
    x = 0.895, y = 0.84, size = 2.8, colour = "black"
  ) +
  annotate(
    "text", label = "Reward/\npunish",
    x = 0.955, y = 0.84, size = 2.8, colour = "black"
  )

ggsave(here::here("figures", "ratings_by_type_annotated.pdf"), height = 4.3, width = 7)
```

# Wordcloud for reviews
```{r}
reviews <- read_csv(here::here("data", "processed", "distinct_reviews.csv"))

library(tidytext)
library(SnowballC)
library(wordcloud2)

data(stop_words)

tidy_review_words <- reviews %>% 
  unnest_tokens(word, review_text) %>% 
  anti_join(stop_words) %>% 
  filter(!is.na(word))

tidy_review_words %>%
  count(word, sort = TRUE) %>% 
  filter(!word %in% c("app", "apps", "phone", "extension")) %>% 
  wordcloud2(size = 1)

# tidy_review_words %>% 
#   mutate(word = case_when(
#     str_detect(word, "^love") ~ "love",
#     str_detect(word, "^block") ~ "block", 
#     str_detect(word, "^addict") ~ "addiction", 
#     str_detect(word, "^productiv") ~ "productivity", 
#     str_detect(word, "^wast") ~ "waste", 
#     str_detect(word, "track") ~ "tracking", 
#     str_detect(word, "^notification") ~ "notification", 
#     str_detect(word, "^feature") ~ "feature", 
#     str_detect(word, "^distract") ~ "distract", 
#     str_detect(word, "^focus") ~ "focus", 
#     word %in% c("recommend", "recommended") ~ "recommend", 
#     word == "use" ~ "use/usage", 
#     word == "usage" ~ "use/usage", 
#     TRUE ~ word
#   )) %>% 
#   count(word, sort = TRUE) %>% 
#   filter(!word %in% c("app", "apps", "phone", "extension")) %>% 
#   wordcloud2(size = 1)

# tidy_review_words %>% 
#   mutate(word = wordStem(word)) %>% 
#   count(word, sort = TRUE) %>% 
#   filter(!word %in% c("app", "apps", "phone", "extension")) %>% 
#   wordcloud2(size = 1)
```

### 2022-03-08 materials/scrapers/apple-app-store/apple_scraper.Rmd
first commit

---
title: "Apple app meta data extraction"
author: "Ulrik Lyngs"
date: "18 Feb 2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(magrittr)
library(lubridate)
library(xml2)
library(here)
library(jsonlite)
```

This sript was originally used to scrape tools from the Apple App store on 19 March 2019.

# Scraping with rvest
## Get page sources
### Rescrape
If you want to rescrape the Apple tools from the store, run this

```{r, warning=FALSE, eval=FALSE}
tools_to_scrape <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv")) %>% 
  filter(store == "Apple App Store") %>% 
  select(title, url, appId)

safe_read <- safely(read_html)

tools_with_html <- tools_to_scrape %>% 
  mutate(html = map(url, safe_read)) %>% 
  mutate(is_ok = map_lgl(html, ~ is_null(.$error))) %>%  # add column saying whether the scrape worked
  filter(is_ok == TRUE) %>% 
  mutate(html = map(html, "result"))

# create a folder to store the page sources in
dir.create(here("data", "raw", "apple", str_c("own_scraper_", today(), "page_sources")))

# save them out
tools_with_html %>% 
  select(html, appId) %>% 
  pwalk(~write_xml(.x, str_c(str_c(here("data", "raw", "apple", str_c("own_scraper_", today(), "page_sources/")), .y, ".xml"))))

```

### Use original data
If you want to do this with the original data from March 2019, run this

```{r, warning=FALSE}
tools_with_html <- tibble(
  app_id = list.files(here("data", "raw", "apple", "own_scraper_2019-03-20_page_sources")) %>% str_remove("\\.xml$"),
  html_paths = list.files(here("data", "raw", "apple", "own_scraper_2019-03-20_page_sources"), full.names = TRUE)
) %>% 
  mutate(html = map(html_paths, read_html))

```

## Extract app info
We write a function for extracting the right data from a store page html

```{r}
extract_star_proportion <- function(numStar, html){
  child_order <- 6 - numStar
  html_node(html, paste0('.we-star-bar-graph__row:nth-child(', child_order, ')')) %>% html_node('.we-star-bar-graph__bar__foreground-bar') %>% as.character() %>% str_extract("\\d+%")
}

safe_extract_star_proportion <- possibly(extract_star_proportion, otherwise = "no_ratings")

extract_data <- function(html, app_id){
  tibble(
    app = app_id,
    app_header_title = html_node(html, '.app-header__title') %>% html_text(),
    app_header_subtitle = html_node(html, '.app-header__subtitle') %>% html_text(),
    price = html_node(html, '.app-header__list__item--price') %>% html_text(),
    company = html_node(html, '.app-header__identity') %>% html_text(),
    rating_and_count = html_node(html, '.star-rating__count') %>% html_text(),
    truncated_description = html_node(html, '.we-truncate') %>% html_text(),
    version = html_node(html, '.whats-new__latest__version') %>% html_text(),
    average_rating = html_node(html, '.we-customer-ratings__averages__display') %>% html_text(),
    num_ratings = html_node(html, '.we-customer-ratings__count') %>% html_text(),
    prop_stars_5 = safe_extract_star_proportion(5, html),
    prop_stars_4 = safe_extract_star_proportion(4, html),
    prop_stars_3 = safe_extract_star_proportion(3, html),
    prop_stars_2 = safe_extract_star_proportion(2, html),
    prop_stars_1 = safe_extract_star_proportion(1, html)
  )
}

```

Then we apply it.

```{r}
meta_data <- tools_with_html %>% 
  select(html, app_id) %>% 
  pmap_dfr(~extract_data(.x, .y))

```

Now let's do a little additional cleaning and save it out

```{r}
clean_title <- function(title_raw){
  title_raw %>% 
    str_replace_all("[\n]","") %>% 
    str_replace("4\\+", "") %>% 
    str_trim()
}

clean_company <- function(company){
  company %>% 
    str_replace_all("[\n]","") %>% 
    str_trim()
}

clean_meta_data <- meta_data %>% 
  mutate(app_header_title = map_chr(app_header_title, clean_title),
         company = map_chr(company, clean_company),
         version = str_replace(version, "Version ", ""), 
         num_ratings = str_replace(num_ratings, "Ratings", "") %>% str_replace("K", "00") %>% str_replace("\\.", "") %>% parse_integer(),
         num_ratings = if_else(prop_stars_1 == "no_ratings", as.integer(0), num_ratings))

write_csv(clean_meta_data, here("data", "processed", "apple", "2019-03-20-app_info_extracted_from_page_sources.csv"))

```

### 2022-03-08 materials/scrapers/chrome-web/chrome_ext_info_scraper.Rmd
first commit

---
title: "Chrome extensions info scraper"
author: "Ulrik Lyngs"
date: "21 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(lubridate)
library(here)
library(xml2)
library(janitor)
```

This sript was originally used to scrape tools from the Apple App store on 19 March 2019.

# Scraping with rvest
## Get page sources

### Rescrape
If you want to rescrape the tools from the store, run this

```{r, eval=FALSE}
tools_to_scrape <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv")) %>% 
  filter(store == "Chrome Web Store") %>% 
  select(title, url) %>% 
  mutate(filename = str_c("ext", row_number(), "_", make_clean_names(title)))

safe_read <- safely(read_html)

tools_with_html <- tools_to_scrape %>% 
  mutate(html = map(url, safe_read)) %>% # SCRAPE
  mutate(is_ok = map_lgl(html, ~ is_null(.$error))) %>% # add column saying whether the scrape worked
  filter(is_ok == TRUE) %>% # only the keep the ones that were available
  mutate(html = map(html, "result")) %>% # grab just the html part

# create a folder to store the page sources in
dir.create(here("data", "raw", "chrome", str_c(today(), "-page_sources")))

# save them out
tools_with_html %>% 
  select(html, filename) %>% 
  pwalk(~write_xml(.x, str_c(str_c(here("data", "raw", "chrome", str_c(today(), "-page_sources_info/")), .y, ".xml"))))

```

### Use original data
If you want to do this with the original data from March 2019, run this

```{r, warning=FALSE}
tools_with_html <- tibble(
  filename = list.files(here("data", "raw", "chrome", "2019-03-21-page_sources_info")) %>% str_remove("\\.xml$"),
  html_paths = list.files(here("data", "raw", "chrome", "2019-03-21-page_sources_info"), full.names = TRUE)
) %>% 
  mutate(html = map(html_paths, read_html)) %>% 
  select(-html_paths)

```

## Extract app info

```{r}
extract_data <- function(html){
  tibble(
    extension_url = html_node(html, '[property="og:url"]') %>% html_attr("content"),
    extension_header_title = html_node(html, '.e-f-w') %>% html_text(),
    company = html_node(html, '.e-f-Me') %>% html_text(),
    summary = html_node(html, '.C-b-p-j-Pb') %>% html_text(),
    description = html_node(html, '.C-b-p-j-Oa') %>% html_text(),
    version = html_node(html, '.C-b-p-D-Xe.h-C-b-p-D-md') %>% html_text(),
    average_rating = html_node(html, '.q-N-nd') %>% html_attrs() %>% enframe() %>% filter(name == 'aria-label') %>% pull(value),
    num_ratings = html_node(html, '.q-N-nd') %>% html_attrs() %>% enframe() %>% filter(name == 'aria-label') %>% pull(value),
    num_users = html_node(html, '.e-f-ih') %>% html_text()
  )
}

# get the meta data
extension_info <- map_dfr(tools_with_html$html, extract_data)

```

Now let's do a bit of cleaning of the data and save it out

```{r}
info_cleaned <- extension_info %>% 
  mutate(average_rating = str_extract(average_rating, "\\d[^o]+") %>% parse_number(),
         num_ratings = str_extract(num_ratings, "[^\\.]+ user") %>% str_replace("user", "") %>% str_replace("One", "1") %>% parse_number(),
         num_users = str_replace(num_users, "user", "") %>% str_replace("s", "") %>% parse_number(),
         company = str_replace(company, "offered by ", "") %>% str_replace("\n", ""))

info_cleaned %>% 
  write_csv(here("data", "processed", "chrome", "2019-03-21-chrome_meta_data.csv"))

```

### 2022-03-08 materials/scrapers/chrome-web/chrome_review_scraper.Rmd
first commit

---
title: "chrome_reviews_scraper"
author: "Ulrik Lyngs"
date: "05/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(robotstxt)
library(jsonlite)
library(tidyverse)
library(rvest)
library(RSelenium)
library(here)
library(RCurl)
```

# Rvest <3 RSelenium
## Websites' own stated rules - robots.txt
From a research ethics perspective, we should start by inspecting **robots.txt**, which is a file most websites contain saying how they want others to behave when accessing the website using robots.

```{r}
get_robotstxt(domain = "chrome.google.com/")

paths_allowed("https://chrome.google.com/webstore/detail/")

```

So we see that Google actually don't want us to search through the Chrome Web Store with automated scrapers, but everything else should be fair game.

We also add delays when we crawl so that we're not mistaken for a DDoS attack.

We'll do this with `Sys.sleep()`, e.g. `Sys.sleep(5)` to add a 5-second delay. However, using a constant might also get you recognised as a robot. We could use e.g. `Sys.sleep(runif(n = 1, min = 5, max = 10))` to add a random sleep between 5 and 10 seconds.

## Scraping with RSelenium
### Getting things up and running

1. install RSelenium
```{r}
# to run RSelenium we need these:
# devtools::install_github("johndharrison/binman")
# devtools::install_github("johndharrison/wdman")
# devtools::install_github("ropenSci/RSelenium")
```

2. Get Docker up and running
  - install it from https://docs.docker.com/docker-for-mac/install/
  - Open the app (on first run you will need to accept its terms and conditions)
  - then in a terminal start a chrome server with `docker run --name chrome -d -p 4445:4444 selenium/standalone-chrome`
  - there's some test things shown here that you can run (commented out)

```{r, eval=FALSE}
# in a terminal run: docker run --name chrome -d -p 4445:4444 selenium/standalone-chrome
# this will start up a server named 'chrome'
# to stop it you go 'docker stop chrome'
# to see what's currently running, go 'docker ps'
# to delete a server instance name chrome, do 'docker rm chrome'

# -name names your container
# -v mount volume
# -d detached mode (if you don't do this, it'll actually enter the image rather than just)
# -p port mapping (selenium connects to port 4444 by default; if you go like 4445:4444 then you can connect to 4445 on the outside which is mapped to 4444; with -p 5901:5900 you can open up ports for viewers)

# remDr <- remoteDriver(port=4445L, browserName = "chrome")
# remDr$open()
# remDr$getStatus()
# remDr$navigate("https://www.google.com/")
# remDr$getCurrentUrl()
# remDr$navigate("https://www.bing.com/")
# remDr$getCurrentUrl()
# remDr$goBack()
# remDr$getCurrentUrl()
# remDr$goForward()
# remDr$getCurrentUrl()
# remDr$close()
```

### Create function to scrape the sources of review pages

```{r, echo=TRUE, eval=FALSE, results='hide'}
scrape_review_pages <- function(url_string){
  # navigate to url
  print(str_c("navigating to ", url_string))
  remDr$navigate(url_string)
  Sys.sleep(1) # wait for one second
  print(remDr$getCurrentUrl())
  remDr$screenshot(display = TRUE) # show screenshot

  # click 'reviews' button
  reviewButton <- remDr$findElement(using = 'class', "h-e-f-z-b")
  Sys.sleep(0.5)
  reviewButton$clickElement()
  remDr$screenshot(display = TRUE) # show screenshot

  # save the source
    # we don't know how many reviews there are and because
    # speed is not really a concern, we won't preallocate output length
  n <- 1
  reviews <- list()

  reviews[[n]] <- remDr$getPageSource()

  # while there is a visible 'next' button, click it and save the information
  nextButton <- remDr$findElement(using = 'css', "a.Aa.dc-se")
  Sys.sleep(0.5)
  remDr$executeScript("arguments[0].scrollIntoView(true);", args = list(nextButton)) # scroll to it
  remDr$screenshot(display = TRUE)

  while (nextButton$isElementDisplayed()[[1]]) {
    nextButton$clickElement()
    remDr$screenshot(display = TRUE) # show screenshot

    # store the page
    n <- n + 1
    print(stringr::str_c("storing page ", n))
    reviews[[n]] <- remDr$getPageSource()

    print(str_c("can we see a 'next' button?", nextButton$isElementDisplayed()[[1]]))

    print("sleeping for one second")
    Sys.sleep(1)
  }

  return(reviews)
}
```

### Create function to iterate over list of titles and urls and scrape their sources

```{r}
# make helper function to strip out special characters from titles
remove_special_chars <- function(annoying_title) {
  str_replace_all(annoying_title, "[^[A-Za-z]]", " ") %>% str_replace_all(.,"[ ]+", " ")  
}

# make helper function to name the list with the pages
set_review_page_names <- function(review_pages){
  page_names <- str_c("page ", seq_along(review_pages))

  review_pages %>% 
    set_names(page_names)
}

scrape_and_store_page_sources <- function(title_and_url_tbl) {
  for (i in 1:nrow(title_and_url_tbl)) {
    print(stringr::str_c("there are ", nrow(title_and_url_tbl), " extensions to scrape."))

    print(stringr::str_c("scraping extension number ", i, " titled ", title_and_url_tbl$title[i], " from url ", title_and_url_tbl$url[i]))

    # get the review pages
    review_pages <- scrape_review_pages(title_and_url_tbl$url[i])

    # put the result and its parsing in a tibble
    source_tibble <- review_pages %>% 
      set_review_page_names() %>% 
      enframe() %>% 
      mutate(extension_title = title_and_url_tbl$title[i],
             extension_url = title_and_url_tbl$url[i],
             time_scraped = lubridate::now())

    # store it
    source_tibble %>% 
      write_rds(str_c(here::here("data", "raw", "chrome", str_c(lubridate::today(), "-page_sources_reviews/")), "extension_num_", i, remove_special_chars(title_and_url_tbl$title[i]), ".rds"))
  }
}

```

### Error handling
So we need this to be able to handle errors. Let us try with three extensions - two we know works and one we know doesn't work.

```{r}
# Lights off will give an error - the others won't
test_extensions <- tibble(
  title = c("(NO)tifications", "Lights off", "Afairo"),
  url = c("https://chrome.google.com/webstore/detail/notifications/ocfjkjbhicdjhmgkolejpojieddlogbg?hl=gb", "https://chrome.google.com/webstore/detail/lights-off/ehfaompdnhilblcmkkakofflcjfijhai?hl=gb", "https://chrome.google.com/webstore/detail/afairo/imecoioakflafgnkkklfhopabameijck?hl=gb")
)
```

So rather than running this directly, we'll wrap the scraper function in `safely` to make sure we don't crash when hitting extensions that have become unavailable. See [R4DS 'Dealing with failure'](https://r4ds.had.co.nz/iteration.html#dealing-with-failure).

```{r}
safe_scrape_sources <- safely(scrape_and_store_page_sources)
```

Now let's test it.

```{r}
# TEST: scrape review page sources for forest
forest_reviews <- safe_scrape_sources("https://chrome.google.com/webstore/detail/forest-stay-focused-be-pr/kjacjjdnoddnpbbcjilcajfhhbdhkpgk?hl=gb")
```

### Go ahead and do it!

```{r}
extensions_to_scrape <- read_csv(here("materials", "2018-12-30_lyngs_tool_coding.csv")) %>% 
  filter(is.na(Exclude)) %>% 
  filter(str_detect(url, "chrome.google.com/webstore")) %>% 
  mutate(is_available = url.exists(url)) %>%  # this will take a while
  filter(is_available == TRUE) %>% 
  select(title, url)

scrape_and_store_page_sources(extensions_to_scrape)

```

### 2022-03-08 paper.Rmd
first commit

---
journal: International Journal of Human-Computer Studies
title: "The Goldilocks Level of Friction: Using User Numbers, Ratings, and Reviews to Evaluate Digital Self-Control Tools"
keywords: Attention, self-control, self-regulation, distraction, digital wellbeing, focus
abstract: |
  Much effort has been invested in designing digital systems that keep people 'hooked', but comparatively little is known about how designers can best support people in re-gaining control over digital device use. Online, hundreds of apps and browser extensions promise to help people self-regulate their use, and their reviews and popularity metrics can indicate which design patterns are found useful in the will. Moreover, they reveal how platforms such as Android and iOS differ in the ecosystems they enable for such tools, which has important practical implications for users struggling with self-regulation difficulties. We analysed user reviews, installation numbers, and ratings for 334 digital self-control tools (DSCTs) on the Google Play, Chrome Web, and Apple App stores, investigating what user reviews reveal about usage contexts and key design challenges, and how functionality relates to popularity metrics. Our thematic analysis of 1,529 reviews (sampled from a data set of 54,320 reviews scraped in March 2019) found that DSCTs are seen as highly important for staying focused on less instantly rewarding tasks when digital distractions are easily available. Users seek DSCTs that adapt to their personal definitions of 'distracting use', and provide incentives sufficient to affect behaviour change without feeling overly intrusive. DSCTs' interventions, even if highly useful, can elicit negative emotions, which is an important reason that the goals of the user and the DSCT should align. Tools that combine two or more types of design patterns (e.g., website blocking and goal reminders) received significantly higher ratings than those implementing a single type, which resonated with feature suggestions in the reviews. We discuss implications for research and design, including how basic psychological research on self-regulation can inform evaluation measures for DSCTs, as well as guide future exploration of how design patterns and framings of 'distraction' interact.

bibliography: references.bib

output:
  bookdown::pdf_document2:
    template: elsarticle-template/elsarticle-template.tex
    citation_package: natbib
    keep_tex: true
    latex_engine: xelatex
---

```{r setup, include=FALSE}
library(tidytext)
library(tidyverse)
library(readxl)
library(ngram)
library(scales)
library(ircor)

knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(scipen = 999)
```

# Introduction

The constant connectivity of smartphones and laptops provides innumerable benefits, but also unprecedented opportunities for distraction.
Indeed, as the business models of many tech companies have come to rely on nudging people into using digital services extensively (cf. the 'attention economy' [@davenportAttentionEconomyUnderstanding2001; @einsteinBlackOpsAdvertising2016; @wuAttentionMerchantsEpic2016]), many people routinely struggle to exercise self-control over their device use [@coxDesignFrictionsMindful2016; @koNUGUGroupbasedIntervention2015; @marottaOnlineDistractionsWebsite2017; @tranModelingEngagementDisengagementCycle2019; @hinikerMyTimeDesigningEvaluating2016].
As a result, public debate about negative effects of 'persuasive technologies' has been growing [@OptimizingEngagementUnderstanding2019; @orlowski2020] alongside increased political pressure on technology platforms to safeguard users [@allpartyparliamentarygrouponsocialmediaandyoungpeoplesmentalhealthandwellbeingNewFiltersManageImpact2019].

On online stores, a flourishing market exists for 'digital self-control tools', where hundreds of apps and browser extensions promise to help people take back control over their device use.
These tools offer interventions such as blocking apps or removing distracting website elements, tracking and visualising usage, or providing rewards or punishments for how devices are used [@lyngsSelfControlCyberspaceApplying2019].
Similarly, Apple and Google now offer various usage logging and self-limiting features in their 'ScreenTime' and 'Digital Wellbeing' tools, respectively [@appleIOS12Introduces2018; @gonzalezResearchGoogleNew2018; @googleDigitalWellbeing2018], and even Facebook have introduced tools 'for managing your time on Facebook and Instagram' [@ranadiveNewToolsManage2018].

In academia, a growing amount of HCI research is focusing on developing and evaluating various design patterns meant to help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019]^[We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders), and the term ‘digital self-control tool’ (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook’s newsfeed). We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user’s usual digital environment, or when it is evaluated in a research study.].
This area of HCI work has gradually been building an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how much people vary in how useful they find particular interventions, and how specific design patterns' usefulness depends on details of their implementation, such as the level of 'friction' they introduce [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].

To answer such questions, most HCI studies have taken an experimental approach and compared user behaviour and perceptions when applying a given intervention to its absence and/or some alternative intervention [@lyngs2021].
Yet, while controlled studies are appropriate and feasible for evaluating and comparing a small number of design patterns, they are difficult to scale to broadly assess a large numbers of patterns and implementations.

A complementary approach, which may help scope the range of design patterns and implementations to explore in controlled studies, is to investigate the existing landscape of digital self-control tools on app and web stores [@lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]. 
Here, user reviews and popularity metrics can provide information about the relative utility of different patterns and implementations, contexts of use, and user needs 'in the wild'.
However, few studies to date have taken this approach.
Hence, in the present paper, we contribute to the growing HCI literature on design patterns for digital self-control by using app and web store information as a resource for understanding user self-experimentation with digital self-control tools (DSCTs).
Specifically, we present an analysis of user reviews, functionality, installation numbers, and ratings for 334 digital self-control tools, drawn from Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] who reviewed functionality in such tools on the Google Play, Chrome Web, and Apple App stores.

Our analysis of user reviews finds that digital self-control tools are seen as crucial to support focus on less instantly rewarding goals when digital distractions are easily available, specifically in contexts related to productivity (including work, studying, reading, and writing), and going to sleep.
DSCTs seem especially useful for people who describe themselves as 'addicts' or as struggling with distraction due to conditions like ADHD.

Users want DSCTs to flexibly adapt to their personal definition of ‘distraction’, as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
A key reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions. This may cause the effects of interventions to be counterproductive, especially when users perceive DSCTs' goals or framing to differ from their own.

Finally, users want DSCTs to provide carefully calibrated incentives that are sufficient to incentivise intended behaviour change without feeling overly intrusive.
In our analysis, this 'Goldilocks level'^[After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'. The concept of the 'just right amount' is often referred to as the 'Goldilocks principle'. For example, a widely cited paper on the relationship between screen time and mental well-being documented that a balance point in between "too much" and "too little" tech use seemed optimal, which they referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017].] varied between reviewers, but they commonly wanted tools to provide incentives powerful enough to keep themselves on track in moments of temptation or absentmindedness.
Overly strong incentives could backfire if they simply made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).

One way for tools to provide properly calibrated incentives might be by carefully combining different types of design patterns (e.g., using both block/removal and goal-advancement features). 
This was often asked for in reviews' feature suggestions, and supported by higher overall ratings given to tools combining two or more types compared to one.
A final consideration related to the amount of effort involved in operating a tool, where low friction to set up and initiate use (e.g., via schedules where a tool automatically turns on), but high friction to override it, was important.

## Motivation and background

In HCI, a range of studies have developed and evaluated design patterns for digital self-control, ranging from ambient visualisations of time spent [@whittakerDonWasteMy2016] and distraction blocking [@kimGoalKeeperExploringInteraction2019] to goal-setting and social support [@koNUGUGroupbasedIntervention2015].
This research is starting to build an evidence base of how specific interventions to support self-regulation influence user behaviour and perceptions, but basic questions remain open.
These include *conceptual challenges* (e.g., what metrics digital self-control tools should seek to optimise, as simply measuring impact on 'screen time' is deemed inadequate by most researchers [@cecchinatoDesigningDigitalWellbeing2019; @hinikerMovingGreatScreen2019; @lukoffDigitalWellbeingWay2019]); related gaps in our understanding of *individual variation in user needs* (such as how how widely people differ in what they consider 'distracting' usage [@cecchinatoDesigningDigitalWellbeing2019; @lukoffDigitalWellbeingWay2019]); and *practical implementation challenges* (such as how much 'friction' tools should apply to hold users accountable for their past preferences [@kimGoalKeeperExploringInteraction2019; @coxDesignFrictionsMindful2016; @lyngsTellMeWhat2018]).

One approach to addressing these questions is to evaluate new or existing interventions in lab and field studies, and/or conduct surveys or interviews with users.
A complementary approach, which  we explore in the present paper, is to investigate the landscape of digital self-control tools available online.
Here, many hundreds of tools are exploring a wide range of strategies, from limiting the amount of functionality available on home screens (e.g.~the Android app \emph{LessPhone Launcher} [@mohanLessPhoneLauncher2019]) and hiding or modifying content on distracting websites (e.g. *Newsfeed Eradicator* [@jdevNewsFeedEradicator2019]), to gamifying self-control by tying device use to the wellbeing of virtual creatures (e.g. *Forest* [@seekrtechForestStayFocused2018]).

Some of these tools have millions of users [@lyngsSelfControlCyberspaceApplying2019], and they effectively represent a myriad of natural studies in which individuals self-experiment with tools that represent not only one or more design patterns, but specific ways to implement those patterns [@daskalovaSleepCoacher2018; @leeSelfExperimentation2017].
Indicators of the outcome of these experiments include user reviews, store ratings, and installation numbers, which may provide valuable information about the usefulness of different design patterns and implementations.

Other areas of HCI research have benefited from taking such an approach.
For example, in mental health research, thematic analysis of user reviews has been used to explore factors important for optimising user experience and support engagement in apps for cognitive behavioural therapy [@bakkerMentalHealthSmartphone2016; @huguetSystematicReviewCognitive2016; @luiEvidencebasedAppsReview2017; @sucalaAnxietyThereApp2017; @stawarzUserExperienceCognitive2018].
Similar research found that content in popular apps marketed as reducing symptoms of anxiety and worry was largely inconsistent with evidence-based treatments [@kertzReviewFreeIPhone2017]. 
This opened avenues for further research including how established CBT practices could more effectively be disseminated and, conversely, what might be learned from alternative approaches that proved popular in the wild.
HCI research on design patterns for digital self-control might in a similar way derive actionable insights from the large number of tools for this purpose available online.

However, few studies on digital self-control have so far taken this route:
To the best of our knowledge, four published studies have described and/or categorised functionality in samples of digital self-control tools available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al. [@vanvelthovenProblematicSmartphoneUse2018] presented aims and features in 21 tools identified via the software recommendation platform [alternativeto.net](https://alternativeto.net); Biskjær et al. [@biskjaerTakingActionDistraction2016] presented a taxonomy of functionality based on 10 tools from the Google Play, Chrome Web, and Apple App stores, and online tech magazines; Roffarello & Russis [@mongeroffarelloRaceDigitalWellbeing2019] presented a taxonomy of functionality based on 42 mobile apps from the Google Play store; and Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] presented a taxonomy of functionality based on 367 tools from the Google Play, Chrome Web and Apple App stores, which was mapped to psychological mechanisms in a dual systems model.

Only one of these studies analysed user reviews, namely Roffarello & Russis [@mongeroffarelloRaceDigitalWellbeing2019], which included thematic analysis of 1,128 reviews from 42 smartphone apps, and none incorporated popularity metrics.
Moreover, Roffarello & Russis' analysis of user reviews was restricted to 42 smartphone apps drawn from the Google Play store and did not specify how the reviews were distributed across apps with different functionality.
Thematic analysis is likely to yield different results depending on how reviews are sampled across tools with different functionality, and depending on the analysis strategy [@chenARminerMiningInformative2014; @braunThematicAnalysis2018]. Therefore, additional analyses with broader sampling of tools, and transparent sampling of reviews from those tools, might be helpful.

On this background, we extended existing studies of publicly available digital self-control tools by 
(i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across stores and types of functionality, 
(ii) analysing review content using 'reflexive' thematic analysis [@braunThematicAnalysis2018], 
(iii) analysing popularity metrics' relationship to tools' design features, and 
(iv) making our materials, analysis pipelines and data openly available for the purpose of transparency and reusability [@harozUpdatesCHISubmission2019].

# Methods

Data, analysis scripts, and materials for the present paper are available at \url{https://osf.io/cdqub/}.

We investigated the 367 digital self-control tools reviewed by Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019], as this paper provided the largest currently available review of DSCTs.
Lyngs et al. identified these tools by downloading search results for the terms 'distraction', 'addiction', 'motivation', and 'self-control', individually as well as prefixed by 'internet', 'smartphone', and 'laptop', on the Chrome Web, Google Play, and Apple App stores.
From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in a data set of 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).^[See Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.]

We adopted Lyngs et al.'s classification of functionality within these tools, which clustered their features into four types of design patterns for digital self-control:

- \emph{block/removal} (features for blocking distractions, such as temporarily locking the user out of specific apps, or for removing them in the first place, such as hiding recommended videos on YouTube)
- \emph{self-tracking} (features for tracking and/or visualising device use, such as showing how much time the user has spent in particular apps or on their devices overall, or letting the user start a countdown timer when they need to focus)
- \emph{goal-advancement} (features for reminding the user of their usage goals, such as displaying pop-ups when a specific amount of time has been spent, or replacing the content of new browser tabs with to-do lists)
- \emph{reward/punishment} (features that add incentives for using devices in particular ways, such as rewarding the user with points/streaks when they do not use their phone, or specific apps)

## Collecting tool information and user reviews

In March 2019, we scraped additional information about these tools, including installation numbers, ratings, and store descriptions, as well as publicly available user reviews.
For apps on the Google Play store, we used a pre-existing script [@olanoGoogleplayscraper2018] to scrape tool information and reviews.
For apps on the Apple App store, we used a pre-existing script [@olanoAppstorescraper2018] to scrape reviews, and used our own script (using the R packages \href{https://rvest.tidyverse.org}{\texttt{rvest}} and \href{http://ropensci.github.io/RSelenium/}{\texttt{RSelenium}}) to scrape tool information.
As the pre-existing script sometimes failed to retrieve all reviews visible on tools' pages on the Apple App store, we wrote an additional R script to scrape reviews when manual inspection found some to be missing.
For browser extensions on the Chrome Web store, we wrote our own R script to scrape both tool information and user reviews.

## Data and analysis approach
### User numbers and ratings
All stores displayed average ratings, as well as a count of how many ratings a tool had received.
For installation numbers, the stores provided different information: the Chrome Web store provided an exact number of users, and the Google Play store provided a 'minimum number of installs' (e.g. '100,000+').
The Apple App store provided no direct information about user numbers.
For processing which ranked tools by installation numbers, we used number of ratings as a proxy for relative numbers of users for tools on the Apple App store.

\subsection{Sampling and thematic analysis of user reviews}

To get a broad coverage of reviews for tools across stores and types of design patterns, we used an R script to systematically sample user reviews for qualitative analysis from the full scraped data.
To ensure that the reviews we analysed came from tools with a minimum level of engagement from users, we excluded reviews from tools on the Google Play and Chrome Web stores that had fewer than 100 installations/users, and tools from the Apple App Store that had fewer than 5 ratings.
To make the sampled reviews more likely to be informative, we also excluded reviews shorter than 10 words.
Our script then iterated through the four types of design patterns described above, on each iteration picking the tool on each store that included that specific type of design pattern and had the highest number of users (Google Play and Chrome Web) or ratings (Apple App store). 
The script then randomly sampled up to 30 user reviews from each of those tools (if a specific tool had fewer than 30 reviews, all of its reviews would be included), and started over. 
The script ran until more than 1,500 reviews were sampled in total. A tool could only be sampled from once.

Next, we conducted inductive thematic analysis of the sampled reviews, following the 'reflexive' approach described in [@braunUsingThematicAnalysis2006] and [@braunThematicAnalysis2018].
Our work was informed by an earlier analysis of 961 reviews randomly sampled from the same data set by the lead author of the present paper [@lyngs2021a, ch. 4].
However, our analysis in the present paper was conducted from scratch, with a new sample of user reviews drawn in the manner described.
First, the lead author (UL) read through all user reviews sampled for thematic analysis and conducted initial coding of recurrent patterns relevant to the research topic.
A second author (LC) read through one third of the reviews and independently coded recurrent patterns of meaning.
Afterwards, these two authors read through the coded excerpts, recoded them, iteratively sorted the codes into potential themes, and discussed emerging themes.

Thematic coding was conducted using [NVivo](https://www.qsrinternational.com/nvivo-qualitative-data-analysis-software/home/) v1.6.1; all quantitative analyses were conducted in R.

# Results

```{r, include=FALSE}
meta_data <- read_csv(here::here("data", "processed", "meta_data_all_w_design_pattern_types.csv"))
original_tools <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv"))

reviews_w_main_cat_data <- read_csv(here::here("data", "processed", "distinct_reviews.csv")) %>% 
  mutate(review_word_count = stringi::stri_count_words(review_text))

(num_missing <- original_tools %>% 
  filter(!(tool_id %in% meta_data$tool_id)) %>% 
  nrow())

(num_w_info_by_store <- meta_data %>% count(store))

## tools w reviews
got_reviews_from <- reviews_w_main_cat_data %>% 
  distinct(tool_id)

(num_tools_w_reviews_by_store <- reviews_w_main_cat_data %>% 
  distinct(tool_id, store) %>% 
  count(store))

(num_reviews_by_store <- reviews_w_main_cat_data %>% 
  count(store))

```

At the time of scraping, `r num_missing` of the tools from Lyngs et al.'s review had become unavailable.
Hence, we scraped information for `r nrow(meta_data)` tools: 
`r num_w_info_by_store %>% filter(store == "chrome") %>% pull(n)` extensions from Chrome Web, 
`r num_w_info_by_store %>% filter(store == "play") %>% pull(n)` apps from Google Play, and 
`r num_w_info_by_store %>% filter(store == "apple") %>% pull(n)` apps from the Apple App store.

`r nrow(got_reviews_from)` of these tools had received user reviews (`r num_tools_w_reviews_by_store %>% filter(store == "chrome") %>% pull(n)` from Chrome Web, `r num_tools_w_reviews_by_store %>% filter(store == "google_play") %>% pull(n)` from Google Play, and `r num_tools_w_reviews_by_store %>% filter(store == "apple") %>% pull(n)` from the Apple App store).
From these, we collected a total of `r format(nrow(reviews_w_main_cat_data), big.mark = ",")` reviews (`r num_reviews_by_store %>% filter(store == "chrome") %>% pull(n) %>% format(big.mark = ",")` from Chrome Web, `r num_reviews_by_store %>% filter(store == "google_play") %>% pull(n) %>% format(big.mark = ",")` from Google Play, and `r num_reviews_by_store %>% filter(store == "apple") %>% pull(n) %>% format(big.mark = ",")` from the Apple App store).

## User reviews

(ref:word-cloud-caption) Word cloud depicting frequencies of terms across all `r format(nrow(reviews_w_main_cat_data), big.mark = ",")` reviews. Font size indicates relative frequency. The word cloud excludes the terms 'app/s', 'phone', and 'extension', as well as stop words (i.e., common English words unlikely to be informative such as "the", "of", "to", etc. Stop words were excluded using the R package `tidytext` [@silge2016tidytext]). Colouring is aesthetic and does not map to any characteristics of the data.

```{r wordcloud, fig.cap="(ref:word-cloud-caption)", out.width='100%'}
knitr::include_graphics("figures/wordcloud.png")
```

```{r, include=FALSE}
(num_reviews_summary <- reviews_w_main_cat_data %>% 
  group_by(tool_id) %>% 
  summarise(num_reviews = n()) %>% 
  summarise(median = median(num_reviews),
            mean = mean(num_reviews),
            min = min(num_reviews),
            max = max(num_reviews),
            IQR = IQR(num_reviews)))

(word_count_summary <- reviews_w_main_cat_data %>% 
  summarise(median = median(review_word_count),
            mean = mean(review_word_count),
            min = min(review_word_count),
            max = max(review_word_count),
            IQR = IQR(review_word_count)))

(ptc_app <- reviews_w_main_cat_data %>% 
  mutate(has_app = str_detect(review_text, "app")) %>% 
  count(has_app) %>% 
  mutate(ptc = (n/sum(n))*100) %>% 
  filter(has_app == TRUE) %>% 
  pull(ptc) %>% 
  round())

(ptc_time <- reviews_w_main_cat_data %>% 
  mutate(has_time = str_detect(review_text, "time")) %>% 
  count(has_time) %>% 
  mutate(ptc = (n/sum(n))*100) %>% 
  filter(has_time == TRUE) %>% 
  pull(ptc) %>% 
  round())
```

For the tools that had received reviews, the median number of reviews was `r num_reviews_summary$median` (min = `r num_reviews_summary$min`, max = `r num_reviews_summary$max`, IQR = `r num_reviews_summary$IQR`), and the median number of words was `r word_count_summary$median` (min = `r word_count_summary$min`, max = `r word_count_summary$max`, IQR = `r word_count_summary$IQR`; reviews containing only emojis have a word count of 0).
Figure \@ref(fig:wordcloud) shows the overall frequency of terms used in the reviews, excluding 'app/s', 'phone' and 'extension' as well as common stop words in English (e.g., 'the', 'of', 'to'; stop words were excluded using the R package `tidytext` [@silge2016tidytext]).
Aside from 'app' (46% of reviews), the most frequent term was 'time', which was included in 17% of all reviews.
The terms most commonly following or preceding 'time' included 'limit', 'usage', 'spent', 'screen', 'waste', and 'management', suggesting that managing time spent on digital devices was a central topic in the reviews.

```{r review-bigrams, eval=FALSE}
review_bigrams <- reviews_w_main_cat_data %>% 
  unnest_tokens(bigram, review_text, token = "ngrams", n = 2)

bigrams_separated <- review_bigrams %>%
  separate(bigram, c("word1", "word2"), sep = " ")

bigrams_filtered <- bigrams_separated %>%
  filter(!word1 %in% stop_words$word) %>%
  filter(!word2 %in% stop_words$word)

bigram_counts <- bigrams_filtered %>% 
  count(word1, word2, sort = TRUE)

```

### Thematic analysis

(ref:ta-reviews-by-design-patterns) Reviews sampled for thematic analysis (total n = 1,529), by design patterns included in the tools.

```{r sampled-reviews}
reviews_for_ta <- read_excel(here::here("data", "processed", "2022-01-22-reviews_for_ta.xlsx"))

num_tools_within_categories <- reviews_for_ta %>% 
  distinct(categories_alphabetically, tool_id) %>% 
  group_by(categories_alphabetically) %>% 
  summarise(num_tools = n())

num_reviews_within_categories <- reviews_for_ta %>% 
  group_by(categories_alphabetically) %>% 
  summarise(num_reviews = n())

# create a table with the count of tools and review from each combination of functionality
num_tools_within_categories %>% 
  left_join(num_reviews_within_categories) %>% 
  mutate(num_categories = str_split(categories_alphabetically, "\\+")) %>% 
  rowwise() %>% mutate(num_categories = length(num_categories)) %>% ungroup() %>% 
  arrange(num_categories, categories_alphabetically) %>% 
  mutate(`Colour code` = categories_alphabetically %>% str_replace("Block/removal", "\\\\typeBlock") %>% str_replace("Reward/punish", "\\\\typeReward") %>% str_replace("Self-tracking", "\\\\typeTrack") %>% str_replace("Goal-advancement", "\\\\typeGoal") %>% str_remove_all(" \\+ ")) %>% 
  select("Design patterns included" = categories_alphabetically,
         "Colour code",
         "Number of tools" = num_tools,
         "Number of reviews" = num_reviews) %>% 
  knitr::kable(caption = "(ref:ta-reviews-by-design-patterns)", escape = FALSE) %>% 
  kableExtra::landscape()

```

We conducted thematic analysis of `r nrow(reviews_for_ta)` reviews from a total of `r reviews_for_ta %>% distinct(tool_id) %>% nrow()` different apps (Table \@ref(tab:sampled-reviews) shows how the reviews were distributed across tools with different combinations of design patterns).
In the following analysis, we accompany review quotes with the name of the tool the review was for, a colour coded indication of which design patterns the tool included, and which store it came from:

- \typeBlock\hspace{1mm}= block/removal, \typeTrack\hspace{1mm}= self-tracking,\
\typeGoal\hspace{1mm}= goal-advancement, \typeReward\hspace{1mm}= reward/punish
- C = Chrome Web, P = Google Play, A = Apple App store
- Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}

Our early codes were 'domain summaries' [@braunThematicAnalysis2018] organised in an evolving hierarchy of sub-codes. For the first coder (UL), early codes were organised in relation to usage contexts and benefits, user characteristics, key design patterns, and praise or criticism of the developer or app. For the second coder (LC), early codes were organised in relation to feature requests, positive feedback, negative feedback, and reasons for use.

An early observation as we explored reported benefits of use was that a substantial proportion of reviews expressed highly positive and important impact of the tools used (n = 480, 22% of reviews).
Thus, the reviews described a common feeling of lacking control over digital device use, which resulted in using devices in unintended ways, or in using them so much that it interfered with other goals.
However, digital self-control tools were seen as effectively helping reviewers take back control, save time, and improve their lives (*"I can happily say that this extension has changed my life!!!! It hides EVERY ANNOYING YouTube feature"*, R385, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock\typeGoal \textbf{C}), and many expressed deep appreciation towards the developers ("Who so ever is working behind this software, my deepest gratitude to them", R271, \href{https://chrome.google.com/webstore/detail/blocksite-block-websites/eiimnmioipafcokbfikbljfdeojpcgbh?hl=en-GB}{\emph{BlockSite}} \typeBlock \textbf{C}). 
Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \textbf{C}).
Importantly, the reviews also contained a wealth of information on how specific implementations of design patterns could fail to fulfil their purpose, and how they might be improved.

```{r, include=FALSE}
theme_one <- "DSCTs as essential for focusing on less instantly rewarding tasks when digital distractions are easily available"
theme_two <- "DSCTs should flexibly adapt to personal definitions of 'distraction'"
theme_three <- "DSCTs should provide incentives that are sufficient to change behaviour without feeling overly intrusive"
```

As we iteratively worked with the codes and excerpts, we developed three broad themes capturing digital self-control tools' (DSCTs) use contexts and design challenges, assembled from recurrent patterns in the early codes:
(i) `r theme_one`,
(ii) `r theme_two`, 
(iii) `r theme_three`

The NVivo files with our early codes, and their later re-organisation by relation to the three main themes, as well as exports of the code hierarchy to other formats, are available on \href{https://osf.io/cdqub/}{osf.io/cdqub/}.

#### **Theme 1: `r theme_one`**

Users saw DSCTs as critical to avoiding time-wasting patterns of habitual or impulsive digital device use.
Many reviews described this challenge as related to 'addictive' design patterns and other adversarial dynamics online, to which DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{Inbox When Ready for Gmail} \typeBlock\typeGoal\typeTrack \textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}). 
DSCTs seemed especially helpful when it was not possible to avoid using distracting services (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, \href{https://chrome.google.com/webstore/detail/kill-news-feed/hjobfcedfgohjkaieocljfcppjbkglfd}{Kill News Feed} \typeBlock \textbf{C}).

The usage contexts mentioned suggested that, aside from wanting to limit time on devices overall or on specific services (n = 62, 4\% of reviews), reviewers used DSTCs to help them focus on a given task amidst all too easily available distractions on their devices (n = 201, 13\%).
The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding, expressed in some reviews as DSCTs helping overcome 'procrastination' (*"seeing my short-term to-do list nudges me to get back to work when I am opening a new website to procrastinate"*, R324, Dayboard \typeBlock\typeTrack \textbf{C}). 
In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \textbf{C}), and/or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack\typeReward \textbf{C}). 
Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11\%; "Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity", R776, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{\emph{Inbox When Ready for Gmail}} \typeBlock\typeGoal\typeTrack \textbf{C}), including studying, writing, or reading (n = 90, 6\%).

Another important use context was \emph{sleep}, where many reviewers said their digital devices often interfered with going to bed at the intended time. However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \textbf{P}).

**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges**
In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely.
This was often expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (n = 58, 3.8\%; *"This was the perfect solution to my phone addiction. It was consuming my life and my free time"*, R612, Forest \typeTrack\typeReward \textbf{P}).
Sometimes reviewers used the clinical labels ADD or ADHD to explain why tools had been helpful to them (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack\typeReward \textbf{P}).

#### **Theme 2: `r theme_two`**

DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 237, 16\%), a theme often expressed in the form of feature suggestions that would make tools more customisable.

One aspect involved getting tools to accurately capture \emph{what} was considered a distraction.
Thus, for block/removal tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (\emph{"I would like to block reddit but still be able to visit productivity related subreddits"}, R1401, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}).
Similarly, self-tracking tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage""*, R989, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}).

Another aspect involved \emph{when} specific activities should count as distractions.
For example, tools for curbing nightly use received feature requests to handle variation in bedtime schedules (*"it needs to allow for the real life fact that people can't go to bed and wake up at the same time everyday. If it was done on your amount of sleep you get in hours that would be better"*, R1250, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \textbf{A}).

**Sub-theme: Alignment between users' and tools' framing of distraction as important for navigating negative emotions elicited by DSCTs**
Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor was the negative emotions that interventions could elicit.
Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and essential, users also needed to carefully navigate the emotional reactions they could create (*"I hate it and love it at the same time. \emojiTearsOfJoy"*, R1299, Stay Focused \typeBlock\typeTrack \textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock\typeTrack \textbf{P}). 
In order for experiences of "shock", "pressure", etc., to be useful, it was important that the user felt that a DSCT's purpose and framing of usage were closely aligned with their personal perspective.

Thus, tools that provided standards for 'good' or 'bad' usage, for example in their presentation of usage statistics or reminder pop-ups, needed to tread carefully.
Being reminded that one did not behave in line with one's intentions could be very helpful (*"This tells it like I need to be told sometimes...LOVE IT! Simple and effective :)"*, R690, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock\typeGoal \textbf{C}). 
However, normative messaging could also backfire and be profoundly annoying if reviewers did not agree with it (*"I was a bit annoyed when Big Brother started getting judgemental on me. “You have hit your target of 50 pickups!” “You’ve now picked up your phone 90 times!” Well, first I didn’t set a target, and second, I don’t think that 54 minutes across a day should make me feel guilty. Especially since I was reading a book during that time"* R911, Moment \typeGoal\typeTrack \textbf{A}).

#### **Theme 3: `r theme_three`**

Reviewers searched for DSCTs that provided interventions whose incentives were 'just right' to bring about an intended change in behaviour without feeling overly intrusive.
The reviews implied a spectrum of individual differences in where this 'Goldilocks level' was. A few reviewers stated that simply receiving reminders or knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, \emph{Mute Screen Time Tracker} \typeGoal\typeTrack \textbf{A}). More commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9\%).

Thus, tools implementing block/removal design patterns were often criticised if they were too easy to disable (*"This is fantastic but the thing is I can simply remove the extension from chrome which defeats the purpose. Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \textbf{C}).
However, if they were *too* difficult to override, they ran into another problem, namely how to deal with emergencies or otherwise unforeseen circumstances.
One developer followed reviewers' suggestion to add a 'pause' button, only to receive critique that blocking was now too easy to override.
Reviewers suggested that adding varying forms of friction to turn off blocking would be a solution, such as having to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock\typeTrack \textbf{P}), or simply adding passwords, which could be set by a trusted other.

Similarly, for tools implementing reward/punishment, some reviewers suggested increasing incentives to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack\typeReward \textbf{C}). Here too, however, tools needed to find the right balance, because punishments that were perceived as too harsh might make users feel like failures (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed'"*, R1247, \href{https://itunes.apple.com/gb/app/sleeptown/id1210251567?mt=8\&uo=4}{SleepTown} \typeTrack\typeReward \textbf{P}).

For tools including goal advancement design patterns, the 'right' level of incentive often related to the frequency or framing of features such as reminder pop-ups, which needed to be finely calibrated to be helpful rather than annoying (*"the todo popup on each page load is quite useful (...) it should be limited (across multiple tabs) to only pop-up every minute or so (...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock\typeGoal \textbf{C}).

**Sub-theme 1: Combining types of design patterns to find the Goldilocks level of incentive**
In addition to tweaking parameters of a tool's current design patterns, many reviewers suggested including other types of patterns to make tools more effective.
Thus, some reviewers for self-tracking tools said that simply being informed about their behaviour was not sufficient for them to change anything.
However, incorporating other types of elements might make it more effective, such as combing self-tracking with goal-setting.

Similarly, some reviewers wanted block/removal tools to not simply block distractions, but also actively nudge their behaviour in ways that supported intended habits.
This might happen by, for example, forwarding blocked websites to productive alternatives, or by replacing algorithmic recommendations with content that they wanted to engage with more ("\emph{if, where the home-screen recommended section is, this could be replaced by a website of our choice} (\ldots) \emph{This way I could do music theory, or learn German. Things I always intend to do, but don't do enough. But if these things came up every time I log into Youtube I'd definitely use them."}, R372, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \textbf{C}).

**Sub-theme 2: DSCTs should have low friction to start and use, but high friction to stop**
Finally, finding the right level of friction also related to the level of effort required to use a tool.
The amount of effort needed to \emph{initiate} use was important to keep low, because a lack of motivation could easily lead users to not use an otherwise effective tool (*"Most users have procrastination issues which make them postpone the moment in which they will click on the "Start focusing" button. Therefore those users (including me) will get distracted and they will click on that button after several hours, or even never, sometimes wasting a whole working day"*, R545, Focus \typeBlock\typeGoal\typeTrack \textbf{C}).
Therefore, a common feature request for blocking tools was the ability to set a schedule where blocking would automatically turn on. An interesting suggestion related to blocking tools alternating between work and break cycles: 
these could flip the ordering of cycles so as to begin with a break, followed by enforced blocking, which might lower the amount of motivation needed to initiate a focus session (*"Hitting the start button the very first time is hard! Starting in a short break would allow me to hit this without a second thought, because I have a few short moments to surf before the focus is forced."*, R232, Block \& Focus \typeBlock\typeTrack \textbf{C}).

The amount of effort involved in routine use of a tool was also important to keep low, because tools that were confusing or demanding were likely to be abandoned.
This seemed to present an especially difficult challenge for self-tracking tools on iOS, because Apple's restrictive developer permissions do not allow apps to easily track time spent in other apps.
As a work-around, some self-tracking tools for iOS required users to take screenshots of their phone's battery page, from which the tool was able to extract information about time spent in different apps.
However, this work-around was simply seen as too effort-demanding by many reviewers (\emph{"I should have looked at how you use this. Take a picture everyday????? Really?"}, R121, App Usage Tracker \typeTrack \textbf{A}).

Thus, whereas high friction related to \emph{turning off} a tool could be actively helpful when it served the purpose of holding the user accountable to their own intentions, tools were criticised if they required too much effort to operate.

## User numbers and ratings

```{r tool-num-users-and-ratings-plot, fig.cap="Distributions of installation numbers and number of ratings on the Google Play, Chrome Web, and Apple App stores (tools with zero ratings scored as 0.5 ratings, to show them in this plot with a log scale). The Apple App store provided no direct information about installation numbers.", out.width='100%'}
to_visualise <- meta_data %>% 
  mutate(installs = case_when(
    store == "play" ~ min_installs,
    store == "chrome" ~ num_users,
    store == "apple" ~ NA_real_
  )) %>% 
  select(tool_id, store, installs, num_ratings) %>% 
  pivot_longer(-c(tool_id, store)) %>% 
  mutate(name = case_when(
    name == "installs" ~ "Min. installs (Google Play) / Users (Chrome Web)",
    name == "num_ratings" ~ "Number of ratings")) %>% 
  mutate(store = factor(store, levels = c("play", "chrome", "apple")),
         name = factor(name))

to_visualise %>% 
  mutate(value = if_else(value == 0, 0.5, value)) %>% # the log transformation removes tools with 0 ratings
  ggplot() +
    aes(x = store, y = value) +
    geom_boxplot(outlier.shape = NA, aes(fill = store), position = position_dodge2(preserve = "single"), alpha = 0.3) +
    scale_y_log10(labels = label_number(big.mark = ",")) +
    geom_jitter(aes(color = store), width = 0.2, height = 0, alpha = 0.5) +
    labs(x = "", y = "") +
    facet_wrap(~name) +
    guides(fill=FALSE, color = FALSE) +
    theme_minimal()

#ggsave("tool-num-users-and-ratings-plot.pdf", width = 7, height = 5)
```

```{r, users-and-category, fig.cap="Top tools in terms of installation numbers (Chrome Web and Play) or number of ratings (Apple App store). Only 13 of the Apple App store tools had received any ratings. Differences between the stores in terms of design patterns are readily apparent: Block/removal patterns are especially common on the Chrome Web store (87% of all tools), where browser extensions can easily modify distracting elements on web pages. Reward/punishment patterns are included in 43% of tools on the Apple App store, compared to, respectively, 6 and 15% on the Chrome Web and Play stores.", out.width="70%"}
knitr::include_graphics("figures/user_numbers_combined_plot.png")
```

```{r user_stats, include=FALSE}
(play_user_stats <- meta_data %>% 
  filter(store == "play") %>% 
  summarise(median_min_installs = median(min_installs),
            min_min_installs = min(min_installs),
            max_min_installs = max(min_installs),
            IQR = IQR(min_installs)))

(chrome_user_stats <- meta_data %>% 
  filter(store == "chrome") %>% 
  summarise(median_installs = median(num_users),
            min_installs = min(num_users),
            max_installs = max(num_users),
            IQR = IQR(num_users)))

(ratings_overall <- meta_data %>% 
  filter(num_ratings >= 1) %>% 
  summarise(median_rating = median(average_rating, na.rm = TRUE),
            min_rating = min(average_rating, na.rm = TRUE),
            max_rating = max(average_rating, na.rm = TRUE),
            IQR = IQR(average_rating, na.rm = TRUE)))

```

The distribution of numbers of users/installations and ratings are shown in Figure \ref{fig:tool-num-users-and-ratings-plot}.
For Google Play tools, the median 'minimum number of installs' was `r format(play_user_stats$median_min_installs, big.mark = ",")` (min = `r format(play_user_stats$min_min_installs, big.mark = ",")`, max = `r format(play_user_stats$max_min_installs, big.mark = ",")`, IQR = `r format(play_user_stats$IQR, big.mark=",")`).
For Chrome Web tools, the median number of users was `r format(chrome_user_stats$median_min_installs, big.mark=",")` (min = `r format(chrome_user_stats$min_min_installs, big.mark=",")`, max = `r format(chrome_user_stats$max_min_installs, big.mark=",")`, IQR = `r format(chrome_user_stats$IQR, big.mark=",")`).
(The Apple App store did not display information about number of users.)
In terms of the ratings given to tools, the median average rating was `r ratings_overall$median_rating` (min = `r ratings_overall$min_rating`, max = `r ratings_overall$max_rating`, IQR = `r round(ratings_overall$IQR, 1)`; ratings given on a scale from 1 to 5 stars on all stores).

Figure \ref{fig:users-and-category} shows the top tools, ranked by number of users (Chrome Web and Play) or number of ratings (Apple App store), alongside the design patterns they include.
The distributions of design patterns included in top tools differed by store, mirroring differences between the stores overall: for example, block/removal functionality was rare on the Apple App store compared to the Google Play and Chrome Web store [@lyngsSelfControlCyberspaceApplying2019].

```{r, include=FALSE}
chrome_and_play_rankings <- meta_data %>% 
  filter(num_ratings >= 30) %>% 
  group_by(store) %>% 
  mutate(installs = case_when(
    store == "chrome" ~ num_users,
    store == "play" ~ min_installs
  )) %>% 
  select(store, tool_id, average_rating, installs) %>% 
  mutate(rank_rating = rank(desc(average_rating)),
         rank_installs = rank(desc(installs)))

cor.test(chrome_and_play_rankings %>% filter(store == "chrome") %>% pull(rank_rating),
      chrome_and_play_rankings %>% filter(store == "chrome") %>% pull(rank_installs), 
      method = "kendall")

cor.test(chrome_and_play_rankings %>% filter(store == "play") %>% pull(rank_rating),
      chrome_and_play_rankings %>% filter(store == "play") %>% pull(rank_installs), 
      method = "spearman")

meta_data %>% 
  mutate( has_blocking = str_detect(categories_alphabetically, "Block/removal")) %>% 
  group_by(store, has_blocking) %>% 
  summarise(number = n()) %>% 
  mutate(prop = number / sum(number)) %>% 
  filter(has_blocking == TRUE)

meta_data %>% 
  mutate( has_goal = str_detect(categories_alphabetically, "Goal-advancement")) %>% 
  group_by(store, has_goal) %>% 
  summarise(number = n()) %>% 
  mutate(prop = number / sum(number)) %>% 
  filter(has_goal == TRUE)

meta_data %>% 
  mutate( has_track = str_detect(categories_alphabetically, "Self-tracking")) %>% 
  group_by(store, has_track) %>% 
  summarise(number = n()) %>% 
  mutate(prop = number / sum(number)) %>% 
  filter(has_track == TRUE)

meta_data %>% 
  mutate( has_reward = str_detect(categories_alphabetically, "Reward/punish")) %>% 
  group_by(store, has_reward) %>% 
  summarise(number = n()) %>% 
  mutate(prop = number / sum(number)) %>% 
  filter(has_reward == TRUE)
```

```{r number-of-design-patterns, include=FALSE}
meta_data %>% 
  count(num_categories) %>% 
  mutate(prop = n/sum(n))

data_test_one_vs_two <- meta_data %>% 
    filter(num_ratings >= 15) %>% 
    mutate(num_categories = as.numeric(num_categories),
           num_categories_one_or_more = ifelse(num_categories == 1, "1", "2-4"))

wilcox.test(average_rating ~ num_categories_one_or_more, data = data_test_one_vs_two)

(median_one <- data_test_one_vs_two %>% filter(num_categories_one_or_more == "1") %>% pull(average_rating) %>% median(na.rm = TRUE))
(median_two_to_four <- data_test_one_vs_two %>% filter(num_categories_one_or_more == "2-4") %>% pull(average_rating) %>% median(na.rm = TRUE))

```

Across all tools, 64% included a single type of design pattern for digital self-control, whereas 32% combined two - very few combined three (3%) or all four (1%).
Tools that combined two or more types had significantly higher average ratings than tools which implemented a single type (median average rating with a single type = 4.2, median for tools combining two or more = 4.39, *p* = 0.003 in Wilcoxon's rank-sum test; analysis excludes tools with fewer than 15 ratings; see Figure \ref{fig:average-ratings-by-functionality}).

On the Chrome Web and Play stores, Block/removal was the most common type of design pattern (included in 87% and 65% of all tools, respectively, compared to 37% on the Apple App store), whereas Reward/punishment was the least common (included in 6% and 15% of tools, respectively, compared to 43% on the Apple App store).
Notably, the app 'Forest' (which combines reward/punishment and self-tracking in the form of virtual trees that grow while a timer counts down) ranked highest on the Play store and 5th on the Chrome Web store, while a very similar app, 'Flora', ranked highest on the Apple App store in terms of number of ratings received.

(ref:caption-ratings-by-functionality) Average ratings by design patterns in digital self-control tools (for tools with 15 or more ratings). Tools including more than one type of design pattern had higher average ratings (median = 4.39) than tools implementing a single type (median = 4.2, *p* = 0.003, Wilcoxon's rank-sum test).

```{r, average-ratings-by-functionality, fig.cap="(ref:caption-ratings-by-functionality)", out.width="100%"}
knitr::include_graphics("figures/ratings_by_type_annotated.pdf")
```

# Discussion

To sum up, the user reviews described DSCTs as crucial support for focusing on less instantly rewarding tasks amidst easily available digital distractions. Specific usage contexts related mainly to productivity/work (including studying, reading, and writing), as well as sleeping.
Individuals with greater self-regulation difficulties seemed to find DSCTs particularly useful, as evidenced by reviewers who explained usefulness in terms of coping with 'addiction' or conditions like ADHD.

Reviewers wanted DSCTs to provide interventions that accurately targeted usage they wanted to better control, often expressed as feature requests for, for example, being able to target blocking to specific functionality on a website, or exempt specific apps from usage tracking.
Close alignment between user and tool in their definitions of 'distracting use' was important, partly to reduce negative emotional reactions to specific interventions, which might otherwise be considered too annoying.

Reviewers also wanted DSCTs to provide a level of incentive that was sufficient for affecting intended behaviour change without feeling too intrusive.
Reviewers differed in the level of incentive they wanted, but were commonly looking for interventions strict or rewarding enough to stay on track when they felt tempted to indulge in digital distractions or drift into unwanted habitual use.
DSCTs needed to find a delicate balance, as overly strong incentives might backfire if they made users feel dejected when they failed to live up to their intentions, or if users had no ways to handle emergency use or shifting daily schedules.

Some reviewers suggested that different types of design patterns should be combined to provide the right level of incentive. This matched our observation in the popularity metrics that DSCTs which combined two or more types of design patterns received higher overall ratings than tools including only one.
A final consideration related to how much effort a tool required to use: Whereas a high level of friction to *stop* a tool could help hold users accountable to their goals, the level of friction required to *start* use should be low, to handle initial lack of motivation on the part of the user.

In the following, we discuss implications for future research.

\subsection{Using self-regulation measures to evaluate DSCTs and assess individual user needs}
Our work provides additional data points for the discussion around how to evaluate effectiveness of DSCTs, and understand how needs for support vary between users.
In terms of behavioural measures, our thematic analysis and the term frequencies in user reviews overall suggested that limiting time spent on devices, or in specific apps and services, was an important concern for many.
Hence, whereas simple measurements of 'screen time' may have many inadequacies in assessing DSCT effectiveness [@cecchinatoDesigningDigitalWellbeing2019; @hinikerMovingGreatScreen2019; @lukoffDigitalWellbeingWay2019], they remain important.

More commonly than limiting time, however (observed three times as frequently in our thematic analysis), the reason for using DSCTs was specifically to support focus on less instantly rewarding tasks in the face of easily available digital distractions.
This matches findings from research using other methods [@reineckeSlackingWindingExperience2016], such as Tran et al. [@tranModelingEngagementDisengagementCycle2019]'s interview study showing that tedious or effortful activities are common triggers of compulsive smartphone use, and Kim et al.'s finding that their phone-restriction app was found particularly useful in "boring classes" [@kimLetFOCUSMitigating2017].
What are the most useful ways for us to assess DSCTs' effectiveness for this purpose?

In a given context, simple behavioural measures may provide useful proxies, such as error rates or output in a work task [@borghoutsLookingInformationEmail2018; @marottaOnlineDistractionsWebsite2017] or time spent in functionality 'relevant' or 'distracting' relative to a goal [@lottridgeBrowserDesignImpacts2012; @hinikerMyTimeDesigningEvaluating2016; @whittakerDonWasteMy2016; @kimTechnologySupportedBehavior2017].
However, we also need effective ways to assess users' own experience.

Some existing work, especially on 'problematic use', has used self-report scales related to behavioural addiction to evaluate DSCTs and/or assess baseline difficulties with managing device use [@leeSAMSSmartphoneAddiction2014; @koLockLoLGroupbased2016].
Many researchers have pushed back against this approach, because it risks pathologising everyday patterns of use.
Thus, it has been argued that even frequent, extensive use that distracts users from daily life should not be considered "addiction" unless it also leads to functional impairment and psychological distress [@billieuxDysfunctionalUseMobile2015; @kardefelt-wintherHowCanWe2017; @tranModelingEngagementDisengagementCycle2019].
Moreover, references to 'addiction' among ordinary users often reflect a loose use of the term common in media and society [@tikuWIREDGuideInternet2018; @lanetteSmartphoneAddictionNarrative2018; @orbenTeensScreensWellBeing2019].
In the user review data, we did observe a number of mentions of 'addiction' (n = 52, 3%), but overall the data seemed to reflect a situation in which most users' challenges with focusing and avoiding wasted time reflected more mundane struggles and annoyances than what would deserve the clinical term 'addiction'.

A range of alternatives to addiction measures for assessing users' experience with DSCTs have been explored, including measurements of self-reported stress (e.g., the Perceived Stress scale, \cite{collinsBarriersEngagementPersonal2014}, from \cite{Cohen1983}), workload (e.g., NASA-TLX, \cite{kimLocknTypeLockoutTask2019}, from \cite{Hart1988}), and absorption in a task (e.g., Cognitive Absorption scale, \cite{markEffectsIndividualDifferences2018} from \cite{agarwalTimeFliesWhen2000}).
We suggest, however, that the most directly relevant literature to draw on is existing psychological work on self-regulation and self-control.
Here, a substantial body of research has investigated how individual differences and environmental influences affect the ability to regulate behaviour towards desired goals in the face of conflicting impulses and distractions.
HCI research on DSCTs can be viewed as a sub-field in this respect, which investigates self-regulation in relation to digital devices such as smartphones and laptops, and how to support it by design.

Thus, early research on 'internet addiction', and later 'smartphone addiction', did derive great value from adapting behavioural addiction research to investigate severe cases of self-regulation breakdown among a minority of users [@chakrabortyInternetAddictionConsensus2010].
Future HCI research on design remedies for more mundane self-regulation challenges experienced by the great majority of users could benefit from a similar effort to adapt common measurement scales from basic psychological research on self-regulation to the context of digital device use.
As an example, exploratory work by Mark et al. \cite{markEffectsIndividualDifferences2018} found that blocking tools preferentially benefits individuals low in trait self-control.
The measure of self-control used was sub-items from the Cognitive Absorption scale, a scale developed for measuring a 'state of deep involvement with software'.
Follow-up confirmatory work could adapt e.g., the widely used UPPS Impulsivity Behaviour scale [@Whiteside2001] to investigate the robustness of Mark et al.'s finding and how it relates to well-established personality factors underpinning self-regulation ability.

Similarly, the Brief Self-Control Scale [@tangneyHighSelfcontrolPredicts2004] has been widely used by psychologists as a uni-dimensional measure of ability to self-regulate behaviour in the face of undesired impulses, and found to predict positive life outcomes in various domains \cite{Ferrari2009}.
This scale has also been used as a measure assessing change in self-control ability over time \cite{Duckworth2010}, and might fruitfully be adapted as an effectiveness metric for interventions in DSCTs.

Adapting such measures to the domain of digital device use could provide metrics that more directly measure impact on the ability to self-regulate behaviour in the face of digital distractions than the alternatives commonly used by HCI researchers. 
Therefore, they might have a closer construct fit with what users commonly want DSCTs to achieve, as evidenced by the user review data and existing research, and also more readily allow for knowledge transfer with fundamental psychology research relevant to DSCTs.

## Using psychological theory to explore interactions between design patterns and framings of 'distraction'
Existing work has sometimes implied a single spectrum of 'friction' on which the intervention intensity of a given DSCT may be increased or decreased to find an optimal level \cite{kimGoalKeeperExploringInteraction2019, kimLocknTypeLockoutTask2019, kovacsConservationProcrastinationProductivity2019}.
However, our observations that reviewers often wanted different types of design patterns to be combined, and that a poor fit between the user's and tool's framing of 'good' and 'bad' device use might make people rebel against a tool, suggest that designers need to consider how several psychological dimensions interact.

One simple framework for this purpose is the dual-systems model presented by Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019], which summarises main mechanisms of self-regulation as viewed by current cognitive psychology.
According to this model, the effects of DSCTs on self-regulation can be considered in terms of three factors: 
(i) automatic habits and reflexes (System 1 control), 
(ii) conscious goals and self-monitoring (System 2 control), and 
(iii) the expected value of control (mediator of System 2 control).

At any given point, these factors activate potential behaviours and thoughts, and the action that receives the highest activation controls our behaviour.
For example, if you take out your smartphone to check Instagram, it can happen via automatic System 1 control in response to a boring situation, because you have a habit of doing so in the presence of certain cues.
It might also happen via conscious System 2 control if you remember that you need to message a friend, and so this conscious goal triggers behaviour [@lyngsSelfControlCyberspaceApplying2019].

Self-control challenges occur when actions triggered by System 1 and System 2 are in conflict. 
For example, to meet a goal of not checking Instagram at family dinners, you might need to suppress an impulse to check. 
If you fail, it can be because you had forgotten this goal, or because the 'expected value of control', which mediates the strength of System 2 control, was too low:
The more value a conscious goal has (whether because its rewards are greater or more immediate, or because you are more confident in your ability to achieve it), the more powerfully it triggers action.
In other words, you might fail to suppress your impulse to stay off your phone, because your valuation of staying off was not high enough to compete with the urge to check.

We find this simple partitioning of the main psychological dimensions of self-regulation helpful for considering the purpose and design space of DSCTs.
Thus, from this perspective, the purpose of DSCTs is to bias the internal competition between potential actions such that the ones that win out are in line with the user's enduringly valued conscious goals [@duckworthStitchTimeStrategic2016].
DSCTs can achieve this via interventions that target a limited number of factors --- blocking tools can help prevent unwanted System 1 responses from being triggered or acted on, goal advancement and self-tracking tools can enable System 2 control by activating certain goals in working memory, and reward/punishment tools can make System 2 control more likely to succeed by increasing the expected value of control [@lyngsSelfControlCyberspaceApplying2019].

We can use this model to consider users' suggestions of combining distinct types of design patterns:
it is possible that by combining design patterns which target distinct psychological mechanisms, each can be implemented at a more gentle level of intensity while still in aggregate provide enough input to the internal competition to bring about an intended behaviour change.
By contrast, tools that focus on a single design pattern targeting a single psychological mechanism may need to ramp up its intensity more highly before sufficient impetus for behaviour change is provided --- and in so doing be at greater risk of being perceived as overly annoying.
For example, \emph{Forest} combines growing of virtual trees with a countdown timer, which influences both the expected reward for control and our sensitivity to delays [@lyngsSelfControlCyberspaceApplying2019].
Based on this tool's popularity, with over 10 million users on Android alone, this approach seems highly effective for many.
An alternative version which simply grew virtual trees without a countdown timer might need to correspondingly increase the magnitude of reward to remain useful, but in so doing be more likely to be perceived by users as too forceful or manipulative [@Lee2011Mining].
^[We do note that the higher ratings given to tools combining two or more distinct design patterns has an alternative explanation: tools that combine design patterns may simply have received more developer attention, and so higher ratings could simply be due to such tools being better designed.]

Similarly, we may reach to the psychological literature for design guidance related to the observation that alignment between the user and tool in the framing of distraction is important for the user in managing potential negative emotions.
In Lyngs et al.'s dual-systems model, this might be considered a mediator of the expected value of control, but the model is too high level to provide useful advice in this respect.
Instead, we propose that future research draw on Self-Determination Theory (SDT), which has been frequently used and well-validated in HCI research but rarely applied in research on DSCTs \cite{ballouSelfDeterminationTheoryInHCI2022, ryan2017}.
SDT research on the factors that determine whether people identify and integrate with a source of extrinsic motivation, or whether they experience it as a less authentic external influence, could prove highly useful for understanding how to frame DSCTs' interventions so that they are more likely to be accepted by the user as a desirable extension of their personal will.

We encourage future research to explicitly investigate interactions between interventions types and framings of 'distraction' using these psychological theories as guide.

## Navigating platform challenges for future DSCTs
Looking towards the future, we note that the underlying technical infrastructures and platforms create very different conditions for the emergence of ecosystems of DSCT in the wild.
The largest degree of flexibility exists on the web, where browser extensions make it possible for developers --- and tech-savvy end-users --- to easily customise the look and functionality of any website via injected stylesheets and scripts.
Ecosystems of browser extensions today exist for all major browsers, which all share the cross-browser \href{https://developer.mozilla.org/en-US/docs/Mozilla/Add-ons/WebExtensions/API}{Web Extensions API}.

The situation is very different if a user wishes to experiment with DSCTs on a smartphone or tablet.
Unless users jailbreak their devices, no options exist for users to customise the look and function of mobile apps beyond the options that app developers chooses to provide.
Moreover, the options available for self-tracking or blocking tools are dramatically different between Android and iOS:
Because of Google's less restrictive approach to developer permissions, a diverse ecosystem exists on Android for apps that provide bespoke usage visualisation or customised blocking schedules, in addition to what Google provides themselves in their 'Digital Wellbeing' app.
On iOS, by contrast, Apple's Screen Time tool was, at the time data for the present paper was scraped, the only option with which users could block functionality and see their usage of different apps^[A Screen Time API, introduced in the summer of 2021 after long campaigning by developers [@digitalwellnesswarriorsAppleLetDevelopers2018], now allows creation of blocking tools for parental control purposes, but does not enable apps for tracking and visualising screen time.].
As we observed in the user reviews, workarounds by iOS developers, such as asking users to routinely take screenshots of the battery page on their phone settings, seemed to be too effort-demanding on the part of users to be workable.

We make no judgment of whether Apple or Google are 'right' in terms of the level of permissions to make available to developers, as greater operating system access also entails greater security risks.
We do note, however, that such risks are already present in browser extensions, where both Apple and Google, and other companies, already allow extensions that in many respects have greater powers than Android apps.
Here we simply wish to highlight the distinct roads ahead, in relation to providing the interventions that users seek to stay in control:

Google's less restrictive approach has meant that a diverse range of potent digital self-control tools are available on Android.
Therefore, Google may face less pressure to 'get it right' with their own Digital Wellbeing tools, because less common user needs can be addressed by tools within an ecosystem of DSCTs on their platform.
Apple's approach, however, implies that their own Screen Time app needs to accommodate a very wide range of user needs.
This will be challenging, as there can be a trade-off between customisability and effort of use, where providing sufficiently fine-grained controls to account for all users' needs can make it too complicated to navigate.

## Limitations and future work
First, the data used in the present paper was scraped in March 2019. Hence, the findings presented may deviate from what would by learned by re-scraping the app and browser extension stores in 2022.
However, all of our data and materials are openly available, and so our process should be readily repeatable.
As such, we believe the present paper provides a useful snapshot of DSCTs across three online stores, which can serve as point of comparison for future research tracking how the landscape of these tools have developed.

Second, the content of user reviews is subject to various forms of bias. 
For example, reviews may be more likely to represent views from users whose experiences with a particular tool are either highly positive or negative. 
That is, for a user to be sufficiently motivated to bother reviewing a tool in the first place, their level of excitement about a tool might exceed the excitement of the average user [@panichellaHowCanImprove2015].
Neverthelesss, the fact that our findings from reflexive thematic analysis resonated well with findings from studies using other methods such as surveys or interviews [@koNUGUGroupbasedIntervention2015; @tranModelingEngagementDisengagementCycle2019] suggests that user reviews of DSCTs can provide valid and useful insights.

Finally, for the present paper, we programmatically sampled reviews for thematic analysis from our full data set of reviews with the goal of sampling broadly across stores and types of design patterns.
However, reviews can be sampled in many different ways to provide data tailored to more narrow research questions, e.g. for researchers specifically investigating design parameters relevant for lockout mechanisms [@kimGoalKeeperExploringInteraction2019].
We encourage future research on DSCTs to make their data sets and analysis procedures open for others to re-use, to encourage synergies between the efforts of different research groups in the field.

# Conclusion
The landscape of digital self-control tools available online amounts to hundreds of thousands of natural experiments in supporting self-regulation of device use.
As such, it provides a running commentary on user needs and struggles in common digital environments, as well as concrete suggestions for solutions.
Therefore, studying this landscape can be a valuable complementary approach to evaluating interventions in controlled studies, as the latter is difficult to scale to broad assessment of strategies and implementations across the design space.

In this paper, we investigated what user reviews reveal about digital self-control tools' contexts of use and design challenges, and how design features relate to ratings and installation numbers, using data from 334 tools on the Google Play, Chrome Web, and Apple App stores.
Compared to existing research, our work followed a more transparent procedure for sampling user reviews for thematic analysis from the full data set, with reproducible, programmatic sampling and open data and materials.

As most daily tasks come to be mediated by digital devices, it is important that users are given the tools they need to effectively manage time and attention in their digital environments.
As such, discovering effective interventions for support self-regulated digital device use is a practical and urgent research challenge.
In this respect, we encourage future work to make further use of the landscape of digital self-control tools on app and browser extension stores as a resource for investigating effective design solutions.

# Acknowledgments
Funding: Ulrik Lyngs was funded by the UK Engineering and Physical Sciences Research Council (EPSRC) [grant number EP/J017728/2]; and the Carlsberg Foundation [grant number CF20-0678].
The other authors received no specific grant for this research.
We thank Claudine Tinsman for proof-reading a late version of the manuscript.

### 2022-03-08 materials/scrapers/apple-app-store/github-scraper/2019-03-19-scrape_reviews.js

@@ -1,0 +1,94 @@
+const store = require('app-store-scraper');
+
+const fs = require('fs')
+
+// couldn't scrape for
+
+const app_list = ['cpimhoff.Tabula.iOS', 'com.masaoyamashita.myfirstapp', 'com.emberify.appusagetracker', 'com.richhenderson.banish', 'com.stogaheimllc.BeePresent', 'com.blacklistapp.Blacklist', 'pabloweb.net.SelfControl', 'com.caramba.phoneuse', 'com.anti-apps.bSociable', 'com.imobiapp.screentimer', 'com.bateman.charityJar', 'nl.schungel.checkout', 'io.magpiedigital.Clar', 'com.getcluster.Compose', 'com.naruthk.DailyNotes', 'org.DinnerMode.ios', 'club.donutdog.ios', 'com.questmobile.draco', 'com.getfeedless.feedless', 'com.riko.suyasaso', 'com.appfinca.flora.ios', 'com.Rave.Focus', 'noxa.Focus', 'com.NextGen.Focus-Drive', 'Guleed-Abdi.FocusApp', 'com.talkingcucumberltd.FocussuPro', 'com.ezoneapps.ForestGrowless', 'com.sen.lim', 'com.jeremy.Goyfaw', 'com.bellostudios.hooked', 'com.bellostudios.hookedpro', 'com.InternetAddiction.Calendar', 'VTT.Just-Write', 'com.jordan-carney.Liberate', 'org.mobpage.markdowneditor', 'com.kevinholesh.Moment', 'to.glued.Mute', 'cn.bahamut.notouchphonealarm', 'com.alexlongerbeam.NoPhone-Timer', 'org.aliyilmaz.notelr', 'com.One-Studio.NoteFlow', 'ru.phoneless.Phoneless', 'com.7CD2H2J37A.moment', 'ca.genoe.Refrain', 'com.imobiapp.screenfree', 'vn.wehelp.SelfControlLite', 'nl.sidekickmattie.sidekickmattie', 'com.yeowjin.simpletext', 'com.lukejmann.SiteBlocker', 'com.seekrtech.sleep', 'com.usedopamine.app.space', 'com.ron.StayFocused', 'com.Milk.milkthemoment', 'com.erichuju.pomodoro', 'io.timeguard.app', 'com.bluecocoa.time2', 'com.globussoft.unhookme', 'com.emberify.unplug2']
+
[…]

### 2022-03-08 materials/scrapers/apple-app-store/github-scraper/test.js

@@ -1,0 +1,9 @@
+var store = require('app-store-scraper');
+
+store.reviews({
+  appId: 'com.stogaheimllc.BeePresent',
+  sort: store.sort.HELPFUL,
+  page: 2
+})
+.then(console.log)
+.catch(console.log);

### 2022-03-08 materials/scrapers/play-store/2019-03-19-scrape-play-metadata.js

@@ -1,0 +1,43 @@
+const gplay = require('google-play-scraper');
+
+const fs = require('fs')
+
+const app_list = ['co.offtime.kit', 'olsencheung.lockyourphoneproject', 'com.tobiasschuerg.smartphonestats', 'com.antiaddication.innomotion', 'com.goozix.antisocial_personal', 'com.appblockernew', 'com.timelesssky.appgatekeeper', 'com.aozora_create.appofftimer', 'com.a0soft.gphone.uninstaller', 'cz.mobilesoft.appblock', 'de.dfki.appdetox', 'bis.mccaffrey.myfinalyearproject', 'com.incubasys.apptym', 'me.noip.AppUseFree', 'com.adventurize.artiquill', 'com.runnably.bashful', 'notstupid.directsearch', 'com.blackout', 'com.wverlaek.block', 'co.blocksite', 'com.calm.checky', 'com.felixlogic.coldturkey', 'forinnovation.phoneaddiction', 'com.urbandroid.ddc', 'com.Appdesk.MuteIt', 'bg.bozho.dumbphone', 'com.Dyve.Dyve', 'com.kaamchor.android.fameorshame', 'kr.co.rinasoft.yktime', 'com.flipd.app', 'com.lazygeniouz.focus', 'com.NoxfallStudios.StudyCatApp', 'com.curlybrace.ruchir.keepfocusneverprocrastinateagain', 'com.fokison.mobile', 'com.manifest.fomo', 'cc.forestapp', 'com.apps.adrcotfas.goodtime', 'ee.totalstar.gooffline', 'com.andre3.wzlimitdistraction', 'com.apps.dsimpletools.helpmefocus', 'de.cherrypit.idletime', 'com.ifocusmode.phone.addiction', 'com.deGans.itsStudyTime', 'com.jotterpad.x', 'vikesh.dass.lockmeout', 'com.teqtic.lockmeout', 'com.andr.slema.delayedapplock', 'dodolabs.lokeet', 'nl.sidekickmattie.mattie', 'meetingtime.app', 'in.blogspot.alcory.mobileaddictionmeterpro', 'com.it.appstracker', 'com.monitro.android', 'com.assembtec.nodistractions', 'me.aswinmohan.nophone', 'com.offthegrid', 'jp.s122107.phonescreentimer', 'phonesmart.lock.self.control', 'com.ryosoftware.phoneusagemonitor', 'me.noip.PhoneUseFree', 'com.pock
[…]

### 2022-03-08 materials/scrapers/play-store/2019-03-19-scrape-play-reviews.js

@@ -1,0 +1,93 @@
+const gplay = require('google-play-scraper');
+
+const fs = require('fs')
+
+const app_list = ['co.offtime.kit', 'olsencheung.lockyourphoneproject', 'com.tobiasschuerg.smartphonestats', 'com.antiaddication.innomotion', 'com.goozix.antisocial_personal', 'com.appblockernew', 'com.timelesssky.appgatekeeper', 'com.aozora_create.appofftimer', 'com.a0soft.gphone.uninstaller', 'cz.mobilesoft.appblock', 'de.dfki.appdetox', 'bis.mccaffrey.myfinalyearproject', 'com.incubasys.apptym', 'me.noip.AppUseFree', 'com.adventurize.artiquill', 'com.runnably.bashful', 'notstupid.directsearch', 'com.blackout', 'com.wverlaek.block', 'co.blocksite', 'com.calm.checky', 'com.felixlogic.coldturkey', 'forinnovation.phoneaddiction', 'com.urbandroid.ddc', 'com.Appdesk.MuteIt', 'bg.bozho.dumbphone', 'com.Dyve.Dyve', 'com.kaamchor.android.fameorshame', 'kr.co.rinasoft.yktime', 'com.flipd.app', 'com.lazygeniouz.focus', 'com.NoxfallStudios.StudyCatApp', 'com.curlybrace.ruchir.keepfocusneverprocrastinateagain', 'com.fokison.mobile', 'com.manifest.fomo', 'cc.forestapp', 'com.apps.adrcotfas.goodtime', 'ee.totalstar.gooffline', 'com.andre3.wzlimitdistraction', 'com.apps.dsimpletools.helpmefocus', 'de.cherrypit.idletime', 'com.ifocusmode.phone.addiction', 'com.deGans.itsStudyTime', 'com.jotterpad.x', 'vikesh.dass.lockmeout', 'com.teqtic.lockmeout', 'com.andr.slema.delayedapplock', 'dodolabs.lokeet', 'nl.sidekickmattie.mattie', 'meetingtime.app', 'in.blogspot.alcory.mobileaddictionmeterpro', 'com.it.appstracker', 'com.monitro.android', 'com.assembtec.nodistractions', 'me.aswinmohan.nophone', 'com.offthegrid', 'jp.s122107.phonescreentimer', 'phonesmart.lock.self.control', 'com.ryosoftware.phoneusagemonitor', 'me.noip.PhoneUseFree', 'com.pocketpoints
[…]

### 2022-03-08 01-processing-raw-data_reviews.Rmd
fix typo
get rid of excess whitespace in review text

@@
 ## Extract reviews from page sources
-In this case, we've store .RDS files that contain the page sources.
+In this case, we've stored .RDS files that contain the page sources.
 
@@
   rename(rating = numstars) %>% 
+  mutate(text = str_squish(text)) %>% # get rid of excess whitespace 
   distinct(across(-review_page)) %>% 
@@
 play_reviews_tibble %>% 
-  mutate(date = lubridate::mdy(date)) %>% 
+  mutate(date = lubridate::mdy(date),
+         text = str_squish(text)) %>% 
   distinct() %>% 
@@
   select(-reviews) %>% 
-  unnest()
+  unnest(cols = c(review_info))
 
@@
   select(-review_title) %>% 
-  # turn the rating into an integer
-  mutate(rating = str_replace(rating, "out of 5", "") %>% as.integer()) %>% 
-  # change date format
-  mutate(date = lubridate::dmy(date)) %>% 
+  # turn the rating into an integer and change date format
+  mutate(rating = str_replace(rating, "out of 5", "") %>% as.integer(),
+         date = lubridate::dmy(date)) %>% 
+  # get rid of excess whitespace
+  mutate(review_text = str_squish(review_text)) %>% 
   distinct() %>% 

### 2022-03-08 04-sample_reviews_for_ta.Rmd
add script for sampling for thematic analysis

---
title: "01-review-wrangling"
output: html_document
---

```{r}
library(tidyverse)
library(writexl)
```

# Extracting reviews for analysis
## Our exclusion process
- we exclude tools that have fewer than 5 ratings (apple) or less than 100 installs / users (google play and chrome web)
- we exclude reviews with fewer than 10 words
- we want: broad distribution across functionality and stores
- we do, in order of the most popular: for a main functionality type, we slice off a tool from each store which has that type of functionality and sample up to 30 reviews from it; then we move on to the next functionality type; we repeat until we have about 1500 reviews

## Read in the review data
Get the reviews that have more than 10 words

```{r}
review_data <- read_csv(here::here("data", "processed", "distinct_reviews.csv")) %>% 
  mutate(review_word_count = stringi::stri_count_words(review_text)) %>% 
  filter(review_word_count >= 10) %>% 
  select(-review_word_count)

# get also the review counts
review_counts <- review_data %>% 
  group_by(tool_id) %>% 
  summarise(num_reviews = n())

```

## Get the tool info
We apply the exclusion criteria and order them by our ranking measure

```{r}
store_info <- read_csv(here::here("data", "processed", "meta_data_all_w_design_pattern_types.csv"))

# exclude if they have fewer than 5 ratings (apple) or less than 100 installs / users on the google play and chrome web stores
# arrange by numbers of users (num of ratings for the apple store)
tool_by_num_users <- store_info %>% 
  select(store, tool_id, title, average_rating, num_ratings, min_installs, num_users, categories_alphabetically) %>% 
  left_join(review_counts) %>% 
  filter(!is.na(num_reviews)) %>% 
  filter((store == "apple" & num_ratings >= 5) | min_installs >= 100 | num_users >= 100) %>% 
  mutate(ranking_measure = case_when(
    store == "play" ~ min_installs,
    store == "chrome" ~ num_users,
    store == "apple" ~ num_ratings
  )) %>% 
  arrange(store, desc(ranking_measure))

```

Ok, so let's write a nested for loop to grab the reviews.

```{r}
feature_categories <- c("Block/removal", "Goal-advancement", "Self-tracking", "Reward/punish")
stores <- c("apple", "chrome", "play")

# initialise empty results and apps sampled from
review_results <- tibble(
)
apps_sampled_from <- c()

while(nrow(review_results) < 1500){
  for (category in feature_categories){
    current_category_tools <- tool_by_num_users %>% 
      filter(!tool_id %in% apps_sampled_from) %>% 
      filter(str_detect(categories_alphabetically, category))

    for (cur_store in stores) {
      # print out how many we have and break the loop if we're about 1500
      print(str_c("num reviews sampled: ", nrow(review_results)))

      if(nrow(review_results) > 1500) {
        print("breaking loop")
        break
      }

      #print(cur_store)
      current_store_tools <- current_category_tools %>% 
        filter(store == cur_store)

      current_app <- current_store_tools %>% 
        slice(1) %>% 
        pull(tool_id)

      if (length(current_app) > 0){
        current_reviews <- review_data %>% 
          filter(tool_id == current_app) %>% 
          slice(1:30) %>% 
          slice_sample(n = 30, replace = FALSE)

        # update reviews and apps sampled from
        review_results <- bind_rows(review_results, current_reviews)
        apps_sampled_from <- c(apps_sampled_from, current_app) 
      }
    }
  }
}

```

We'll add a date to the reviews

```{r}
all_reviews_w_most_recent_date <- read_csv(here::here("data", "processed", "all_reviews.csv")) %>% 
  group_by(row_number()) %>% 
  top_n(1, date)

review_results %>% 
  mutate(review_id = str_c("R", row_number())) %>% 
  left_join(all_reviews_w_most_recent_date) %>% 
  left_join(store_info) %>% 
  select(review_id, store, title, tool_id, categories_alphabetically, user_name, date, rating, review_text) %>% 
  arrange(title, tool_id, date) %>% 
  write_xlsx(str_c(here::here("data", "processed/"), lubridate::today(), "-reviews_for_ta.xlsx"))

```

### 2022-03-08 02-wrangling_reviews.Rmd
save out all reviews too

@@
   filter(type == "review") %>% 
-  select(tool_id = extension_url, user_name, date = date_modified, rating, review_text = text, store) %>% 
+  # use the unique ending of the URL as tool id
+  mutate(tool_id = str_remove(extension_url, "\\?hl=.+") %>%  str_extract("/[^/]+$") %>% str_remove("/")) %>% 
+  select(tool_id, user_name, date = date_modified, rating, review_text = text, store) %>% 
   bind_rows(play_reviews %>% rename(review_text = text, tool_id = app_id)) %>% 
   bind_rows(apple_reviews %>% select(tool_id = app_id, user_name, date, rating, review_text, store))
+
+all_reviews %>% 
+  write_csv(here::here("data", "processed", "all_reviews.csv"))
 
@@
   # simplify toolid for chrome web extensions
-  mutate(tool_id = if_else(store == "chrome", 
-                           str_remove(tool_id, "\\?hl=.+") %>%  str_extract("/[^/]+$") %>% str_remove("/"),
-                           tool_id)) %>% 
   relocate(tool_id, store) %>% 

### 2022-03-08 02-wrangling_reviews.Rmd
rm obsolete comment

@@
 all_reviews_distinct %>%
-  # simplify toolid for chrome web extensions
   relocate(tool_id, store) %>% 

### 2022-04-29 paper.Rmd
use visual editor

@@
 
-In academia, a growing amount of HCI research is focusing on developing and evaluating various design patterns meant to help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019]^[We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders), and the term ‘digital self-control tool’ (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook’s newsfeed).
-We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user’s usual digital environment, or when it is evaluated in a research study.].
+In academia, a growing amount of HCI research is focusing on developing and evaluating various design patterns meant to help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019][^1].
 This area of HCI work has gradually been building an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
 However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how much people vary in how useful they find particular interventions, and how specific design patterns' usefulness depends on details of their implementation, such as the level of 'friction' they introduce [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].
+
+[^1]: We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders), and the term 'digital self-control tool' (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook's newsfeed).
+    We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user's usual digital environment, or when it is evaluated in a research study.
 
@@
 
-Users want DSCTs to flexibly adapt to their personal definition of ‘distraction’, as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
+Users want DSCTs to flexibly adapt to their personal definition of 'distraction', as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
 A key reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions.
@@
 Finally, users want DSCTs to provide carefully calibrated incentives that are sufficient to incentivise intended behaviour change without feeling overly intrusive.
-In our analysis, this 'Goldilocks level'^[After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'.
+In our analysis, this 'Goldilocks level'[^2] varied between reviewers, but they commonly wanted tools to provide incentives powerful enough to keep themselves on track in moments of temptation or absentmindedness.
+Overly strong incentives could backfire if they simply made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).
+
+[^2]: After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'.
     The concept of the 'just right amount' is often referred to as the 'Goldilocks principle'.
-For example, a widely cited paper on the relationship between screen time and mental well-being documented that a balance point in between "too much" and "too little" tech use seemed optimal, which they referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017].] varied between reviewers, but they commonly wanted tools to provide incentives powerful enough to keep themselves on track in moments of temptation or absentmindedness.
-Overly strong incentives could backfire if they simply made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).
+    For example, a widely cited paper on the relationship between screen time and mental well-being documented that a balance point in between "too much" and "too little" tech use seemed optimal, which they referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017].
 
@@
 A complementary approach, which we explore in the present paper, is to investigate the landscape of digital self-control tools available online.
-Here, many hundreds of tools are exploring a wide range of strategies, from limiting the amount of functionality available on home screens (e.g.~the Android app \emph{LessPhone Launcher} [@mohanLessPhoneLauncher2019]) and hiding or modifying content on distracting websites (e.g.
+Here, many hundreds of tools are exploring a wide range of strategies, from limiting the amount of functionality available on home screens (e.g.\~the Android app \emph{LessPhone Launcher} [@mohanLessPhoneLauncher2019]) and hiding or modifying content on distracting websites (e.g.
 *Newsfeed Eradicator* [@jdevNewsFeedEradicator2019]), to gamifying self-control by tying device use to the wellbeing of virtual creatures (e.g.
@@
 
-However, few studies on digital self-control have so far taken this route:
-To the best of our knowledge, four published studies have described and/or categorised functionality in samples of digital self-control tools available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al.
+However, few studies on digital self-control have so far taken this route: To the best of our knowledge, four published studies have described and/or categorised functionality in samples of digital self-control tools available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al.
 [@vanvelthovenProblematicSmartphoneUse2018] presented aims and features in 21 tools identified via the software recommendation platform [alternativeto.net](https://alternativeto.net); Biskjær et al.
@@
 
-On this background, we extended existing studies of publicly available digital self-control tools by 
-(i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across stores and types of functionality, 
-(ii) analysing review content using 'reflexive' thematic analysis [@braunThematicAnalysis2018], 
-(iii) analysing popularity metrics' relationship to tools' design features, and 
-(iv) making our materials, analysis pipelines and data openly available for the purpose of transparency and reusability [@harozUpdatesCHISubmission2019].
+On this background, we extended existing studies of publicly available digital self-control tools by (i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across stores and types of functionality, (ii) analysing review content using 'reflexive' thematic analysis [@braunThematicAnalysis2018], (iii) analysing popularity metrics' relationship to tools' design features, and (iv) making our materials, analysis pipelines and data openly available for the purpose of transparency and reusability [@harozUpdatesCHISubmission2019].
 
@@
 identified these tools by downloading search results for the terms 'distraction', 'addiction', 'motivation', and 'self-control', individually as well as prefixed by 'internet', 'smartphone', and 'laptop', on the Chrome Web, Google Play, and Apple App stores.
-From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in a data set of 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).^[See Lyngs et al.
-[@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.]
+From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in a data set of 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).[^3]
+
+[^3]: See Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.
 
@@
 ## Data and analysis approach
+
 ### User numbers and ratings
+
 All stores displayed average ratings, as well as a count of how many ratings a tool had received.
@@
 The script then randomly sampled up to 30 user reviews from each of those tools (if a specific tool had fewer than 30 reviews, all of its reviews would be included), and started over.
-The script ran until more than 1,500 reviews were sampled in total. A tool could only be sampled from once.
+The script ran until more than 1,500 reviews were sampled in total.
+A tool could only be sampled from once.
 
@@
 At the time of scraping, `r num_missing` of the tools from Lyngs et al.'s review had become unavailable.
-Hence, we scraped information for `r nrow(meta_data)` tools: 
-`r num_w_info_by_store %>% filter(store == "chrome") %>% pull(n)` extensions from Chrome Web, 
-`r num_w_info_by_store %>% filter(store == "play") %>% pull(n)` apps from Google Play, and 
-`r num_w_info_by_store %>% filter(store == "apple") %>% pull(n)` apps from the Apple App store.
+Hence, we scraped information for `r nrow(meta_data)` tools: `r num_w_info_by_store %>% filter(store == "chrome") %>% pull(n)` extensions from Chrome Web, `r num_w_info_by_store %>% filter(store == "play") %>% pull(n)` apps from Google Play, and `r num_w_info_by_store %>% filter(store == "apple") %>% pull(n)` apps from the Apple App store.
 
@@
 -   C = Chrome Web, P = Google Play, A = Apple App store
-- Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}
+-   Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \\textbf{C}
 
@@
 However, digital self-control tools were seen as effectively helping reviewers take back control, save time, and improve their lives (*"I can happily say that this extension has changed my life!!!!
-It hides EVERY ANNOYING YouTube feature"*, R385, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock\typeGoal \textbf{C}), and many expressed deep appreciation towards the developers ("Who so ever is working behind this software, my deepest gratitude to them", R271, \href{https://chrome.google.com/webstore/detail/blocksite-block-websites/eiimnmioipafcokbfikbljfdeojpcgbh?hl=en-GB}{\emph{BlockSite}} \typeBlock \textbf{C}).
-Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \textbf{C}).
+It hides EVERY ANNOYING YouTube feature"*, R385, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \\textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock\typeGoal \\textbf{C}), and many expressed deep appreciation towards the developers ("Who so ever is working behind this software, my deepest gratitude to them", R271, \href{https://chrome.google.com/webstore/detail/blocksite-block-websites/eiimnmioipafcokbfikbljfdeojpcgbh?hl=en-GB}{\emph{BlockSite}} \typeBlock \\textbf{C}).
+Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \\textbf{C}).
 Importantly, the reviews also contained a wealth of information on how specific implementations of design patterns could fail to fulfil their purpose, and how they might be improved.
@@
 
-As we iteratively worked with the codes and excerpts, we developed three broad themes capturing digital self-control tools' (DSCTs) use contexts and design challenges, assembled from recurrent patterns in the early codes:
-(i) `r theme_one`,
-(ii) `r theme_two`, 
-(iii) `r theme_three`
+As we iteratively worked with the codes and excerpts, we developed three broad themes capturing digital self-control tools' (DSCTs) use contexts and design challenges, assembled from recurrent patterns in the early codes: (i) `r theme_one`, (ii) `r theme_two`, (iii) `r theme_three`
 
@@
 Users saw DSCTs as critical to avoiding time-wasting patterns of habitual or impulsive digital device use.
-Many reviews described this challenge as related to 'addictive' design patterns and other adversarial dynamics online, to which DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{Inbox When Ready for Gmail} \typeBlock\typeGoal\typeTrack \textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}).
-DSCTs seemed especially helpful when it was not possible to avoid using distracting services (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, \href{https://chrome.google.com/webstore/detail/kill-news-feed/hjobfcedfgohjkaieocljfcppjbkglfd}{Kill News Feed} \typeBlock \textbf{C}).
+Many reviews described this challenge as related to 'addictive' design patterns and other adversarial dynamics online, to which DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{Inbox When Ready for Gmail} \typeBlock\typeGoal\typeTrack \\textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \\textbf{A}).
+DSCTs seemed especially helpful when it was not possible to avoid using distracting services (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, \href{https://chrome.google.com/webstore/detail/kill-news-feed/hjobfcedfgohjkaieocljfcppjbkglfd}{Kill News Feed} \typeBlock \\textbf{C}).
 
-The usage contexts mentioned suggested that, aside from wanting to limit time on devices overall or on specific services (n = 62, 4\% of reviews), reviewers used DSTCs to help them focus on a given task amidst all too easily available distractions on their devices (n = 201, 13\%).
-The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding, expressed in some reviews as DSCTs helping overcome 'procrastination' (*"seeing my short-term to-do list nudges me to get back to work when I am opening a new website to procrastinate"*, R324, Dayboard \typeBlock\typeTrack \textbf{C}).
-In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \textbf{C}), and/or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack\typeReward \textbf{C}).
-Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11\%; "Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity", R776, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{\emph{Inbox When Ready for Gmail}} \typeBlock\typeGoal\typeTrack \textbf{C}), including studying, writing, or reading (n = 90, 6\%).
+The usage contexts mentioned suggested that, aside from wanting to limit time on devices overall or on specific services (n = 62, 4% of reviews), reviewers used DSTCs to help them focus on a given task amidst all too easily available distractions on their devices (n = 201, 13%).
+The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding, expressed in some reviews as DSCTs helping overcome 'procrastination' (*"seeing my short-term to-do list nudges me to get back to work when I am opening a new website to procrastinate"*, R324, Dayboard \typeBlock\typeTrack \\textbf{C}).
+In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \\textbf{C}), and/or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack\typeReward \\textbf{C}).
+Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11%; "Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity", R776, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{\emph{Inbox When Ready for Gmail}} \typeBlock\typeGoal\typeTrack \\textbf{C}), including studying, writing, or reading (n = 90, 6%).
 
 Another important use context was \emph{sleep}, where many reviewers said their digital devices often interfered with going to bed at the intended time.
-However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \textbf{P}).
+However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \\textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \\textbf{P}).
 
-**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges**
-In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely.
-This was often expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (n = 58, 3.8\%; *"This was the perfect solution to my phone addiction.
-It was consuming my life and my free time"*, R612, Forest \typeTrack\typeReward \textbf{P}).
-Sometimes reviewers used the clinical labels ADD or ADHD to explain why tools had been helpful to them (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack\typeReward \textbf{P}).
+**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges** In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely.
+This was often expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (n = 58, 3.8%; *"This was the perfect solution to my phone addiction.
+It was consuming my life and my free time"*, R612, Forest \typeTrack\typeReward \\textbf{P}).
+Sometimes reviewers used the clinical labels ADD or ADHD to explain why tools had been helpful to them (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack\typeReward \\textbf{P}).
 
@@
 
-DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 237, 16\%), a theme often expressed in the form of feature suggestions that would make tools more customisable.
+DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 237, 16%), a theme often expressed in the form of feature suggestions that would make tools more customisable.
 
 One aspect involved getting tools to accurately capture \emph{what} was considered a distraction.
-Thus, for block/removal tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (\emph{"I would like to block reddit but still be able to visit productivity related subreddits"}, R1401, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}).
-Similarly, self-tracking tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage""*, R989, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}).
+Thus, for block/removal tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (\emph{"I would like to block reddit but still be able to visit productivity related subreddits"}, R1401, Timewarp \typeBlock\typeTrack\typeGoal \\textbf{C}).
+Similarly, self-tracking tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage""*, R989, Mute Screen Time Tracker \typeGoal\typeTrack \\textbf{A}).
 
@@
 For example, tools for curbing nightly use received feature requests to handle variation in bedtime schedules (*"it needs to allow for the real life fact that people can't go to bed and wake up at the same time everyday.
-If it was done on your amount of sleep you get in hours that would be better"*, R1250, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \textbf{A}).
+If it was done on your amount of sleep you get in hours that would be better"*, R1250, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \\textbf{A}).
 
-**Sub-theme: Alignment between users' and tools' framing of distraction as important for navigating negative emotions elicited by DSCTs**
-Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor was the negative emotions that interventions could elicit.
-Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and essential, users also needed to carefully navigate the emotional reactions they could create (*"I hate it and love it at the same time.
-\emojiTearsOfJoy"*, R1299, Stay Focused \typeBlock\typeTrack \textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock\typeTrack \textbf{P}).
+**Sub-theme: Alignment between users' and tools' framing of distraction as important for navigating negative emotions elicited by DSCTs** Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor was the negative emotions that interventions could elicit.
+Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and essential, users also needed to carefully navigate the emotional reactions they could create (*"I hate it and love it at the same time.* \emojiTearsOfJoy", R1299, Stay Focused \typeBlock\typeTrack \\textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock\typeTrack \\textbf{P}).
 In order for experiences of "shock", "pressure", etc., to be useful, it was important that the user felt that a DSCT's purpose and framing of usage were closely aligned with their personal perspective.
@@
 Being reminded that one did not behave in line with one's intentions could be very helpful (*"This tells it like I need to be told sometimes...LOVE IT!
-Simple and effective :)"*, R690, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock\typeGoal \textbf{C}).
+Simple and effective :)"*, R690, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \\textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock\typeGoal \\textbf{C}).
 However, normative messaging could also backfire and be profoundly annoying if reviewers did not agree with it (*"I was a bit annoyed when Big Brother started getting judgemental on me.
-“You have hit your target of 50 pickups!” “You’ve now picked up your phone 90 times!” Well, first I didn’t set a target, and second, I don’t think that 54 minutes across a day should make me feel guilty.
-Especially since I was reading a book during that time"* R911, Moment \typeGoal\typeTrack \textbf{A}).
+"You have hit your target of 50 pickups!" "You've now picked up your phone 90 times!" Well, first I didn't set a target, and second, I don't think that 54 minutes across a day should make me feel guilty.
+Especially since I was reading a book during that time"* R911, Moment \typeGoal\typeTrack \\textbf{A}).
 
@@
 The reviews implied a spectrum of individual differences in where this 'Goldilocks level' was.
-A few reviewers stated that simply receiving reminders or knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, \emph{Mute Screen Time Tracker} \typeGoal\typeTrack \textbf{A}).
-More commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9\%).
+A few reviewers stated that simply receiving reminders or knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, \emph{Mute Screen Time Tracker} \typeGoal\typeTrack \\textbf{A}).
+More commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9%).
 
 Thus, tools implementing block/removal design patterns were often criticised if they were too easy to disable (*"This is fantastic but the thing is I can simply remove the extension from chrome which defeats the purpose.
-Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \textbf{C}).
+Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \\textbf{C}).
 However, if they were *too* difficult to override, they ran into another problem, namely how to deal with emergencies or otherwise unforeseen circumstances.
 One developer followed reviewers' suggestion to add a 'pause' button, only to receive critique that blocking was now too easy to override.
-Reviewers suggested that adding varying forms of friction to turn off blocking would be a solution, such as having to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock\typeTrack \textbf{P}), or simply adding passwords, which could be set by a trusted other.
+Reviewers suggested that adding varying forms of friction to turn off blocking would be a solution, such as having to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock\typeTrack \\textbf{P}), or simply adding passwords, which could be set by a trusted other.
 
-Similarly, for tools implementing reward/punishment, some reviewers suggested increasing incentives to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack\typeReward \textbf{C}).
-Here too, however, tools needed to find the right balance, because punishments that were perceived as too harsh might make users feel like failures (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed'"*, R1247, \href{https://itunes.apple.com/gb/app/sleeptown/id1210251567?mt=8\&uo=4}{SleepTown} \typeTrack\typeReward \textbf{P}).
+Similarly, for tools implementing reward/punishment, some reviewers suggested increasing incentives to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack\typeReward \\textbf{C}).
+Here too, however, tools needed to find the right balance, because punishments that were perceived as too harsh might make users feel like failures (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed'"*, R1247, \href{https://itunes.apple.com/gb/app/sleeptown/id1210251567?mt=8\&uo=4}{SleepTown} \typeTrack\typeReward \\textbf{P}).
 
-For tools including goal advancement design patterns, the 'right' level of incentive often related to the frequency or framing of features such as reminder pop-ups, which needed to be finely calibrated to be helpful rather than annoying (*"the todo popup on each page load is quite useful (...) it should be limited (across multiple tabs) to only pop-up every minute or so (...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock\typeGoal \textbf{C}).
+For tools including goal advancement design patterns, the 'right' level of incentive often related to the frequency or framing of features such as reminder pop-ups, which needed to be finely calibrated to be helpful rather than annoying (*"the todo popup on each page load is quite useful (...) it should be limited (across multiple tabs) to only pop-up every minute or so (...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock\typeGoal \\textbf{C}).
 
-**Sub-theme 1: Combining types of design patterns to find the Goldilocks level of incentive**
-In addition to tweaking parameters of a tool's current design patterns, many reviewers suggested including other types of patterns to make tools more effective.
+**Sub-theme 1: Combining types of design patterns to find the Goldilocks level of incentive** In addition to tweaking parameters of a tool's current design patterns, many reviewers suggested including other types of patterns to make tools more effective.
 Thus, some reviewers for self-tracking tools said that simply being informed about their behaviour was not sufficient for them to change anything.
@@
 Things I always intend to do, but don't do enough.
-But if these things came up every time I log into Youtube I'd definitely use them."}, R372, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \textbf{C}).
+But if these things came up every time I log into Youtube I'd definitely use them."}, R372, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \\textbf{C}).
 
-**Sub-theme 2: DSCTs should have low friction to start and use, but high friction to stop**
-Finally, finding the right level of friction also related to the level of effort required to use a tool.
+**Sub-theme 2: DSCTs should have low friction to start and use, but high friction to stop** Finally, finding the right level of friction also related to the level of effort required to use a tool.
 The amount of effort needed to \emph{initiate} use was important to keep low, because a lack of motivation could easily lead users to not use an otherwise effective tool (*"Most users have procrastination issues which make them postpone the moment in which they will click on the "Start focusing" button.
-Therefore those users (including me) will get distracted and they will click on that button after several hours, or even never, sometimes wasting a whole working day"*, R545, Focus \typeBlock\typeGoal\typeTrack \textbf{C}).
+Therefore those users (including me) will get distracted and they will click on that button after several hours, or even never, sometimes wasting a whole working day"*, R545, Focus \typeBlock\typeGoal\typeTrack \\textbf{C}).
 Therefore, a common feature request for blocking tools was the ability to set a schedule where blocking would automatically turn on.
-An interesting suggestion related to blocking tools alternating between work and break cycles:
-these could flip the ordering of cycles so as to begin with a break, followed by enforced blocking, which might lower the amount of motivation needed to initiate a focus session (*"Hitting the start button the very first time is hard!
-Starting in a short break would allow me to hit this without a second thought, because I have a few short moments to surf before the focus is forced."*, R232, Block \& Focus \typeBlock\typeTrack \textbf{C}).
+An interesting suggestion related to blocking tools alternating between work and break cycles: these could flip the ordering of cycles so as to begin with a break, followed by enforced blocking, which might lower the amount of motivation needed to initiate a focus session (*"Hitting the start button the very first time is hard!
+Starting in a short break would allow me to hit this without a second thought, because I have a few short moments to surf before the focus is forced."*, R232, Block & Focus \typeBlock\typeTrack \\textbf{C}).
 
@@
 Take a picture everyday?????
-Really?"}, R121, App Usage Tracker \typeTrack \textbf{A}).
+Really?"}, R121, App Usage Tracker \typeTrack \\textbf{A}).
 
@@
 For Chrome Web tools, the median number of users was `r format(chrome_user_stats$median_min_installs, big.mark=",")` (min = `r format(chrome_user_stats$min_min_installs, big.mark=",")`, max = `r format(chrome_user_stats$max_min_installs, big.mark=",")`, IQR = `r format(chrome_user_stats$IQR, big.mark=",")`).
-(The Apple App store did not display information about number of users.)
-In terms of the ratings given to tools, the median average rating was `r ratings_overall$median_rating` (min = `r ratings_overall$min_rating`, max = `r ratings_overall$max_rating`, IQR = `r round(ratings_overall$IQR, 1)`; ratings given on a scale from 1 to 5 stars on all stores).
+(The Apple App store did not display information about number of users.) In terms of the ratings given to tools, the median average rating was `r ratings_overall$median_rating` (min = `r ratings_overall$min_rating`, max = `r ratings_overall$max_rating`, IQR = `r round(ratings_overall$IQR, 1)`; ratings given on a scale from 1 to 5 stars on all stores).
 
@@
 \subsection{Using self-regulation measures to evaluate DSCTs and assess individual user needs}
+
 Our work provides additional data points for the discussion around how to evaluate effectiveness of DSCTs, and understand how needs for support vary between users.
@@
 ## Using psychological theory to explore interactions between design patterns and framings of 'distraction'
+
 Existing work has sometimes implied a single spectrum of 'friction' on which the intervention intensity of a given DSCT may be increased or decreased to find an optimal level \cite{kimGoalKeeperExploringInteraction2019, kimLocknTypeLockoutTask2019, kovacsConservationProcrastinationProductivity2019}.
@@
 [@lyngsSelfControlCyberspaceApplying2019], which summarises main mechanisms of self-regulation as viewed by current cognitive psychology.
-According to this model, the effects of DSCTs on self-regulation can be considered in terms of three factors: 
-(i) automatic habits and reflexes (System 1 control), 
-(ii) conscious goals and self-monitoring (System 2 control), and 
-(iii) the expected value of control (mediator of System 2 control).
+According to this model, the effects of DSCTs on self-regulation can be considered in terms of three factors: (i) automatic habits and reflexes (System 1 control), (ii) conscious goals and self-monitoring (System 2 control), and (iii) the expected value of control (mediator of System 2 control).
 
@@
 For example, to meet a goal of not checking Instagram at family dinners, you might need to suppress an impulse to check.
-If you fail, it can be because you had forgotten this goal, or because the 'expected value of control', which mediates the strength of System 2 control, was too low:
-The more value a conscious goal has (whether because its rewards are greater or more immediate, or because you are more confident in your ability to achieve it), the more powerfully it triggers action.
+If you fail, it can be because you had forgotten this goal, or because the 'expected value of control', which mediates the strength of System 2 control, was too low: The more value a conscious goal has (whether because its rewards are greater or more immediate, or because you are more confident in your ability to achieve it), the more powerfully it triggers action.
 In other words, you might fail to suppress your impulse to stay off your phone, because your valuation of staying off was not high enough to compete with the urge to check.
@@
 
-We can use this model to consider users' suggestions of combining distinct types of design patterns:
-it is possible that by combining design patterns which target distinct psychological mechanisms, each can be implemented at a more gentle level of intensity while still in aggregate provide enough input to the internal competition to bring about an intended behaviour change.
+We can use this model to consider users' suggestions of combining distinct types of design patterns: it is possible that by combining design patterns which target distinct psychological mechanisms, each can be implemented at a more gentle level of intensity while still in aggregate provide enough input to the internal competition to bring about an intended behaviour change.
 By contrast, tools that focus on a single design pattern targeting a single psychological mechanism may need to ramp up its intensity more highly before sufficient impetus for behaviour change is provided --- and in so doing be at greater risk of being perceived as overly annoying.
@@
 An alternative version which simply grew virtual trees without a countdown timer might need to correspondingly increase the magnitude of reward to remain useful, but in so doing be more likely to be perceived by users as too forceful or manipulative [@Lee2011Mining].
-^[We do note that the higher ratings given to tools combining two or more distinct design patterns has an alternative explanation: tools that combine design patterns may simply have received more developer attention, and so higher ratings could simply be due to such tools being better designed.]
+[^4]
+
+[^4]: We do note that the higher ratings given to tools combining two or more distinct design patterns has an alternative explanation: tools that combine design patterns may simply have received more developer attention, and so higher ratings could simply be due to such tools being better designed.
 
@@
 ## Navigating platform challenges for future DSCTs
+
 Looking towards the future, we note that the underlying technical infrastructures and platforms create very different conditions for the emergence of ecosystems of DSCT in the wild.
@@
 Unless users jailbreak their devices, no options exist for users to customise the look and function of mobile apps beyond the options that app developers chooses to provide.
-Moreover, the options available for self-tracking or blocking tools are dramatically different between Android and iOS:
-Because of Google's less restrictive approach to developer permissions, a diverse ecosystem exists on Android for apps that provide bespoke usage visualisation or customised blocking schedules, in addition to what Google provides themselves in their 'Digital Wellbeing' app.
-On iOS, by contrast, Apple's Screen Time tool was, at the time data for the present paper was scraped, the only option with which users could block functionality and see their usage of different apps^[A Screen Time API, introduced in the summer of 2021 after long campaigning by developers [@digitalwellnesswarriorsAppleLetDevelopers2018], now allows creation of blocking tools for parental control purposes, but does not enable apps for tracking and visualising screen time.].
+Moreover, the options available for self-tracking or blocking tools are dramatically different between Android and iOS: Because of Google's less restrictive approach to developer permissions, a diverse ecosystem exists on Android for apps that provide bespoke usage visualisation or customised blocking schedules, in addition to what Google provides themselves in their 'Digital Wellbeing' app.
+On iOS, by contrast, Apple's Screen Time tool was, at the time data for the present paper was scraped, the only option with which users could block functionality and see their usage of different apps[^5].
 As we observed in the user reviews, workarounds by iOS developers, such as asking users to routinely take screenshots of the battery page on their phone settings, seemed to be too effort-demanding on the part of users to be workable.
+
+[^5]: A Screen Time API, introduced in the summer of 2021 after long campaigning by developers [@digitalwellnesswarriorsAppleLetDevelopers2018], now allows creation of blocking tools for parental control purposes, but does not enable apps for tracking and visualising screen time.
 
@@
 ## Limitations and future work
-First, the data used in the present paper was scraped in March 2019. Hence, the findings presented may deviate from what would by learned by re-scraping the app and browser extension stores in 2022.
+
+First, the data used in the present paper was scraped in March 2019.
+Hence, the findings presented may deviate from what would by learned by re-scraping the app and browser extension stores in 2022.
 However, all of our data and materials are openly available, and so our process should be readily repeatable.
@@
 # Conclusion
+
 The landscape of digital self-control tools available online amounts to hundreds of thousands of natural experiments in supporting self-regulation of device use.
@@
 # Acknowledgments
+
 Funding: Ulrik Lyngs was funded by the UK Engineering and Physical Sciences Research Council (EPSRC) [grant number EP/J017728/2]; and the Carlsberg Foundation [grant number CF20-0678].

### 2022-04-29 paper.Rmd
tweak abstract

@@
 abstract: |
-  Much effort has been invested in designing digital systems that keep people 'hooked', but comparatively little is known about how designers can best support people in re-gaining control over digital device use.
-Online, hundreds of apps and browser extensions promise to help people self-regulate their use, and their reviews and popularity metrics can indicate which design patterns are found useful in the will.
-Moreover, they reveal how platforms such as Android and iOS differ in the ecosystems they enable for such tools, which has important practical implications for users struggling with self-regulation difficulties.
-We analysed user reviews, installation numbers, and ratings for 334 digital self-control tools (DSCTs) on the Google Play, Chrome Web, and Apple App stores, investigating what user reviews reveal about usage contexts and key design challenges, and how functionality relates to popularity metrics.
-Our thematic analysis of 1,529 reviews (sampled from a data set of 54,320 reviews scraped in March 2019) found that DSCTs are seen as highly important for staying focused on less instantly rewarding tasks when digital distractions are easily available.
-Users seek DSCTs that adapt to their personal definitions of 'distracting use', and provide incentives sufficient to affect behaviour change without feeling overly intrusive.
-DSCTs' interventions, even if highly useful, can elicit negative emotions, which is an important reason that the goals of the user and the DSCT should align.
-Tools that combine two or more types of design patterns (e.g., website blocking and goal reminders) received significantly higher ratings than those implementing a single type, which resonated with feature suggestions in the reviews.
-We discuss implications for research and design, including how basic psychological research on self-regulation can inform evaluation measures for DSCTs, as well as guide future exploration of how design patterns and framings of 'distraction' interact.
-
+  Much effort has been invested in designing digital systems that keep people 'hooked'.
+By contrast, comparatively little is known about how designers can support people in re-gaining control.
+Online, however, hundreds of apps and browser extensions promise to help people self-regulate use of digital devices.
+Reviews and popularity metrics for these digital self-control tools (DSCTs) can indicate which design patterns are useful in the wild.
+Moreover, they reveal how platforms such as Android and iOS differ in the ecosystems they enable for DSCTs, which has important practical implications for end users.
+We analysed reviews, installation numbers, and ratings for 334 DSCTs on the Google Play, Chrome Web, and Apple App stores, investigating what user reviews reveal about usage contexts and key design challenges, and how functionality relates to popularity metrics.
+Our thematic analysis of 1,529 reviews (sampled from a data set of 54,320 reviews scraped in March 2019) found that DSCTs are seen as highly important to stay focused on less instantly rewarding tasks when digital distractions are easily available.
+Users seek DSCTs that adapt to their personal definitions of 'distracting use', and which provide incentives sufficient to change behaviour without feeling overly intrusive.
+DSCTs' interventions can elicit negative emotions, which appeared to be an important reason that the DSCT accurately capture users' definition of distraction.
+Tools that combine two or more types of design patterns (e.g., website blocking and goal reminders) received slightly higher ratings than those implementing a single type.
+This mirrored suggestions in the reviews to combine different types of design patterns to provide a incentive for behaviour change that is 'just right'.
+  UPDATE FINAL SENTENCE AFTER REWRITE: We discuss implications for research and design, including how basic psychological research on self-regulation can inform evaluation measures for DSCTs, as well as guide future exploration of how design patterns and framings of 'distraction' interact.
 bibliography: references.bib

### 2022-04-29 paper.Rmd
incorporate initial proofing

@@
 the 'attention economy' [@davenportAttentionEconomyUnderstanding2001; @einsteinBlackOpsAdvertising2016; @wuAttentionMerchantsEpic2016]), many people routinely struggle to exercise self-control over their device use [@coxDesignFrictionsMindful2016; @koNUGUGroupbasedIntervention2015; @marottaOnlineDistractionsWebsite2017; @tranModelingEngagementDisengagementCycle2019; @hinikerMyTimeDesigningEvaluating2016].
-As a result, public debate about negative effects of 'persuasive technologies' has been growing [@OptimizingEngagementUnderstanding2019; @orlowski2020] alongside increased political pressure on technology platforms to safeguard users [@allpartyparliamentarygrouponsocialmediaandyoungpeoplesmentalhealthandwellbeingNewFiltersManageImpact2019].
+As a result, public debate about negative effects of 'persuasive technologies' has grown [@OptimizingEngagementUnderstanding2019; @orlowski2020], alongside increased political pressure on technology platforms to safeguard users [@allpartyparliamentarygrouponsocialmediaandyoungpeoplesmentalhealthandwellbeingNewFiltersManageImpact2019].
 
-On online stores, a flourishing market exists for 'digital self-control tools', where hundreds of apps and browser extensions promise to help people take back control over their device use.
+On online stores, a flourishing market exists for 'digital self-control tools' (DSCTs), where hundreds of apps and browser extensions promise to help people take back control over their device use.
 These tools offer interventions such as blocking apps or removing distracting website elements, tracking and visualising usage, or providing rewards or punishments for how devices are used [@lyngsSelfControlCyberspaceApplying2019].
@@
 
-In academia, a growing amount of HCI research is focusing on developing and evaluating various design patterns meant to help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019][^1].
-This area of HCI work has gradually been building an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
-However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how much people vary in how useful they find particular interventions, and how specific design patterns' usefulness depends on details of their implementation, such as the level of 'friction' they introduce [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].
+In academia, a growing amount of HCI research is similarly focused on developing and evaluating various design patterns that help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019][^1].
+This work has gradually built an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
+However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how implementation details of specific design patterns --- such as the level of 'friction' --- influence their usefulness, and how much people vary in how useful they find particular interventions [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].
 
-[^1]: We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders), and the term 'digital self-control tool' (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook's newsfeed).
-    We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user's usual digital environment, or when it is evaluated in a research study.
+[^1]: We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders).
+    We use the term 'digital self-control tool' (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook's newsfeed).
+    We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user's usual digital environment, or when it is being evaluated in a research study.
 
-To answer such questions, most HCI studies have taken an experimental approach and compared user behaviour and perceptions when applying a given intervention to its absence and/or some alternative intervention [@lyngs2021].
+To answer such questions, most existing studies have taken an experimental approach and compared user behaviour and perceptions when applying a given intervention to its absence and/or some alternative intervention [@lyngs2021].
 Yet, while controlled studies are appropriate and feasible for evaluating and comparing a small number of design patterns, they are difficult to scale to broadly assess a large numbers of patterns and implementations.
 
-A complementary approach, which may help scope the range of design patterns and implementations to explore in controlled studies, is to investigate the existing landscape of digital self-control tools on app and web stores [@lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018].
-Here, user reviews and popularity metrics can provide information about the relative utility of different patterns and implementations, contexts of use, and user needs 'in the wild'.
+A complementary approach, which can help scope the range of design patterns and implementations to explore in controlled studies, is to investigate the existing landscape of DSCTs on app and web stores [@lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018].
+Here, user reviews and popularity metrics can provide information about user needs, contexts of use, and the relative utility of different patterns and implementations 'in the wild'.
 However, few studies to date have taken this approach.
-Hence, in the present paper, we contribute to the growing HCI literature on design patterns for digital self-control by using app and web store information as a resource for understanding user self-experimentation with digital self-control tools (DSCTs).
+Hence, in the present paper, we contribute to the HCI literature on design patterns for digital self-control by using app and web store information as a resource for understanding user self-experimentation with DSCTs.
 Specifically, we present an analysis of user reviews, functionality, installation numbers, and ratings for 334 digital self-control tools, drawn from Lyngs et al.
@@
 
-Our analysis of user reviews finds that digital self-control tools are seen as crucial to support focus on less instantly rewarding goals when digital distractions are easily available, specifically in contexts related to productivity (including work, studying, reading, and writing), and going to sleep.
-DSCTs seem especially useful for people who describe themselves as 'addicts' or as struggling with distraction due to conditions like ADHD.
+Our analysis of user reviews finds that digital self-control tools are seen as highly important to supports focus on less instantly rewarding goals when digital distractions are easily available, specifically in contexts related to productivity (including work, studying, reading, and writing), and going to sleep.
+DSCTs appeared to be especially useful for people who describe themselves as 'addicts' or as struggling with distraction due to conditions like ADHD.
 
-Users want DSCTs to flexibly adapt to their personal definition of 'distraction', as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
-A key reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions.
-This may cause the effects of interventions to be counterproductive, especially when users perceive DSCTs' goals or framing to differ from their own.
+Users want DSCTs to adapt to their personal definition of 'distraction', as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
+One important reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions.
+This may cause the effects of interventions to be counterproductive, especially if users perceive DSCTs' goals or framing to differ from their own.
 
 Finally, users want DSCTs to provide carefully calibrated incentives that are sufficient to incentivise intended behaviour change without feeling overly intrusive.
-In our analysis, this 'Goldilocks level'[^2] varied between reviewers, but they commonly wanted tools to provide incentives powerful enough to keep themselves on track in moments of temptation or absentmindedness.
-Overly strong incentives could backfire if they simply made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).
+In our analysis, DSCTs 'Goldilocks level'[^2] varied between reviewers.
+However, reviewers commonly wanted tools that provided incentives strong enough to keep themselves on track in moments of temptation or absentmindedness.
+Overly strong incentives could, however, backfire if they made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).
 
@@
 One way for tools to provide properly calibrated incentives might be by carefully combining different types of design patterns (e.g., using both block/removal and goal-advancement features).
-This was often asked for in reviews' feature suggestions, and supported by higher overall ratings given to tools combining two or more types compared to one.
+This was often asked for in reviews' feature suggestions, and tentatively supported by higher overall ratings given to tools that combine two or more types compared to just one.
 A final consideration related to the amount of effort involved in operating a tool, where low friction to set up and initiate use (e.g., via schedules where a tool automatically turns on), but high friction to override it, was important.
@@
 For example, in mental health research, thematic analysis of user reviews has been used to explore factors important for optimising user experience and support engagement in apps for cognitive behavioural therapy [@bakkerMentalHealthSmartphone2016; @huguetSystematicReviewCognitive2016; @luiEvidencebasedAppsReview2017; @sucalaAnxietyThereApp2017; @stawarzUserExperienceCognitive2018].
-Similar research found that content in popular apps marketed as reducing symptoms of anxiety and worry was largely inconsistent with evidence-based treatments [@kertzReviewFreeIPhone2017].
-This opened avenues for further research including how established CBT practices could more effectively be disseminated and, conversely, what might be learned from alternative approaches that proved popular in the wild.
-HCI research on design patterns for digital self-control might in a similar way derive actionable insights from the large number of tools for this purpose available online.
-
-However, few studies on digital self-control have so far taken this route: To the best of our knowledge, four published studies have described and/or categorised functionality in samples of digital self-control tools available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al.
+<!-- Similar research found that content in popular apps marketed as reducing symptoms of anxiety and worry was largely inconsistent with evidence-based treatments [@kertzReviewFreeIPhone2017].
+--> <!-- This opened avenues for further research including how established CBT practices could more effectively be disseminated and, conversely, what might be learned from alternative approaches that proved popular in the wild.
+--> <!-- HCI research on design patterns for digital self-control might in a similar way derive actionable insights from the large number of tools for this purpose available online.
+--> However, few studies on digital self-control have so far taken this route: To the best of our knowledge, four published studies have described and/or categorised functionality in samples of digital self-control tools available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al.
 [@vanvelthovenProblematicSmartphoneUse2018] presented aims and features in 21 tools identified via the software recommendation platform [alternativeto.net](https://alternativeto.net); Biskjær et al.
@@
 Only one of these studies analysed user reviews, namely Roffarello & Russis [@mongeroffarelloRaceDigitalWellbeing2019], which included thematic analysis of 1,128 reviews from 42 smartphone apps, and none incorporated popularity metrics.
-Moreover, Roffarello & Russis' analysis of user reviews was restricted to 42 smartphone apps drawn from the Google Play store and did not specify how the reviews were distributed across apps with different functionality.
+Moreover, Roffarello & Russis' analysis was restricted to 42 smartphone apps drawn from the Google Play store and did not specify how the reviews were distributed across apps with different functionality.
 Thematic analysis is likely to yield different results depending on how reviews are sampled across tools with different functionality, and depending on the analysis strategy [@chenARminerMiningInformative2014; @braunThematicAnalysis2018].
@@
 
-On this background, we extended existing studies of publicly available digital self-control tools by (i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across stores and types of functionality, (ii) analysing review content using 'reflexive' thematic analysis [@braunThematicAnalysis2018], (iii) analysing popularity metrics' relationship to tools' design features, and (iv) making our materials, analysis pipelines and data openly available for the purpose of transparency and reusability [@harozUpdatesCHISubmission2019].
+On this background, we extended existing studies of publicly available digital self-control tools by (i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across stores and types of functionality, (ii) analysing review content using 'reflexive' thematic analysis [@braunThematicAnalysis2018], and (iii) analysing popularity metrics' relationship to tools' design features, and (iv) making our materials, analysis pipelines and data openly available for the purpose of transparency and reusability [@harozUpdatesCHISubmission2019].
 
@@
 
-Data, analysis scripts, and materials for the present paper are available at \url{https://osf.io/cdqub/}.
+Data, analysis scripts, and materials for the present paper are available at [osf.io/cdqub/](https://osf.io/cdqub/).
 

### 2022-05-09 paper.Rmd
calculate quantiles

@@
   summarise(num_reviews = n()) %>% 
-  summarise(median = median(num_reviews),
-            mean = mean(num_reviews),
-            min = min(num_reviews),
-            max = max(num_reviews),
-            IQR = IQR(num_reviews)))
+  summarise(mean = mean(num_reviews),
+            min = quantile(num_reviews, 0),
+            q_25 = quantile(num_reviews, 0.25),
+            median = quantile(num_reviews, 0.5),
+            q_75 = quantile(num_reviews, 0.75),
+            max = quantile(num_reviews, 1))
 
 (word_count_summary <- reviews_w_main_cat_data %>% 
-  summarise(median = median(review_word_count),
-            mean = mean(review_word_count),
-            min = min(review_word_count),
-            max = max(review_word_count),
-            IQR = IQR(review_word_count)))
+  summarise(mean = mean(review_word_count),
+            min = quantile(review_word_count, 0),
+            q_25 = quantile(review_word_count, 0.25),
+            median = quantile(review_word_count, 0.5),
+            q_75 = quantile(review_word_count, 0.75),
+            max = quantile(review_word_count, 1)))
 

### 2022-06-10 paper.Rmd
bring Rmd in line with accepted version

--- 2022-04-29 paper.Rmd
@@
 journal: International Journal of Human-Computer Studies
-title: "The Goldilocks Level of Friction: Using User Numbers, Ratings, and Reviews to Evaluate Digital Self-Control Tools"
-keywords: Attention, self-control, self-regulation, distraction, digital wellbeing, focus
+title: |
+  The Goldilocks Level of Support:\
+  Using User Reviews, Ratings, and Installation Numbers to Investigate Digital Self-Control Tools Tools
+keywords: Attention, distraction, digital self-control, digital wellbeing, focus, Goldilocks principles, self-regulation
 abstract: |
@@
 Reviews and popularity metrics for these digital self-control tools (DSCTs) can indicate which design patterns are useful in the wild.
-Moreover, they reveal how platforms such as Android and iOS differ in the ecosystems they enable for DSCTs, which has important practical implications for end users.
+Moreover, they reveal how platforms like Android and iOS differ in the ecosystems they enable for DSCTs, which has important implications for end users.
 We analysed reviews, installation numbers, and ratings for 334 DSCTs on the Google Play, Chrome Web, and Apple App stores, investigating what user reviews reveal about usage contexts and key design challenges, and how functionality relates to popularity metrics.
-Our thematic analysis of 1,529 reviews (sampled from a data set of 54,320 reviews scraped in March 2019) found that DSCTs are seen as highly important to stay focused on less instantly rewarding tasks when digital distractions are easily available.
-Users seek DSCTs that adapt to their personal definitions of 'distracting use', and which provide incentives sufficient to change behaviour without feeling overly intrusive.
-DSCTs' interventions can elicit negative emotions, which appeared to be an important reason that the DSCT accurately capture users' definition of distraction.
-Tools that combine two or more types of design patterns (e.g., website blocking and goal reminders) received slightly higher ratings than those implementing a single type.
-This mirrored suggestions in the reviews to combine different types of design patterns to provide a incentive for behaviour change that is 'just right'.
-  UPDATE FINAL SENTENCE AFTER REWRITE: We discuss implications for research and design, including how basic psychological research on self-regulation can inform evaluation measures for DSCTs, as well as guide future exploration of how design patterns and framings of 'distraction' interact.
+Our thematic analysis of 1,529 reviews (sampled from a data set of 53,978 distinct reviews scraped in March 2019) found that DSCTs are seen as highly important for focusing on less instantly rewarding tasks when digital distractions are easily available.
+Users seek DSCTs that adapt to their personal definitions of distraction, and provide support that is sufficient to change behaviour without feeling too coercive.
+Reviewers suggested combining design patterns to provide a level of support that is 'just right'.
+This was mirrored in the ratings, where tools combined different types of design patterns (e.g., website blocking and goal reminders) tended to receive higher ratings than those implementing a single type.
+We discuss implications for research and design, including how design patterns in DSCTs interact, and how psychological reactance to DSCTs can be reduced.
+
 bibliography: references.bib
@@
 the 'attention economy' [@davenportAttentionEconomyUnderstanding2001; @einsteinBlackOpsAdvertising2016; @wuAttentionMerchantsEpic2016]), many people routinely struggle to exercise self-control over their device use [@coxDesignFrictionsMindful2016; @koNUGUGroupbasedIntervention2015; @marottaOnlineDistractionsWebsite2017; @tranModelingEngagementDisengagementCycle2019; @hinikerMyTimeDesigningEvaluating2016].
-As a result, public debate about negative effects of 'persuasive technologies' has grown [@OptimizingEngagementUnderstanding2019; @orlowski2020], alongside increased political pressure on technology platforms to safeguard users [@allpartyparliamentarygrouponsocialmediaandyoungpeoplesmentalhealthandwellbeingNewFiltersManageImpact2019].
+In HCI, a growing amount of research attempts to address this challenge by developing and evaluating design patterns that help users self-regulate[^1] [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019].
+This work is gradually building an understanding of how interventions ranging from ambient visualisations of time spent [@whittakerDonWasteMy2016] and distraction blocking [@kimGoalKeeperExploringInteraction2019] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
 
-On online stores, a flourishing market exists for 'digital self-control tools' (DSCTs), where hundreds of apps and browser extensions promise to help people take back control over their device use.
-These tools offer interventions such as blocking apps or removing distracting website elements, tracking and visualising usage, or providing rewards or punishments for how devices are used [@lyngsSelfControlCyberspaceApplying2019].
-Similarly, Apple and Google now offer various usage logging and self-limiting features in their 'ScreenTime' and 'Digital Wellbeing' tools, respectively [@appleIOS12Introduces2018; @gonzalezResearchGoogleNew2018; @googleDigitalWellbeing2018], and even Facebook have introduced tools 'for managing your time on Facebook and Instagram' [@ranadiveNewToolsManage2018].
+Most of this research has taken an experimental approach and compared user behaviour and perceptions when applying a given intervention to its absence and/or some alternate intervention [@lyngs2021]. 
+In the present paper, we take an alternative approach.
+On online stores for apps and browser extensions, hundreds of 'digital self-control tools' (DSCTs) provide similar interventions that block apps or hide distracting website elements, track and visualise use, provide rewards or punishments for how devices are used, and much more [@lyngsSelfControlCyberspaceApplying2019].
+User reviews and popularity metrics from these tools might provide valuable complementary information about the utility of different design patterns and implementations 'in the wild', and help scope future controlled studies.
+However, a limited number of studies have investigated what we might learn from this source of data.
 
-In academia, a growing amount of HCI research is similarly focused on developing and evaluating various design patterns that help users self-regulate digital device use [@lottridgeBrowserDesignImpacts2012; @kimGoalKeeperExploringInteraction2019; @pinderPushAwaySmartphone2019; @tsengOvercomingDistractionsTransitions2019; @mongeroffarelloRaceDigitalWellbeing2019; @lyngsSelfControlCyberspaceApplying2019][^1].
-This work has gradually built an understanding of how a range of design patterns from visualisations of time spent [@whittakerDonWasteMy2016] to goal-setting with social support [@koNUGUGroupbasedIntervention2015] can help people stay in control.
-However, we are still in the early days of answering basic questions including how to define and measure 'success' for these design patterns, how implementation details of specific design patterns --- such as the level of 'friction' --- influence their usefulness, and how much people vary in how useful they find particular interventions [@kimGoalKeeperExploringInteraction2019; @markEffectsIndividualDifferences2018; @tseng2019; @lyngsTellMeWhat2018; @hinikerMovingGreatScreen2019].
+Hence, in this paper, we explore how user reviews and other store information can contribute to HCI research on design patterns for digital self-control.
+Specifically, we analyse user reviews, functionality, installation numbers, and ratings for 334 DSCTs on the Google Play, Chrome Web, and Apple App stores (drawn from [@lyngsSelfControlCyberspaceApplying2019]), investigating user needs, contexts and purposes of use, and key design challenges.
 
-[^1]: We use the term 'design patterns' to refer to specific functions intended to support people in self-regulating their digital device use (e.g., hiding newsfeeds, providing goal reminders).
-    We use the term 'digital self-control tool' (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook's newsfeed).
-    We occasionally use the term 'intervention' to refer to the implementation of a design pattern in the context of it changing a user's usual digital environment, or when it is being evaluated in a research study.
+Our analysis of user reviews found that DSCTs are seen as highly important to support focus on less instantly rewarding tasks when digital distractions are easily available, especially in contexts related to productivity (including work, studying, reading, and writing), and going to sleep.
+DSCTs seemed particularly useful for reviewers describing themselves as 'addicts' or otherwise struggling intensely with distraction, for example due to conditions like ADHD.
 
-To answer such questions, most existing studies have taken an experimental approach and compared user behaviour and perceptions when applying a given intervention to its absence and/or some alternative intervention [@lyngs2021].
-Yet, while controlled studies are appropriate and feasible for evaluating and comparing a small number of design patterns, they are difficult to scale to broadly assess a large numbers of patterns and implementations.
+Reviewers wanted DSCTs to flexibly adapt to their personal definition of 'distracting use', as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
+One important reason was that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions.
+This could cause their interventions to be counterproductive, especially if users perceived DSCTs' framing of distracting use to differ from their own.
 
-A complementary approach, which can help scope the range of design patterns and implementations to explore in controlled studies, is to investigate the existing landscape of DSCTs on app and web stores [@lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018].
-Here, user reviews and popularity metrics can provide information about user needs, contexts of use, and the relative utility of different patterns and implementations 'in the wild'.
-However, few studies to date have taken this approach.
-Hence, in the present paper, we contribute to the HCI literature on design patterns for digital self-control by using app and web store information as a resource for understanding user self-experimentation with DSCTs.
-Specifically, we present an analysis of user reviews, functionality, installation numbers, and ratings for 334 digital self-control tools, drawn from Lyngs et al.
-[@lyngsSelfControlCyberspaceApplying2019] who reviewed functionality in such tools on the Google Play, Chrome Web, and Apple App stores.
+Finally, reviewers wanted DSCTs to provide carefully calibrated support that is sufficient to support intended behaviour change without feeling too coercive.
+This 'Goldilocks level'[^2] varied between reviews, but most wanted DSCTs to provide support powerful enough to keep the user on track in moments of temptation or absentmindedness.
+Thus, blocking tools were often criticised if they were too easy to override because they could then fail to hold people accountable in moments of temptation.
+On the other hand, DSCTs were also criticised if their incentives were *too* strong, for example if blocking tools did not include some way to handle genuine emergency use, or if tools providing reward/punishment made users feel dejected when they failed to use their devices as intended.
 
-Our analysis of user reviews finds that digital self-control tools are seen as highly important to supports focus on less instantly rewarding goals when digital distractions are easily available, specifically in contexts related to productivity (including work, studying, reading, and writing), and going to sleep.
-DSCTs appeared to be especially useful for people who describe themselves as 'addicts' or as struggling with distraction due to conditions like ADHD.
-
-Users want DSCTs to adapt to their personal definition of 'distraction', as seen in common requests for customisation of *what* and *when* particular usage patterns should be blocked or tracked.
-One important reason that users' and tools' conceptualisation of distraction should be in close alignment is that DSCTs' interventions, useful as they may be, can elicit negative emotional reactions.
-This may cause the effects of interventions to be counterproductive, especially if users perceive DSCTs' goals or framing to differ from their own.
-
-Finally, users want DSCTs to provide carefully calibrated incentives that are sufficient to incentivise intended behaviour change without feeling overly intrusive.
-In our analysis, DSCTs 'Goldilocks level'[^2] varied between reviewers.
-However, reviewers commonly wanted tools that provided incentives strong enough to keep themselves on track in moments of temptation or absentmindedness.
-Overly strong incentives could, however, backfire if they made users feel like a failure (in the case of reward/punishment tools) or did not allow some way to handle genuine emergency use (in the case of block/removal tools).
-
-[^2]: After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'.
-    The concept of the 'just right amount' is often referred to as the 'Goldilocks principle'.
-    For example, a widely cited paper on the relationship between screen time and mental well-being documented that a balance point in between "too much" and "too little" tech use seemed optimal, which they referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017].
-
-One way for tools to provide properly calibrated incentives might be by carefully combining different types of design patterns (e.g., using both block/removal and goal-advancement features).
-This was often asked for in reviews' feature suggestions, and tentatively supported by higher overall ratings given to tools that combine two or more types compared to just one.
-A final consideration related to the amount of effort involved in operating a tool, where low friction to set up and initiate use (e.g., via schedules where a tool automatically turns on), but high friction to override it, was important.
+One way DSCTs might provide properly calibrated support is by combining different types of design patterns (e.g., combining distraction blocking with goal-reminders).
+This was sometimes requested in the reviews, and tentatively supported by rating scores, where tools that combine two or more types of design patterns received higher ratings than those implementing a single type.
 
@@
 
-In HCI, a range of studies have developed and evaluated design patterns for digital self-control, ranging from ambient visualisations of time spent [@whittakerDonWasteMy2016] and distraction blocking [@kimGoalKeeperExploringInteraction2019] to goal-setting and social support [@koNUGUGroupbasedIntervention2015].
-This research is starting to build an evidence base of how specific interventions to support self-regulation influence user behaviour and perceptions, but basic questions remain open.
-These include *conceptual challenges* (e.g., what metrics digital self-control tools should seek to optimise, as simply measuring impact on 'screen time' is deemed inadequate by most researchers [@cecchinatoDesigningDigitalWellbeing2019; @hinikerMovingGreatScreen2019; @lukoffDigitalWellbeingWay2019]); related gaps in our understanding of *individual variation in user needs* (such as how how widely people differ in what they consider 'distracting' usage [@cecchinatoDesigningDigitalWellbeing2019; @lukoffDigitalWellbeingWay2019]); and *practical implementation challenges* (such as how much 'friction' tools should apply to hold users accountable for their past preferences [@kimGoalKeeperExploringInteraction2019; @coxDesignFrictionsMindful2016; @lyngsTellMeWhat2018]).
+In recent years, HCI researchers have started to build an understanding of how specific design patterns intended to support self-regulation over digital device can influence user behaviour and perceptions [@whittakerDonWasteMy2016; @kimGoalKeeperExploringInteraction2019; @kimLocknTypeLockoutTask2019; @kovacsRotatingOnlineBehavior2018].
+However, basic questions still remain open.
+These include *conceptual challenges* (e.g., what outcome measures we should seek to optimise, as simply measuring impact on 'screen time' is deemed inadequate by most researchers [@cecchinatoDesigningDigitalWellbeing2019; @hinikerMovingGreatScreen2019; @lukoffDigitalWellbeingWay2019]); related gaps in our understanding of *individual variation in user needs* (such as how how widely people differ in what they consider 'distracting' use [@cecchinatoDesigningDigitalWellbeing2019; @lukoffDigitalWellbeingWay2019]); and *practical implementation challenges* (such as how much 'friction' tools should apply to hold users accountable for their past preferences [@kimGoalKeeperExploringInteraction2019; @coxDesignFrictionsMindful2016; @lyngsTellMeWhat2018; @kimLocknTypeLockoutTask2019]).
 
-One approach to addressing these questions is to evaluate new or existing interventions in lab and field studies, and/or conduct surveys or interviews with users.
-A complementary approach, which we explore in the present paper, is to investigate the landscape of digital self-control tools available online.
-Here, many hundreds of tools are exploring a wide range of strategies, from limiting the amount of functionality available on home screens (e.g.\~the Android app \emph{LessPhone Launcher} [@mohanLessPhoneLauncher2019]) and hiding or modifying content on distracting websites (e.g.
-*Newsfeed Eradicator* [@jdevNewsFeedEradicator2019]), to gamifying self-control by tying device use to the wellbeing of virtual creatures (e.g.
-*Forest* [@seekrtechForestStayFocused2018]).
+One approach to addressing these questions is to evaluate new or existing interventions in lab and field studies.
+While this is appropriate and feasible for evaluating and comparing a small number of design patterns, it is difficult to scale to broadly assess a large numbers of patterns and implementations.
 
-Some of these tools have millions of users [@lyngsSelfControlCyberspaceApplying2019], and they effectively represent a myriad of natural studies in which individuals self-experiment with tools that represent not only one or more design patterns, but specific ways to implement those patterns [@daskalovaSleepCoacher2018; @leeSelfExperimentation2017].
-Indicators of the outcome of these experiments include user reviews, store ratings, and installation numbers, which may provide valuable information about the usefulness of different design patterns and implementations.
+A complementary approach is to investigate the landscape of DSCTs available online.
+Here, hundreds of tools are exploring a wide range of strategies, from hiding or modifying content on distracting websites (e.g. *Newsfeed Eradicator* [@jdevNewsFeedEradicator2019]) or limiting the amount of functionality available on home screens (e.g. the Android app *LessPhone Launcher* [@mohanLessPhoneLauncher2019]), to gamifying self-control by tying device use to the wellbeing of virtual creatures (e.g. *Forest* [@seekrtechForestStayFocused2018]).
+Some of these tools have millions of users [@lyngsSelfControlCyberspaceApplying2019], and they effectively amount to a myriad of natural studies in which individuals self-experiment with tools that represent not only one or more design patterns, but specific ways to implement those patterns [@daskalovaSleepCoacher2018; @leeSelfExperimentation2017].
+Indicators of the outcome of these experiments include user reviews, store ratings, and installation numbers, which can provide valuable information about the usefulness of different design patterns and implementations.
 
-Other areas of HCI research have benefited from taking such an approach.
+Other areas of HCI research have benefited from such an approach.
 For example, in mental health research, thematic analysis of user reviews has been used to explore factors important for optimising user experience and support engagement in apps for cognitive behavioural therapy [@bakkerMentalHealthSmartphone2016; @huguetSystematicReviewCognitive2016; @luiEvidencebasedAppsReview2017; @sucalaAnxietyThereApp2017; @stawarzUserExperienceCognitive2018].
-<!-- Similar research found that content in popular apps marketed as reducing symptoms of anxiety and worry was largely inconsistent with evidence-based treatments [@kertzReviewFreeIPhone2017].
---> <!-- This opened avenues for further research including how established CBT practices could more effectively be disseminated and, conversely, what might be learned from alternative approaches that proved popular in the wild.
---> <!-- HCI research on design patterns for digital self-control might in a similar way derive actionable insights from the large number of tools for this purpose available online.
---> However, few studies on digital self-control have so far taken this route: To the best of our knowledge, four published studies have described and/or categorised functionality in samples of digital self-control tools available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al.
+However, few studies on digital self-control have taken this route: To the best of our knowledge, four published studies have described and/or categorised functionality in samples of DSCTs available online [@biskjaerTakingActionDistraction2016; @lyngsSelfControlCyberspaceApplying2019; @mongeroffarelloRaceDigitalWellbeing2019; @vanvelthovenProblematicSmartphoneUse2018]: Velthoven et al.
 [@vanvelthovenProblematicSmartphoneUse2018] presented aims and features in 21 tools identified via the software recommendation platform [alternativeto.net](https://alternativeto.net); Biskjær et al.
-[@biskjaerTakingActionDistraction2016] presented a taxonomy of functionality based on 10 tools from the Google Play, Chrome Web, and Apple App stores, and online tech magazines; Roffarello & Russis [@mongeroffarelloRaceDigitalWellbeing2019] presented a taxonomy of functionality based on 42 mobile apps from the Google Play store; and Lyngs et al.
+[@biskjaerTakingActionDistraction2016] presented a taxonomy of functionality based on 10 tools from the Google Play, Chrome Web, and Apple App stores, and online tech magazines; Monge Roffarello & De Russis [@mongeroffarelloRaceDigitalWellbeing2019] presented a taxonomy of functionality based on 42 mobile apps from the Google Play store; and Lyngs et al.
 [@lyngsSelfControlCyberspaceApplying2019] presented a taxonomy of functionality based on 367 tools from the Google Play, Chrome Web and Apple App stores, which was mapped to psychological mechanisms in a dual systems model.
 
-Only one of these studies analysed user reviews, namely Roffarello & Russis [@mongeroffarelloRaceDigitalWellbeing2019], which included thematic analysis of 1,128 reviews from 42 smartphone apps, and none incorporated popularity metrics.
-Moreover, Roffarello & Russis' analysis was restricted to 42 smartphone apps drawn from the Google Play store and did not specify how the reviews were distributed across apps with different functionality.
+Only one of these studies analysed user reviews, namely Monge Roffarello & De Russis [@mongeroffarelloRaceDigitalWellbeing2019], which included thematic analysis of 1,128 reviews from 42 smartphone apps, and none incorporated popularity metrics.
+Moreover, Monge Roffarello & De Russis' analysis of user reviews was restricted to 42 apps drawn from the Google Play store and did not specify how the reviews were distributed across apps with different functionality.
 Thematic analysis is likely to yield different results depending on how reviews are sampled across tools with different functionality, and depending on the analysis strategy [@chenARminerMiningInformative2014; @braunThematicAnalysis2018].
@@
 
-On this background, we extended existing studies of publicly available digital self-control tools by (i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across stores and types of functionality, (ii) analysing review content using 'reflexive' thematic analysis [@braunThematicAnalysis2018], and (iii) analysing popularity metrics' relationship to tools' design features, and (iv) making our materials, analysis pipelines and data openly available for the purpose of transparency and reusability [@harozUpdatesCHISubmission2019].
+On this background, we extended existing studies of DSCTs by (i) scraping installation numbers, ratings, and user reviews from a broad sample of tools across different stores and types of functionality, (ii) analysing reviews using 'reflexive' thematic analysis [@braunThematicAnalysis2018], (iii) analysing popularity metrics' relationship to design patterns, and (iv) making our materials, analysis scripts and data openly available in the interest of transparency and reusability [@harozUpdatesCHISubmission2019].
 
@@
 
-Data, analysis scripts, and materials for the present paper are available at [osf.io/cdqub/](https://osf.io/cdqub/).
+Data, analysis scripts, and materials for the present paper are available at <https://osf.io/cdqub/>.
 
-We investigated the 367 digital self-control tools reviewed by Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019], as this paper provided the largest currently available review of DSCTs.
+We investigated the 367 digital self-control tools reviewed by Lyngs et al.
+[@lyngsSelfControlCyberspaceApplying2019], as this paper provided the largest available review of DSCTs.
 Lyngs et al.
-identified these tools by downloading search results for the terms 'distraction', 'addiction', 'motivation', and 'self-control', individually as well as prefixed by 'internet', 'smartphone', and 'laptop', on the Chrome Web, Google Play, and Apple App stores.
-From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in a data set of 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).[^3]
+identified these tools by downloading search results on the Chrome Web, Google Play, and Apple App stores in August 2018 for the terms 'distraction', 'addiction', 'motivation', and 'self-control', individually as well as prefixed by 'internet', 'smartphone', and 'laptop'.
+From the initial results (4890 distinct apps and extensions), they included tools explicitly designed to help people self-regulate digital device use, resulting in 367 tools (222 browser extensions from Chrome Web, 87 apps from Google Play, and 58 apps from the Apple App store).[^3]
 
-[^3]: See Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.
+We adopted Lyngs et al.'s classification of functionality within these tools, which clustered functionality into four types of design patterns for digital self-control:
 
-We adopted Lyngs et al.'s classification of functionality within these tools, which clustered their features into four types of design patterns for digital self-control:
+-   *block/removal* (features for blocking distractions, such as temporarily locking the user out of specific apps, or removing them in the first place, such as by hiding recommended videos on YouTube)
 
--   \emph{block/removal} (features for blocking distractions, such as temporarily locking the user out of specific apps, or for removing them in the first place, such as hiding recommended videos on YouTube)
--   \emph{self-tracking} (features for tracking and/or visualising device use, such as showing how much time the user has spent in particular apps or on their devices overall, or letting the user start a countdown timer when they need to focus)
--   \emph{goal-advancement} (features for reminding the user of their usage goals, such as displaying pop-ups when a specific amount of time has been spent, or replacing the content of new browser tabs with to-do lists)
--   \emph{reward/punishment} (features that add incentives for using devices in particular ways, such as rewarding the user with points/streaks when they do not use their phone, or specific apps)
+-   *self-tracking* (features for tracking and/or visualising device use, such as showing how much time the user has spent in particular apps or on devices overall, or letting the user start a countdown timer when they need to focus)
+
+-   *goal-advancement* (features for reminding the user of their usage goals, such as displaying pop-ups when a specific amount of time has been spent, or replacing the content of new browser tabs with to-do lists)
+
+-   *reward/punishment* (features that provide incentives for using devices in particular ways, such as rewarding the user with points/streaks when they do not use their phone)
 
@@
 
-In March 2019, we scraped additional information about these tools, including installation numbers, ratings, and store descriptions, as well as publicly available user reviews.
-For apps on the Google Play store, we used a pre-existing script [@olanoGoogleplayscraper2018] to scrape tool information and reviews.
-For apps on the Apple App store, we used a pre-existing script [@olanoAppstorescraper2018] to scrape reviews, and used our own script (using the R packages \href{https://rvest.tidyverse.org}{\texttt{rvest}} and \href{http://ropensci.github.io/RSelenium/}{\texttt{RSelenium}}) to scrape tool information.
-As the pre-existing script sometimes failed to retrieve all reviews visible on tools' pages on the Apple App store, we wrote an additional R script to scrape reviews when manual inspection found some to be missing.
+In March 2019, we scraped additional information about these tools, including publicly available user reviews, as well as installation numbers, ratings, and store descriptions.
+For apps on the Google Play store, we used the open-source `google-play-scraper` [@olanoGoogleplayscraper2018] to scrape reviews and tool information.
+For apps on the Apple App store, we used the open-source `app-store-scraper` [@olanoAppstorescraper2018] to scrape reviews, and wrote our own R script (using the packages [`rvest`](https://rvest.tidyverse.org) and [`RSelenium`](http://ropensci.github.io/RSelenium/)) to scrape tool information.
+As the `app-store-scraper` sometimes failed to retrieve all reviews visible on tools' pages on the Apple App store, we wrote an additional R script to scrape reviews when manual inspection found some to be missing.
 For browser extensions on the Chrome Web store, we wrote our own R script to scrape both tool information and user reviews.
@@
 All stores displayed average ratings, as well as a count of how many ratings a tool had received.
-For installation numbers, the stores provided different information: the Chrome Web store provided an exact number of users, and the Google Play store provided a 'minimum number of installs' (e.g.
-'100,000+').
+For installation numbers, the stores provided different information: the Chrome Web store provided an exact number of users, and the Google Play store provided a 'minimum number of installs' (e.g. '100,000+').
 The Apple App store provided no direct information about user numbers.
-For processing which ranked tools by installation numbers, we used number of ratings as a proxy for relative numbers of users for tools on the Apple App store.
+For processing that ranked tools by installation numbers, we used number of ratings as a proxy for relative numbers of users for tools from the Apple App store.
 
-\subsection{Sampling and thematic analysis of user reviews}
+## Sampling and thematic analysis of user reviews
 
-To get a broad coverage of reviews for tools across stores and types of design patterns, we used an R script to systematically sample user reviews for qualitative analysis from the full scraped data.
-To ensure that the reviews we analysed came from tools with a minimum level of engagement from users, we excluded reviews from tools on the Google Play and Chrome Web stores that had fewer than 100 installations/users, and tools from the Apple App Store that had fewer than 5 ratings.
-To make the sampled reviews more likely to be informative, we also excluded reviews shorter than 10 words.
-Our script then iterated through the four types of design patterns described above, on each iteration picking the tool on each store that included that specific type of design pattern and had the highest number of users (Google Play and Chrome Web) or ratings (Apple App store).
-The script then randomly sampled up to 30 user reviews from each of those tools (if a specific tool had fewer than 30 reviews, all of its reviews would be included), and started over.
-The script ran until more than 1,500 reviews were sampled in total.
-A tool could only be sampled from once.
+To get a broad coverage of reviews for tools across stores and types of design patterns, we wrote an R script to systematically sample user reviews for qualitative analysis from the scraped data.
+To ensure that the reviews analysed came from tools with a minimum level of engagement on the stores, we only sampled reviews from tools on the Google Play and Chrome Web stores that had at least 100 installations/users, and tools from the Apple App Store that had at least 5 ratings.
+To make the sampled reviews more likely to be informative, we excluded reviews shorter than 10 words.
+Our script iterated through the four types of design patterns described above, on each iteration picking the tool on each store that included that specific type of design pattern and which had the highest number of users (Google Play and Chrome Web) or ratings (Apple App store).
+The script then randomly sampled 30 user reviews from each of those tools (if a tool had fewer than 30 reviews, all of its reviews would be included), and started over.
+The script ran until at least 1,500 reviews were sampled in total.
+A specific tool could only have its reviews sampled once.
 
 Next, we conducted inductive thematic analysis of the sampled reviews, following the 'reflexive' approach described in [@braunUsingThematicAnalysis2006] and [@braunThematicAnalysis2018].
-Our work was informed by an earlier analysis of 961 reviews randomly sampled from the same data set by the lead author of the present paper [@lyngs2021a, ch. 4].
+Our work was informed by an earlier analysis of 961 reviews randomly sampled from the same data set by the lead author of this paper [@lyngs2021a ch. 4].
 However, our analysis in the present paper was conducted from scratch, with a new sample of user reviews drawn in the manner described.
-First, the lead author (UL) read through all user reviews sampled for thematic analysis and conducted initial coding of recurrent patterns relevant to the research topic.
+After sampling reviews for analysis, the lead author (UL) read through them all and conducted initial coding of recurrent patterns relevant to the research focus.
 A second author (LC) read through one third of the reviews and independently coded recurrent patterns of meaning.
@@
 
-Thematic coding was conducted using [NVivo](https://www.qsrinternational.com/nvivo-qualitative-data-analysis-software/home/) v1.6.1; all quantitative analyses were conducted in R.
+Thematic coding was conducted using [NVivo](https://www.qsrinternational.com/nvivo-qualitative-data-analysis-software/home/) v1.6.1; all quantitative analyses were conducted in R v4.0.5 (see [osf.io/cdqub/](https://osf.io/cdqub/) for information on R package versions used).
 
@@
 
-At the time of scraping, `r num_missing` of the tools from Lyngs et al.'s review had become unavailable.
-Hence, we scraped information for `r nrow(meta_data)` tools: `r num_w_info_by_store %>% filter(store == "chrome") %>% pull(n)` extensions from Chrome Web, `r num_w_info_by_store %>% filter(store == "play") %>% pull(n)` apps from Google Play, and `r num_w_info_by_store %>% filter(store == "apple") %>% pull(n)` apps from the Apple App store.
+At the time of scraping, `r num_missing` tools from Lyngs et al.'s review had become unavailable. 
+Hence, we scraped information for `r nrow(meta_data)` tools: 
+`r num_w_info_by_store %>% filter(store == "chrome") %>% pull(n)` extensions from Chrome Web, `r num_w_info_by_store %>% filter(store == "play") %>% pull(n)` apps from Google Play, and `r num_w_info_by_store %>% filter(store == "apple") %>% pull(n)` apps from the Apple App store.
 
@@
 Font size indicates relative frequency.
-The word cloud excludes the terms 'app/s', 'phone', and 'extension', as well as stop words (i.e., common English words unlikely to be informative such as "the", "of", "to", etc.
+The word cloud excludes the terms 'app/s', 'phone', and 'extension', as well as stop words (i.e., extremely common words unlikely to be informative such as "the", "of", "to", etc.
 Stop words were excluded using the R package `tidytext` [@silge2016tidytext]).
@@
   summarise(num_reviews = n()) %>% 
-  summarise(median = median(num_reviews),
-            mean = mean(num_reviews),
-            min = min(num_reviews),
-            max = max(num_reviews),
-            IQR = IQR(num_reviews)))
+  summarise(mean = mean(num_reviews),
+            min = quantile(num_reviews, 0),
+            q25 = quantile(num_reviews, 0.25),
+            median = quantile(num_reviews, 0.5),
+            q75 = quantile(num_reviews, 0.75),
+            max = quantile(num_reviews, 1)))
 
 (word_count_summary <- reviews_w_main_cat_data %>% 
-  summarise(median = median(review_word_count),
-            mean = mean(review_word_count),
-            min = min(review_word_count),
-            max = max(review_word_count),
-            IQR = IQR(review_word_count)))
+  summarise(mean = mean(review_word_count),
+            min = quantile(review_word_count, 0),
+            q25 = quantile(review_word_count, 0.25),
+            median = quantile(review_word_count, 0.5),
+            q75 = quantile(review_word_count, 0.75),
+            max = quantile(review_word_count, 1)))
 
@@
 
-For the tools that had received reviews, the median number of reviews was `r num_reviews_summary$median` (min = `r num_reviews_summary$min`, max = `r num_reviews_summary$max`, IQR = `r num_reviews_summary$IQR`), and the median number of words was `r word_count_summary$median` (min = `r word_count_summary$min`, max = `r word_count_summary$max`, IQR = `r word_count_summary$IQR`; reviews containing only emojis have a word count of 0).
-Figure \@ref(fig:wordcloud) shows the overall frequency of terms used in the reviews, excluding 'app/s', 'phone' and 'extension' as well as common stop words in English (e.g., 'the', 'of', 'to'; stop words were excluded using the R package `tidytext` [@silge2016tidytext]).
-Aside from 'app' (46% of reviews), the most frequent term was 'time', which was included in 17% of all reviews.
-The terms most commonly following or preceding 'time' included 'limit', 'usage', 'spent', 'screen', 'waste', and 'management', suggesting that managing time spent on digital devices was a central topic in the reviews.
+For the tools that had received reviews, the median number of reviews was `r num_reviews_summary$median` (min = `r num_reviews_summary$min`, max = `r num_reviews_summary$max`, IQR = `r num_reviews_summary$q25` to `r round(num_reviews_summary$q75)`),
+and the median number of words was `r word_count_summary$median` (min = `r word_count_summary$min`, max = `r word_count_summary$max`, IQR = `r word_count_summary$q25` to `r word_count_summary$q75`; reviews containing only emojis have a word count of 0).
+Figure \@ref(fig:wordcloud) shows the overall frequency of terms used in the reviews, excluding the terms 'app/s', 'phone' and 'extension' as well as common stop words ('the', 'of', 'to', etc.; excluded using the R package `tidytext` [@silge2016tidytext]).
+Aside from 'app' (`r ptc_app`% of reviews), the most frequent term was 'time', which was used in `r ptc_time`% of all reviews.
+The terms most commonly following or preceding 'time' included 'limit', 'usage', 'spent', 'screen', 'waste', and 'management', suggesting that managing time spent on digital devices was a central topic (which could reflect users' interests or preferences and/or that currently available tools simply focus on this).
 
@@
 
-(ref:ta-reviews-by-design-patterns) Reviews sampled for thematic analysis (total n = 1,529), by design patterns included in the tools.
-
-```{r sampled-reviews}
+```{r, include=FALSE}
 reviews_for_ta <- read_excel(here::here("data", "processed", "2022-01-22-reviews_for_ta.xlsx"))
@@
   summarise(num_reviews = n())
+```
 
+We conducted thematic analysis of `r nrow(reviews_for_ta)` reviews from a total of `r reviews_for_ta %>% distinct(tool_id) %>% nrow()` different tools (Table \@ref(tab:sampled-reviews) shows how the reviews were distributed across tools with different combinations of design patterns).
+In the following, we accompany review quotes with the name of the tool the review was for, alongside an indication of which types of design patterns the tool included, and which store it came from:
+
+-   \typeBlock\hspace{1mm}= block/removal, \typeTrack\hspace{1mm}= self-tracking,\
+    \typeGoal\hspace{1mm}= goal-advancement, \typeReward\hspace{1mm}= reward/punish
+
+-   C = Chrome Web, P = Google Play, A = Apple App store
+
+-   Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \textbf{C}
+
+(ref:ta-reviews-by-design-patterns) Reviews sampled for thematic analysis (total n = 1,529), by design patterns included in the tools.
+
+```{r sampled-reviews}
 # create a table with the count of tools and review from each combination of functionality
@@
   mutate(`Colour code` = categories_alphabetically %>% str_replace("Block/removal", "\\\\typeBlock") %>% str_replace("Reward/punish", "\\\\typeReward") %>% str_replace("Self-tracking", "\\\\typeTrack") %>% str_replace("Goal-advancement", "\\\\typeGoal") %>% str_remove_all(" \\+ ")) %>%
-  select("Design patterns included" = categories_alphabetically,
+  select("Design patterns included in DSCT" = categories_alphabetically,
          "Colour code",
@@
 
-We conducted thematic analysis of `r nrow(reviews_for_ta)` reviews from a total of `r reviews_for_ta %>% distinct(tool_id) %>% nrow()` different apps (Table \@ref(tab:sampled-reviews) shows how the reviews were distributed across tools with different combinations of design patterns).
-In the following analysis, we accompany review quotes with the name of the tool the review was for, a colour coded indication of which design patterns the tool included, and which store it came from:
-
--   \typeBlock\hspace{1mm}= block/removal, \typeTrack\hspace{1mm}= self-tracking,\
-    \typeGoal\hspace{1mm}= goal-advancement, \typeReward\hspace{1mm}= reward/punish
--   C = Chrome Web, P = Google Play, A = Apple App store
-    -   Example review quote from a Chrome Web tool including design patterns for block/removal, self-tracking, and goal-advancement: *"It would be nice to make it hard to disable, something like a simple maths problem that we could change difficulty, so I am not tempted."*, R1408, Timewarp \typeBlock\typeTrack\typeGoal \\textbf{C}
-
 Our early codes were 'domain summaries' [@braunThematicAnalysis2018] organised in an evolving hierarchy of sub-codes.
-For the first coder (UL), early codes were organised in relation to usage contexts and benefits, user characteristics, key design patterns, and praise or criticism of the developer or app.
-For the second coder (LC), early codes were organised in relation to feature requests, positive feedback, negative feedback, and reasons for use.
+For the first coder (UL), early codes were organised as sub-codes of usage contexts and benefits, user characteristics, key design patterns, and praise or criticism of the developer or app.
+For the second coder (LC), early codes were organised as sub-codes of feature requests, positive feedback, negative feedback, and reasons for use.
 
@@
 Thus, the reviews described a common feeling of lacking control over digital device use, which resulted in using devices in unintended ways, or in using them so much that it interfered with other goals.
-However, digital self-control tools were seen as effectively helping reviewers take back control, save time, and improve their lives (*"I can happily say that this extension has changed my life!!!!
-It hides EVERY ANNOYING YouTube feature"*, R385, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \\textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock\typeGoal \\textbf{C}), and many expressed deep appreciation towards the developers ("Who so ever is working behind this software, my deepest gratitude to them", R271, \href{https://chrome.google.com/webstore/detail/blocksite-block-websites/eiimnmioipafcokbfikbljfdeojpcgbh?hl=en-GB}{\emph{BlockSite}} \typeBlock \\textbf{C}).
-Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \\textbf{C}).
+However, digital self-control tools were seen as effectively helping reviewers take back control, save time, and make their device use fit better into their lives (*"I can happily say that this extension has changed my life!!!!
+It hides EVERY ANNOYING YouTube feature"*, R385, DF Tube \typeBlock \textbf{C}, *"Has saved me countless hours of time-wasting nonsense"*, R1000, NewsFeed Eradicator for Facebook \typeBlock \typeGoal \textbf{C}).
+Many expressed deep appreciation towards the developers as a result ("Who so ever is working behind this software, my deepest gratitude to them", R271, BlockSite \typeBlock \textbf{C}).
+Some reviewers seemed surprised that particular tools had been useful to them (*"I added it more as a joke, but it's really good and helpful for those hours you're supposed to study and don't want to be distracted"*, R675, Go Fucking Work \typeBlock \typeReward \textbf{C}).
 Importantly, the reviews also contained a wealth of information on how specific implementations of design patterns could fail to fulfil their purpose, and how they might be improved.
@@
 theme_two <- "DSCTs should flexibly adapt to personal definitions of 'distraction'"
-theme_three <- "DSCTs should provide incentives that are sufficient to change behaviour without feeling overly intrusive"
+theme_three <- "DSCTs should provide a level of support that is sufficient to change behaviour without feeling too coercive."
 ```
 
-As we iteratively worked with the codes and excerpts, we developed three broad themes capturing digital self-control tools' (DSCTs) use contexts and design challenges, assembled from recurrent patterns in the early codes: (i) `r theme_one`, (ii) `r theme_two`, (iii) `r theme_three`
-
-The NVivo files with our early codes, and their later re-organisation by relation to the three main themes, as well as exports of the code hierarchy to other formats, are available on \href{https://osf.io/cdqub/}{osf.io/cdqub/}.
+As we iteratively worked with the codes and excerpts, we developed three broad themes capturing digital self-control tools' (DSCTs) use contexts and design challenges, assembled from recurrent patterns in the early codes:
+(i) `r theme_one`, (ii) `r theme_two`, (iii) `r theme_three`
+The NVivo files with our early codes, and their later re-organisation by relation to the three main themes, as well as exports of the code hierarchy to other formats, are available on [osf.io/cdqub/](https://osf.io/cdqub/).
 
@@
 
-Users saw DSCTs as critical to avoiding time-wasting patterns of habitual or impulsive digital device use.
-Many reviews described this challenge as related to 'addictive' design patterns and other adversarial dynamics online, to which DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{Inbox When Ready for Gmail} \typeBlock\typeGoal\typeTrack \\textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \\textbf{A}).
-DSCTs seemed especially helpful when it was not possible to avoid using distracting services (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, \href{https://chrome.google.com/webstore/detail/kill-news-feed/hjobfcedfgohjkaieocljfcppjbkglfd}{Kill News Feed} \typeBlock \\textbf{C}).
+A fair number of reviews described reduced time on devices overall or on specific services as the reason why DSCTs were helpful (n = 51, 3.3% of reviews).
+More commonly than limiting time *per se*, however, DSTCs were described as useful because they helped people focus on a specific task amidst easily available distractions on their devices (n = 212, 14%).
 
-The usage contexts mentioned suggested that, aside from wanting to limit time on devices overall or on specific services (n = 62, 4% of reviews), reviewers used DSTCs to help them focus on a given task amidst all too easily available distractions on their devices (n = 201, 13%).
-The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding, expressed in some reviews as DSCTs helping overcome 'procrastination' (*"seeing my short-term to-do list nudges me to get back to work when I am opening a new website to procrastinate"*, R324, Dayboard \typeBlock\typeTrack \\textbf{C}).
-In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \\textbf{C}), and/or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack\typeReward \\textbf{C}).
-Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11%; "Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity", R776, \href{https://chrome.google.com/webstore/detail/inbox-when-ready-for-gmai/cdedhgmbfjhobfnphaoihdfmnjidcpim}{\emph{Inbox When Ready for Gmail}} \typeBlock\typeGoal\typeTrack \\textbf{C}), including studying, writing, or reading (n = 90, 6%).
+Many reviews described their challenges in this respect as caused by time-wasting patterns of habitual or impulsive use, exacerbated by distracting or 'addictive' design patterns and other adversarial dynamics online.
+However, DSCTs provided essential countermeasures that reviewers had searched for (*"Was looking for a solution for this for over 1 year"*, R771, Inbox When Ready for Gmail \typeBlock\typeGoal\typeTrack \textbf{C}; *"Should be on all phones as standard"*, R982, Mute Screen Time Tracker \typeGoal\typeTrack \textbf{A}), and many reviewers noted that they had experimented with different tools (n = 80, 5%).
+DSCTs seemed especially helpful when avoiding use of a distracting service was not an option (*"This helps me avoid the mind numbing news feed that is of Facebook when work has to be done on Facebook itself"*, R873, Kill News Feed \typeBlock \textbf{C}).
 
-Another important use context was \emph{sleep}, where many reviewers said their digital devices often interfered with going to bed at the intended time.
-However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \\textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \\textbf{P}).
+The need for support from DSCTs to stay focused seemed especially acute when tasks were important, but less instantly rewarding than the alternatives.
+Thus, many reviewers generically stated that DSCTs benefited them in productivity/work contexts (n = 167, 11%; *"Not being distracted by unread/misc emails has proven to be of significant benefit to my productivity"*, R776, Inbox When Ready for Gmail \typeBlock \typeGoal \typeTrack \textbf{C}), including studying, writing, or reading (n = 90, 6%).
+Some reviews expressed this as DSCTs helping overcome 'procrastination' (\textit{seeing [my short-term to-do list] nudges me to get back to work when I am opening a new website to procrastinate}, R324, Dayboard \typeBlock \typeTrack \textbf{C}).
 
-**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges** In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely.
-This was often expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (n = 58, 3.8%; *"This was the perfect solution to my phone addiction.
-It was consuming my life and my free time"*, R612, Forest \typeTrack\typeReward \\textbf{P}).
-Sometimes reviewers used the clinical labels ADD or ADHD to explain why tools had been helpful to them (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack\typeReward \\textbf{P}).
+Another important use context was *sleep*, where some reviewers said their digital devices interfered with going to bed at the intended time.
+However, DSCTs helped them adjust their habits in favour of the sleep patterns they wanted (*"I always used to stay up really late on my phone until I got this app - I have consistently been getting 9 hours sleep and waking up early for the first time in my life !!"*, R1252, SleepTown \typeTrack \typeReward \textbf{P}; *"It really forces me to stay off to the phone, and so it's useful especially when I need to go to bed"*, R198, BlackOut \typeBlock \textbf{P}).
+
+In these situations, DSCTs helped reviewers stay in control by reminding them of what they wanted to do, by reducing exposure to 'temptations' (*"Without it, it is too easy for me to be tempted to view and open YouTube sidebar links"*, R376, DF Tube \typeBlock \textbf{C}), or by providing incentives for intended use (*"Not only does it help me stay focused, I smile with delight looking at all my trees and flowers at the end of the day"*, R642, Forest \typeTrack \typeReward \textbf{C}).
+
+**Sub-theme: DSCTs as especially useful for people with more acute self-regulation challenges** In an extension of the main theme, DSCTs seemed especially useful for individuals who experience self-regulation challenges more intensely (n = 53, 3.5%).
+This was commonly expressed as a tool being helpful for tackling 'addiction', and/or as reviewers labelling themselves 'addicts', 'procrastinators' or just low in self-control (*"This was the perfect solution to my phone addiction.
+It was consuming my life and my free time"*, R612, Forest \typeTrack \typeReward \textbf{P}).
+Occasionally, reviews used clinical labels like ADD or ADHD to explain why tools had been helpful (*"As someone with ADHD, the pomodoro technique has been super effective"*, R615, Forest \typeTrack \typeReward \textbf{P}).
 
@@
 
-DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 237, 16%), a theme often expressed in the form of feature suggestions that would make tools more customisable.
+DSCTs needed to accurately capture users' personal definitions of 'distracting use' to be helpful (n = 243, 16%), a theme often expressed in the form of feature suggestions that would make tools more customisable.
 
-One aspect involved getting tools to accurately capture \emph{what} was considered a distraction.
-Thus, for block/removal tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (\emph{"I would like to block reddit but still be able to visit productivity related subreddits"}, R1401, Timewarp \typeBlock\typeTrack\typeGoal \\textbf{C}).
-Similarly, self-tracking tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage""*, R989, Mute Screen Time Tracker \typeGoal\typeTrack \\textbf{A}).
+One aspect involved getting tools to accurately capture *what* was considered a distraction.
+Thus, **self-tracking** tools often received feature requests for the ability to exempt specific functionality from tracking because not all use represented use that reviewers wanted support in managing (*"I use a meditation app and can be meditating several times a day, sometimes up to an hour at a time, so as it doesn't distinguish between that and social media it gives a false impression of"usage" "*, R989, Mute Screen Time Tracker \typeGoal \typeTrack \textbf{A}).
 
-Another aspect involved \emph{when} specific activities should count as distractions.
+Similarly, for **block/removal** tools, reviewers wanted to be able to customise which apps or websites were targeted, or target blocking to specific functionality (*"I would like to block reddit but still be able to visit productivity related subreddits"*, R1401, Timewarp \typeBlock \typeTrack \typeGoal \textbf{C}).
+DSCTs on iOS in particular struggled to provide blocking catered to users' needs, because Apple's restrictive developer permissions do not easily allow apps to block other apps.
+Therefore, iOS tools that provide blocking tended to be limited to blocking websites within the Safari browser.
+Sometimes, users did not realise this limitation at first, and went on to express their frustration in reviews (*"I thought the app would be linked to the apps on my phone I intended to block.
+So disappointed"*, R878, Liberate \typeBlock \textbf{A}).
+A workaround provided by one 'blocking' app, was to generate random passwords that would be shown only once (Self-Control to Focus - Lite \typeBlock ).
+This app instructed users to go to apps they wanted to block, change their account passwords to the randomly generated one without memorising it, then log out out of their accounts (!).
+The app would show users the password again on a set date, until which they would be unable to log back into their accounts.
+
+Another aspect involved *when* specific activities should count as distractions.
 For example, tools for curbing nightly use received feature requests to handle variation in bedtime schedules (*"it needs to allow for the real life fact that people can't go to bed and wake up at the same time everyday.
-If it was done on your amount of sleep you get in hours that would be better"*, R1250, [SleepTown](https://apps.apple.com/gb/app/sleeptown/id1210251567)\typeTrack\typeReward \\textbf{A}).
+If it was done on your amount of sleep you get in hours that would be better"*, R1250, SleepTown \typeTrack \typeReward \textbf{A}).
 
-**Sub-theme: Alignment between users' and tools' framing of distraction as important for navigating negative emotions elicited by DSCTs** Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor was the negative emotions that interventions could elicit.
-Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and essential, users also needed to carefully navigate the emotional reactions they could create (*"I hate it and love it at the same time.* \emojiTearsOfJoy", R1299, Stay Focused \typeBlock\typeTrack \\textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock\typeTrack \\textbf{P}).
+**Sub-theme: Alignment between user and tool as important for navigating negative emotions** Aside from practical reasons that DSCTs should accurately capture behaviour that reviewers wished to change, an important factor seemed to be the negative emotions that interventions could elicit (n = 30, 2%).
+Thus, whereas the various restrictions and reminders provided by DSCTs were seen as helpful and important, users also needed to carefully navigate the emotional reactions they could create (\textit{"I hate it and love it at the same time.
+\emojiTearsOfJoy"}, R1299, Stay Focused \typeBlock \typeTrack \textbf{P}, *"Pisses me off sometimes, but worth it"*, R357, Detox Procrastination Blocker, \typeBlock \typeTrack \textbf{P}).
+
 In order for experiences of "shock", "pressure", etc., to be useful, it was important that the user felt that a DSCT's purpose and framing of usage were closely aligned with their personal perspective.
-
 Thus, tools that provided standards for 'good' or 'bad' usage, for example in their presentation of usage statistics or reminder pop-ups, needed to tread carefully.
-Being reminded that one did not behave in line with one's intentions could be very helpful (*"This tells it like I need to be told sometimes...LOVE IT!
-Simple and effective :)"*, R690, \href{https://chrome.google.com/webstore/detail/go-fucking-work/hibmkkpfegfiinilnlabbfnjcopdiiig}{Go Fucking Work} \typeBlock\typeReward \\textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock\typeGoal \\textbf{C}).
-However, normative messaging could also backfire and be profoundly annoying if reviewers did not agree with it (*"I was a bit annoyed when Big Brother started getting judgemental on me.
+Being reminded that one did not behave in line with one's intentions could be very helpful (*"This tells it like I need to be told sometimes\...LOVE IT!
+Simple and effective :)"*, R690, Go Fucking Work \typeBlock \typeReward \textbf{C}; *"I like the guilt trip it gives me if I attempt to get into facebook or other social networks"*, R311, Dayboard, \typeBlock \typeGoal \textbf{C}).
+However, normative messaging could backfire and be profoundly annoying if reviewers did not agree with it (*"I was a bit annoyed when Big Brother started getting judgemental on me.
 "You have hit your target of 50 pickups!" "You've now picked up your phone 90 times!" Well, first I didn't set a target, and second, I don't think that 54 minutes across a day should make me feel guilty.
-Especially since I was reading a book during that time"* R911, Moment \typeGoal\typeTrack \\textbf{A}).
+Especially since I was reading a book during that time"* R911, Moment \typeGoal \typeTrack \textbf{A}).
 
@@
 
-Reviewers searched for DSCTs that provided interventions whose incentives were 'just right' to bring about an intended change in behaviour without feeling overly intrusive.
+Reviewers searched for DSCTs that provided support which was 'just right' to bring about an intended change in behaviour without feeling too coercive (n = 271, 18%).
 The reviews implied a spectrum of individual differences in where this 'Goldilocks level' was.
-A few reviewers stated that simply receiving reminders or knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, \emph{Mute Screen Time Tracker} \typeGoal\typeTrack \\textbf{A}).
-More commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9%).
+A few reviewers stated that simply knowing that a tracking tool was monitoring them was sufficient (*"I know it's tracking me so I pick up the phone a lot less"*, R992, *Mute Screen Time Tracker* \typeGoal \typeTrack \textbf{A}).
+Commonly, however, reviewers wanted DSCTs to provide more strict or persuasive ways to hold themselves accountable, so that they did not succumb to temptations or unwanted habits in moments of weakness or absentmindedness (n = 132, 9%).
+The specifics differed for different design patterns.
 
-Thus, tools implementing block/removal design patterns were often criticised if they were too easy to disable (*"This is fantastic but the thing is I can simply remove the extension from chrome which defeats the purpose.
-Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \\textbf{C}).
-However, if they were *too* difficult to override, they ran into another problem, namely how to deal with emergencies or otherwise unforeseen circumstances.
-One developer followed reviewers' suggestion to add a 'pause' button, only to receive critique that blocking was now too easy to override.
-Reviewers suggested that adding varying forms of friction to turn off blocking would be a solution, such as having to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock\typeTrack \\textbf{P}), or simply adding passwords, which could be set by a trusted other.
+For patterns related to **distraction blocking/removal**, it was important that blocking was not too easy to disable in moments of temptation (*"This is fantastic but the thing is I can simply remove the extension from chrome which defeats the purpose.
+Does anyone know how I can stop myself from removing the extension?"*, R1217, Simple Blocker \typeBlock \textbf{C}).
+On the other hand, if blocking was *too* difficult to disable, it would be criticised for failing to account for emergencies or otherwise unforeseen circumstances.
+One developer followed reviewers' suggestion to add a 'pause' button, only to then receive critique that blocking was now too easy to turn off.
+Reviewers suggested that adding friction to turn off blocking would provide a solution, such as being required to complete a time-consuming or effortful task (*"make it necessary to touch the screen for a minute or 2 so that it becomes harder for anyone to stop it"*, R8, ( OffTime ) \typeBlock \typeTrack \textbf{P}), or type in a password that could be set by a trusted other.
 
-Similarly, for tools implementing reward/punishment, some reviewers suggested increasing incentives to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack\typeReward \\textbf{C}).
-Here too, however, tools needed to find the right balance, because punishments that were perceived as too harsh might make users feel like failures (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed'"*, R1247, \href{https://itunes.apple.com/gb/app/sleeptown/id1210251567?mt=8\&uo=4}{SleepTown} \typeTrack\typeReward \\textbf{P}).
+Another aspect was that the amount of effort required to *initiate* use of blocking tools should be low, because users' initial motivation might be lacking (*"Most users have procrastination issues which make them postpone the moment in which they will click on the "Start focusing" button"*, R545, Focus \typeBlock \typeGoal \typeTrack \textbf{C}).
+Thus, many feature requests asked for the ability to set a schedule where blocking would turn on automatically (\textit{``can I request an automatic start time option?
+So every M-F, 9-5 the app starts working automatically?
+I need it to cut me off from whatever dumb thing I'm doing before work starts :)'', R238, Block \& Focus \typeBlock\typeTrack \textbf{C}}).
+An interesting suggestion for tools alternating 'focus' and 'break' sessions, was to flip the order to begin with a break, followed by enforced blocking.
+This might lower the amount of motivation needed to initiate a focus session (*"Hitting the start button the very first time is hard!
+Starting in a short break would allow me to hit this without a second thought, because I have a few short moments to surf before the focus is forced."*, R232, Block & Focus \typeBlock \typeTrack \textbf{C}).
 
-For tools including goal advancement design patterns, the 'right' level of incentive often related to the frequency or framing of features such as reminder pop-ups, which needed to be finely calibrated to be helpful rather than annoying (*"the todo popup on each page load is quite useful (...) it should be limited (across multiple tabs) to only pop-up every minute or so (...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock\typeGoal \\textbf{C}).
+For **reward/punishment** design patterns, reviewers sometimes suggested ways to increase the incentives provided to make tools more effective (*"If I could also name the trees (so I'll feel even worse about letting them die) it would be even better!"*, R650, Forest \typeTrack \typeReward \textbf{C}).
+Here too, however, tools needed to find the right balance.
+If a punishment was perceived as *too* harsh, it might make users feel dejected (*"I hate the fact that we get a destroyed building (...) I feel like it is too punishing and almost says 'you've failed' "*, R1247, SleepTown \typeTrack \typeReward \textbf{P}).
 
-**Sub-theme 1: Combining types of design patterns to find the Goldilocks level of incentive** In addition to tweaking parameters of a tool's current design patterns, many reviewers suggested including other types of patterns to make tools more effective.
-Thus, some reviewers for self-tracking tools said that simply being informed about their behaviour was not sufficient for them to change anything.
-However, incorporating other types of elements might make it more effective, such as combing self-tracking with goal-setting.
+For **goal advancement** patterns, the 'just right' level of support mainly related to the frequency with which users could be reminded of their goals: too infrequent, and users might fail to complete their tasks; too frequent, and users would get annoyed.
+For example, one browser extension replaced the content of new tabs with to-do lists, and also provided task reminders via pop-ups.
+This tool was on the one hand criticised for dropping unfinished tasks from the day before, as users would then forget those tasks.
+On the other hand, its pop-up reminders were criticised for being too frequent (*"the todo popup on each page load is quite useful (\...) it should be limited (across multiple tabs) to only pop-up every minute or so (\...) I find myself snoozing it too often just because the popup bothers me"*, R339, Dayboard, \typeBlock \typeGoal \textbf{C}).
 
-Similarly, some reviewers wanted block/removal tools to not simply block distractions, but also actively nudge their behaviour in ways that supported intended habits.
-This might happen by, for example, forwarding blocked websites to productive alternatives, or by replacing algorithmic recommendations with content that they wanted to engage with more ("\emph{if, where the home-screen recommended section is, this could be replaced by a website of our choice} (\ldots) \emph{This way I could do music theory, or learn German.
+For **self-tracking** patterns, reviewers commonly said that DSCTs implementing this helped them understand and assess their usage, which was often described as the first step in changing behaviour.
+However, many reviews said that it was not in itself sufficient support to change, and that tracking needed to be combined with other strategies to be effective (*"It does not stop me spending 10 hours on Facebook but it does inform me (\...) What has worked great is using rescue time with very intentional time blocking"*, R1157, RescueTime, \typeTrack \typeGoal \typeBlock ).
+We elaborate on this in the sub-theme below.
+
+Similarly to the challenges around blocking, self-tracking tools on iOS faced an additional challenge related to the amount of user effort required to track use in the first place.
+This occurred because Apple's developer permissions do not allow apps to easily track time spent in other apps.
+As a work-around, some tools required users to take screenshots of their phone's battery page, from which information about time spent in different apps could be extracted.
+This work-around was, however, seen as too effort-demanding by reviewers (*"I should have looked at how you use this.
+Take a picture everyday?????
+Really?"*, R121, App Usage Tracker \typeTrack \textbf{A}).
+
+**Sub-theme: Combining design patterns to find the Goldilocks level of support** Rather than simply focusing on the parameters of individual design patterns, some reviews focused on how a *combination* of different types of patterns could make DSCTs provide the right level of support for behaviour change.
+For example, some reviews for self-tracking tools suggested that adding a goal-setting or blocking element might make it more effective.
+Similarly, some reviewers wanted block/removal tools to not simply block distractions, but also actively nudge them in ways that supported intended habits.
+This might happen by, for example, redirecting blocked websites to productive alternatives, or replacing recommender feeds with content people wanted to engage with more ("*if, where the home-screen recommended section is, this could be replaced by a website of our choice* (...) *This way I could do music theory, or learn German.
 Things I always intend to do, but don't do enough.
-But if these things came up every time I log into Youtube I'd definitely use them."}, R372, \href{https://chrome.google.com/webstore/detail/df-tube-distraction-free/mjdepdfccjgcndkmemponafgioodelna?hl=en}{DF Tube} \typeBlock \\textbf{C}).
-
-**Sub-theme 2: DSCTs should have low friction to start and use, but high friction to stop** Finally, finding the right level of friction also related to the level of effort required to use a tool.
-The amount of effort needed to \emph{initiate} use was important to keep low, because a lack of motivation could easily lead users to not use an otherwise effective tool (*"Most users have procrastination issues which make them postpone the moment in which they will click on the "Start focusing" button.
-Therefore those users (including me) will get distracted and they will click on that button after several hours, or even never, sometimes wasting a whole working day"*, R545, Focus \typeBlock\typeGoal\typeTrack \\textbf{C}).
-Therefore, a common feature request for blocking tools was the ability to set a schedule where blocking would automatically turn on.
-An interesting suggestion related to blocking tools alternating between work and break cycles: these could flip the ordering of cycles so as to begin with a break, followed by enforced blocking, which might lower the amount of motivation needed to initiate a focus session (*"Hitting the start button the very first time is hard!
-Starting in a short break would allow me to hit this without a second thought, because I have a few short moments to surf before the focus is forced."*, R232, Block & Focus \typeBlock\typeTrack \\textbf{C}).
-
-The amount of effort involved in routine use of a tool was also important to keep low, because tools that were confusing or demanding were likely to be abandoned.
-This seemed to present an especially difficult challenge for self-tracking tools on iOS, because Apple's restrictive developer permissions do not allow apps to easily track time spent in other apps.
-As a work-around, some self-tracking tools for iOS required users to take screenshots of their phone's battery page, from which the tool was able to extract information about time spent in different apps.
-However, this work-around was simply seen as too effort-demanding by many reviewers (\emph{"I should have looked at how you use this.
-Take a picture everyday?????
-Really?"}, R121, App Usage Tracker \typeTrack \\textbf{A}).
-
-Thus, whereas high friction related to \emph{turning off} a tool could be actively helpful when it served the purpose of holding the user accountable to their own intentions, tools were criticised if they required too much effort to operate.
+But if these things came up every time I log into Youtube I'd definitely use them."*, R372, DF Tube \typeBlock \textbf{C}).
+Similarly, some reviews suggested that a particular tool's combination of design patterns was what made it more effective than alternative tools (*"This is the first todo list that actually keeps me on task.
+Normally I go down the rabbit burrow when working and either get way too focused on one task or completely go off track.
+This tool keeps pulling my attention back to the task list where I can make sure I'm not off track."*, R550, Focus \typeGoal\typeBlock\typeTrack \footnote{Focus replaces new tab content with a to-do list and activity log, alongside a 'start focusing' button.
+When the button is clicked, a 25 min work session is started where websites on a blacklist are blocked.
+If the user tries to access a blocked website, a countdown timer is shown.
+At the time of writing, this extension is no longer available on the Chrome Web Store.}).
 
@@
 
-```{r tool-num-users-and-ratings-plot, fig.cap="Distributions of installation numbers and number of ratings on the Google Play, Chrome Web, and Apple App stores (tools with zero ratings scored as 0.5 ratings, to show them in this plot with a log scale).
+```{r tool-num-users-and-ratings-plot, fig.cap="Distributions of installation numbers and number of ratings for DSCTs on the Google Play (n=71), Chrome Web (n=212), and Apple App (n=51) stores.
+Values are shown on a log scale --- tools with zero ratings are scored as having 0.5 ratings, so that they show in the plot.
 The Apple App store provided no direct information about installation numbers.", out.width='100%'}
@@
     geom_boxplot(outlier.shape = NA, aes(fill = store), position = position_dodge2(preserve = "single"), alpha = 0.3) +
-    scale_y_log10(labels = label_number(big.mark = ",")) +
+    scale_y_log10(labels=comma_format(accuracy=1), breaks = c(1, 10, 1000, 100000, 10000000)) +
     geom_jitter(aes(color = store), width = 0.2, height = 0, alpha = 0.5) +
@@
 
-#ggsave("tool-num-users-and-ratings-plot.pdf", width = 7, height = 5)
+#ggsave("figures/tool-num-users-and-ratings-plot.pdf", width = 7, height = 5)
 ```
 
-```{r, users-and-category, fig.cap="Top tools in terms of installation numbers (Chrome Web and Play) or number of ratings (Apple App store).
-Only 13 of the Apple App store tools had received any ratings.
-Differences between the stores in terms of design patterns are readily apparent: Block/removal patterns are especially common on the Chrome Web store (87% of all tools), where browser extensions can easily modify distracting elements on web pages.
-Reward/punishment patterns are included in 43% of tools on the Apple App store, compared to, respectively, 6 and 15% on the Chrome Web and Play stores.", out.width="70%"}
+(ref:top-tools-caption) Top tools in terms of installation numbers (Chrome Web and Play) or number of ratings (Apple App store).
+Only 13 of the Apple App store tools had received ratings.
+Differences between the stores in terms of design patterns are readily apparent: Block/removal patterns are particularly common on the Chrome Web store (87% of all tools), where browser extensions can easily modify distracting elements on web pages.
+Reward/punishment patterns are included in 43% of tools on the Apple App store, compared to, respectively, 6 and 15% on the Chrome Web and Play stores.
+
+```{r users-and-category, fig.cap="(ref:top-tools-caption)", out.width="70%", fig.align='center'}
 knitr::include_graphics("figures/user_numbers_combined_plot.png")
@@
   filter(store == "play") %>% 
-  summarise(median_min_installs = median(min_installs),
-            min_min_installs = min(min_installs),
-            max_min_installs = max(min_installs),
-            IQR = IQR(min_installs)))
+  summarise(min_min_installs = quantile(min_installs, 0),
+            q25_min_installs = quantile(min_installs, 0.25),
+            median_min_installs = quantile(min_installs, 0.5),
+            q75_min_installs = quantile(min_installs, 0.75),
+            max_min_installs = max(min_installs)))
 
@@
   filter(store == "chrome") %>% 
-  summarise(median_installs = median(num_users),
-            min_installs = min(num_users),
-            max_installs = max(num_users),
-            IQR = IQR(num_users)))
+  summarise(min_installs = quantile(num_users, 0),
+            q25_installs = quantile(num_users, 0.25),
+            median_installs = median(num_users),
+            q75_installs = quantile(num_users, 0.75),
+            max_installs = max(num_users)))
+
+(num_ratings_apple <- meta_data %>% 
+  filter(store == "apple") %>% 
+  summarise(min = quantile(num_ratings, 0),
+            q25 = quantile(num_ratings, 0.25),
+            median = quantile(num_ratings, 0.5),
+            q75 = quantile(num_ratings, 0.75),
+            max = max(num_ratings, na.rm = TRUE)))
 
@@
   filter(num_ratings >= 1) %>% 
-  summarise(median_rating = median(average_rating, na.rm = TRUE),
-            min_rating = min(average_rating, na.rm = TRUE),
-            max_rating = max(average_rating, na.rm = TRUE),
-            IQR = IQR(average_rating, na.rm = TRUE)))
+  summarise(min_rating = quantile(average_rating, 0),
+            q25_rating = quantile(average_rating, 0.25),
+            median_rating = quantile(average_rating, 0.5),
+            q75_rating = quantile(average_rating, 0.75),
+            max_rating = max(average_rating, na.rm = TRUE)))
 
@@
 The distribution of numbers of users/installations and ratings are shown in Figure \ref{fig:tool-num-users-and-ratings-plot}.
-For Google Play tools, the median 'minimum number of installs' was `r format(play_user_stats$median_min_installs, big.mark = ",")` (min = `r format(play_user_stats$min_min_installs, big.mark = ",")`, max = `r format(play_user_stats$max_min_installs, big.mark = ",")`, IQR = `r format(play_user_stats$IQR, big.mark=",")`).
-For Chrome Web tools, the median number of users was `r format(chrome_user_stats$median_min_installs, big.mark=",")` (min = `r format(chrome_user_stats$min_min_installs, big.mark=",")`, max = `r format(chrome_user_stats$max_min_installs, big.mark=",")`, IQR = `r format(chrome_user_stats$IQR, big.mark=",")`).
-(The Apple App store did not display information about number of users.) In terms of the ratings given to tools, the median average rating was `r ratings_overall$median_rating` (min = `r ratings_overall$min_rating`, max = `r ratings_overall$max_rating`, IQR = `r round(ratings_overall$IQR, 1)`; ratings given on a scale from 1 to 5 stars on all stores).
+For Google Play tools, the median 'minimum number of installs' was `r format(play_user_stats$median_min_installs, big.mark = ",")` (min = `r format(play_user_stats$min_min_installs, big.mark = ",")`, max = `r format(play_user_stats$max_min_installs, big.mark = ",")`, IQR = `r format(play_user_stats$q25_min_installs, big.mark = ",")` to `r format(play_user_stats$q75_min_installs, big.mark = ",")`).
+For Chrome Web tools, the median number of users was `r format(chrome_user_stats$media_installs, big.mark=",")` (min = `r format(chrome_user_stats$min_installs, big.mark=",")`, max = `r format(chrome_user_stats$max_installs, big.mark=",")`, IQR = `r format(chrome_user_stats$q25_installs, big.mark=",")` to `r format(chrome_user_stats$q75_installs, big.mark=",")`).
+The Apple App store did not display information about number of users (the median number of ratings for DSCTs from the Apple App store was `r format(num_ratings_apple$median, big.mark=",")`; max = `r format(num_ratings_apple$max, big.mark=",")`, IQR = `r format(num_ratings_apple$q25, big.mark=",")` to `r format(num_ratings_apple$q75, big.mark=",")`).
+The median average rating across all tools was `r format(ratings_overall$median_rating, big.mark=",")` (min = `r format(ratings_overall$min_rating, big.mark=",")`, max = `r format(ratings_overall$max_rating, big.mark=",")`, IQR = `r format(ratings_overall$q25_rating, big.mark=",")` to `r format(ratings_overall$q75_rating, big.mark=",")`; ratings given on a scale from 1 to 5 stars on all stores).
 
-Figure \ref{fig:users-and-category} shows the top tools, ranked by number of users (Chrome Web and Play) or number of ratings (Apple App store), alongside the design patterns they include.
-The distributions of design patterns included in top tools differed by store, mirroring differences between the stores overall: for example, block/removal functionality was rare on the Apple App store compared to the Google Play and Chrome Web store [@lyngsSelfControlCyberspaceApplying2019].
+(ref:caption-ratings-by-functionality) Average ratings by design patterns in digital self-control tools (for tools with 15 or more ratings).
+Tools including more than one type of design pattern had higher average ratings (median = 4.39) than tools implementing a single type (median = 4.2, *p* = 0.003, Wilcoxon's rank-sum test).
+
+```{r, average-ratings-by-functionality, fig.cap="(ref:caption-ratings-by-functionality)", out.width="100%"}
+knitr::include_graphics("figures/ratings_by_type_annotated.pdf")
+```
 
@@
 
-Across all tools, 64% included a single type of design pattern for digital self-control, whereas 32% combined two - very few combined three (3%) or all four (1%).
-Tools that combined two or more types had significantly higher average ratings than tools which implemented a single type (median average rating with a single type = 4.2, median for tools combining two or more = 4.39, *p* = 0.003 in Wilcoxon's rank-sum test; analysis excludes tools with fewer than 15 ratings; see Figure \ref{fig:average-ratings-by-functionality}).
+Figure \@ref(fig:users-and-category) shows the top tools, ranked by number of users (Chrome Web and Play) or number of ratings (Apple App store), alongside the design patterns they include. 
+Across all tools, 64% included a single type of design pattern for digital self-control, whereas 32% combined two.
+Very few combined three (3%, n = 11) or all four (1%, n = 2).
+Tools that combined two or more types had slightly higher average ratings than tools which implemented a single type (median average rating with a single type = 4.2, median for tools combining two or more = 4.39, *p* = 0.003 in Wilcoxon's rank-sum test; excluding tools with fewer than 15 ratings; see Figure \@ref(fig:average-ratings-by-functionality)).
 
@@
 
-(ref:caption-ratings-by-functionality) Average ratings by design patterns in digital self-control tools (for tools with 15 or more ratings).
-Tools including more than one type of design pattern had higher average ratings (median = 4.39) than tools implementing a single type (median = 4.2, *p* = 0.003, Wilcoxon's rank-sum test).
+# Discussion
 
-```{r, average-ratings-by-functionality, fig.cap="(ref:caption-ratings-by-functionality)", out.width="100%"}
-knitr::include_graphics("figures/ratings_by_type_annotated.pdf")
+To sum up, reviewers described DSCTs as effective means of taking back control and making device use fit better with their lives.
+A fair number of reviews suggested that limiting time spent on devices, or on specific apps or services, was an important purpose for using DSCTs.
+More commonly than limiting time *per se*, however (observed four times as frequently in our thematic analysis), DSCTs were specifically used for focusing on less instantly rewarding tasks amidst easily available digital distractions.
+Specific contexts of use included productivity (including work, studying, reading, and writing), as well as bedtime habits.
+DSCTs seemed to be especially useful for individuals with higher baseline levels of self-regulation difficulties, as suggested by reviews who framed usefulness in terms of coping with 'addiction' or conditions like ADHD.
+
+Reviewers wanted DSCTs to provide interventions that accurately targeted specific patterns of use.
+This was often expressed in the form of feature requests for, for example, being able to target blocking on a certain functionality on a website, or exempt specific apps from usage tracking.
+A contributing factor in this respect was that DSCTs' interventions could cause negative emotional reactions in the moment of temptation.
+If user and tool were not in agreement on their definitions of 'distracting use', interventions might simply be considered too annoying.
+
+Reviewers also wanted DSCTs to provide a level of support that was sufficient for effecting their intended change in behaviour, without feeling too coercive.
+Reviewers differed in the level of support they wanted, but commonly looked for interventions strict or rewarding enough to stay on track when they felt tempted to indulge in digital distractions or drift into unwanted habitual use.
+DSCTs needed to strike a delicate balance in this respect, because overly strong incentives might backfire if they made users feel dejected when they failed, or if they did not allow for emergency use or shifting daily schedules.
+
+Some reviewers suggested that combining different types of design patterns was an effective way to provide a level of support that was 'just right'.
+This matched the observation that DSCTs which combined two or more types of design patterns received slightly higher overall ratings than tools including only one.
+A final consideration related to how much effort a tool required to use: whereas high effort to *stop* a tool could help hold users accountable to their goals, the level of effort required to *start* use should be low, to handle potential lack of motivation to initiate blocking on part of the user.
+
+In the following, we discuss implications for design and opportunities for future research.
+
+## Goldilocks principles for DSCTs
+
+Across different design dimensions, users' desired level of support from DSCTs reflected a Goldilocks principle: 'too little' and a DSCT would fail to bring about an intended change in behaviour; 'too much' and it could make the user feel frustrated, annoyed, or dejected.
+The considerations for a 'just right' level of support varied for different design dimensions: For example, for distraction blocking/removal the main concern was *how much effort* the user needed to apply to circumvent the blocking.
+For goal-advancement, it was *how frequently* users were reminded of their goals.
+In designing for a Goldilocks level of support, developers therefore need to consider more than a single dimension of engineered 'friction' [@coxDesignFrictionsMindful2016; @gouldSpecialInterestGroup2021].
+How might we conceptualise Goldilocks principles for different design dimensions and assess the level of support that a DSCT provides?
+
+### Using a dual-systems model to frame the Goldilocks level of support
+
+The dual-systems model commonly used in psychology and HCI provides insight into how DSCTs support or disrupt target behaviors through multiple cognitive pathways.
+Lyngs et al.
+[@lyngsSelfControlCyberspaceApplying2019] originally used this model (Figure \@ref(fig:simple-cog-model)) to analyse how the four main types of design patterns in DSCTs support self-regulation.
+We suggest that it is also useful for understanding Goldilocks principles for those design patterns.
+
+(ref:simple-cog-caption) A simplified dual-systems model of self-regulation, based on [@lyngsSelfControlCyberspaceApplying2019; @Shenhav2017; @sheaSuprapersonalCognitiveControl2014; @normanAttentionActionWilled1986].
+Behaviour can be activated via quick, automatic reactions to stimuli (System 1), as well as via slower, controlled reactions to conscious goals and standards (System 2).
+The strength of System 2 control is mediated by the expected value of control.
+DSCTs aim to influence the competition between System 1 and System 2 responses such that the ‘winning’ behaviour is more likely to be aligned with the user's longer-term goals (e.g., focusing on a work task, or simply reducing time spent on a device or service).
+
+```{r simple-cog-model, out.width='100%', fig.cap="(ref:simple-cog-caption)"}
+knitr::include_graphics("figures/simple_cog_model.png")
 ```
 
-# Discussion
+According to the dual-systems model, behaviour can be activated via quick, automatic reactions to stimuli in our external and internal environment ("System 1", e.g.
+taking out your phone as a habitual response to boredom).
+They can also be activated via slower, controlled reactions to our conscious goals and standards ("System 2", e.g.
+deliberately taking out your phone to text a friend).
+Potential behaviours compete to get expressed, and what wins out is the action that receives the greatest total activation.
+The purpose of DSCTs is to influence the competition between System 1 and System 2 responses such that the 'winning' behaviour is more likely to be aligned with our longer-term goals (e.g., focusing on a work task, or reducing time spent on a device or service).
+Different design patterns achieve this in different ways.
+Patterns for **blocking/removing distractions** aim to stop unwanted actions from being triggered or expressed via System 1 control, and to decrease the value of those actions when considered by System 2; **self-tracking** aims to enable conscious System 2 monitoring of how device use compares to the user's goals and standards; **goal-advancement** aims to bring specific usage goals into consciousness awareness (e.g.
+via task reminders), thereby enabling System 2 to initiate certain actions or inhibit others; and **reward/punishment** aims to make it more likely that System 2 control succeeds, by increasing or decreasing the expected value of target actions [@lyngsSelfControlCyberspaceApplying2019; @Shenhav2017].
 
-To sum up, the user reviews described DSCTs as crucial support for focusing on less instantly rewarding tasks amidst easily available digital distractions.
-Specific usage contexts related mainly to productivity/work (including studying, reading, and writing), as well as sleeping.
-Individuals with greater self-regulation difficulties seemed to find DSCTs particularly useful, as evidenced by reviewers who explained usefulness in terms of coping with 'addiction' or conditions like ADHD.
+What insights can we draw about Goldilocks principles for design patterns in DSCTs from our available data and the dual systems model?
+In the following, we illustrate this in relation to distraction blocking/removal, and how design patterns may be combined.
 
-Reviewers wanted DSCTs to provide interventions that accurately targeted usage they wanted to better control, often expressed as feature requests for, for example, being able to target blocking to specific functionality on a website, or exempt specific apps from usage tracking.
-Close alignment between user and tool in their definitions of 'distracting use' was important, partly to reduce negative emotional reactions to specific interventions, which might otherwise be considered too annoying.
+### Blocking/removal of distractions: effortful to override, effortless to initiate
 
-Reviewers also wanted DSCTs to provide a level of incentive that was sufficient for affecting intended behaviour change without feeling too intrusive.
-Reviewers differed in the level of incentive they wanted, but were commonly looking for interventions strict or rewarding enough to stay on track when they felt tempted to indulge in digital distractions or drift into unwanted habitual use.
-DSCTs needed to find a delicate balance, as overly strong incentives might backfire if they made users feel dejected when they failed to live up to their intentions, or if users had no ways to handle emergency use or shifting daily schedules.
+From the dual systems perspective, 'too little' support from a blocking tool means that the effort required to *override blocking* is too low to stop unwanted System 1 habits from being expressed, and to impose costs on unwanted actions that are large enough for System 2 to favour better options (recall the reviewer who was unable to stop themselves from deactivating blocking).
+Conversely, 'too much' effort occurs if blocking interferes with important, but unexpected usage goals.
 
-Some reviewers suggested that different types of design patterns should be combined to provide the right level of incentive.
-This matched our observation in the popularity metrics that DSCTs which combined two or more types of design patterns received higher overall ratings than tools including only one.
-A final consideration related to how much effort a tool required to use: Whereas a high level of friction to *stop* a tool could help hold users accountable to their goals, the level of friction required to *start* use should be low, to handle initial lack of motivation on the part of the user.
+In providing the 'just right' level, the data suggested the following guidelines (which converged with other studies, [@kimGoalKeeperExploringInteraction2019; @kimLocknTypeLockoutTask2019; @zhangMonitoringScreenTime2022; @tseng2019; @kimExploringContextawareProactive2018]):
 
-In the following, we discuss implications for future research.
+1.
+By default, require the user to invest moderate effort to override blocking (e.g., having to manually type in a long string of random letters), in order to enforce commitment while still allowing emergency use
 
-\subsection{Using self-regulation measures to evaluate DSCTs and assess individual user needs}
+2.  Make the level of required effort adjustable by the user, to account for individual differences
 
-Our work provides additional data points for the discussion around how to evaluate effectiveness of DSCTs, and understand how needs for support vary between users.
-In terms of behavioural measures, our thematic analysis and the term frequencies in user reviews overall suggested that limiting time spent on devices, or in specific apps and services, was an important concern for many.
-Hence, whereas simple measurements of 'screen time' may have many inadequacies in assessing DSCT effectiveness [@cecchinatoDesigningDigitalWellbeing2019; @hinikerMovingGreatScreen2019; @lukoffDigitalWellbeingWay2019], they remain important.
+3.  Provide ways to initiate blocking with little effort, to counter instances of low motivation (e.g., via automatic blocking according to a schedule)
 
-More commonly than limiting time, however (observed three times as frequently in our thematic analysis), the reason for using DSCTs was specifically to support focus on less instantly rewarding tasks in the face of easily available digital distractions.
-This matches findings from research using other methods [@reineckeSlackingWindingExperience2016], such as Tran et al.
-[@tranModelingEngagementDisengagementCycle2019]'s interview study showing that tedious or effortful activities are common triggers of compulsive smartphone use, and Kim et al.'s finding that their phone-restriction app was found particularly useful in "boring classes" [@kimLetFOCUSMitigating2017].
-What are the most useful ways for us to assess DSCTs' effectiveness for this purpose?
+Thus, as the reviews suggested, a viable solution is to require users to complete an effortful task to override blocking [@kimLocknTypeLockoutTask2019; @zhangMonitoringScreenTime2022].
+It is an open question how much the Goldilocks amount of effort varies between individuals and with emotional state [@kimLocknTypeLockoutTask2019].
+However, a simple design solution found in commercially available tools is to allow users to choose the level of enforcement they find appropriate.
+For example, Cold Turkey Blocker (<https://getcoldturkey.com>) provides a choice between no enforcement --- tool can be turned off at will --- having to type in a password or string of random letters, having to restart one's computer, or allowing blocking to be modified only at certain times.
 
-In a given context, simple behavioural measures may provide useful proxies, such as error rates or output in a work task [@borghoutsLookingInformationEmail2018; @marottaOnlineDistractionsWebsite2017] or time spent in functionality 'relevant' or 'distracting' relative to a goal [@lottridgeBrowserDesignImpacts2012; @hinikerMyTimeDesigningEvaluating2016; @whittakerDonWasteMy2016; @kimTechnologySupportedBehavior2017].
-However, we also need effective ways to assess users' own experience.
+*Initiating blocking* comes at the short-term cost of excluding ourselves from potentially rewarding actions.
+This puts System 2, which compares the expected value of different actions [@Shenhav2017], in a conflicted position.
+At times, manually initiating blocking is therefore 'too high' effort, which reviewers expressed as procrastinating on starting a block session.
+As a solution, blocking could be automatically applied [@tseng2019; @kimExploringContextawareProactive2018], or the perceived short-term cost of blocking lowered by e.g.
+adding a time lag between initiation and application of blocking.
+Though not directly expressed in our data, initiating distraction blocking with 'too little' effort may conversely increase the risk that users apply blocking at the wrong time, or to the wrong content.
+This risk might be acceptable, however, if blocking is not impossible, but only effortful to override.
 
-Some existing work, especially on 'problematic use', has used self-report scales related to behavioural addiction to evaluate DSCTs and/or assess baseline difficulties with managing device use [@leeSAMSSmartphoneAddiction2014; @koLockLoLGroupbased2016].
-Many researchers have pushed back against this approach, because it risks pathologising everyday patterns of use.
-Thus, it has been argued that even frequent, extensive use that distracts users from daily life should not be considered "addiction" unless it also leads to functional impairment and psychological distress [@billieuxDysfunctionalUseMobile2015; @kardefelt-wintherHowCanWe2017; @tranModelingEngagementDisengagementCycle2019].
-Moreover, references to 'addiction' among ordinary users often reflect a loose use of the term common in media and society [@tikuWIREDGuideInternet2018; @lanetteSmartphoneAddictionNarrative2018; @orbenTeensScreensWellBeing2019].
-In the user review data, we did observe a number of mentions of 'addiction' (n = 52, 3%), but overall the data seemed to reflect a situation in which most users' challenges with focusing and avoiding wasted time reflected more mundane struggles and annoyances than what would deserve the clinical term 'addiction'.
+Our thematic analysis yielded less specific information about the 'just right' level of support for the other three types of design patterns.
+Overall, however, our data and the dual systems model suggest that **goal-advancement** patterns should find the 'just right' frequency with which to bring goals into conscious System 2 awareness: too infrequent, and goals are held in mind too briefly for users to complete their tasks; too frequent and users get annoyed.
+**Reward/punishment** should provide a 'just right' incentive that raises the expected value of target use over that of alternative options: too little, and users will still succumb to unwanted habits; too much, and the incentives can undermine users' sense of self-efficacy when they fail.
+**Self-tracking** should present usage information in a form that facilitates goal-setting and activates related behaviour.
+Too vague a connection to goals, and information merely raises awareness without prompting behaviour; too specific, and users might feel judged by the tool and rebel against it.
 
-A range of alternatives to addiction measures for assessing users' experience with DSCTs have been explored, including measurements of self-reported stress (e.g., the Perceived Stress scale, \cite{collinsBarriersEngagementPersonal2014}, from \cite{Cohen1983}), workload (e.g., NASA-TLX, \cite{kimLocknTypeLockoutTask2019}, from \cite{Hart1988}), and absorption in a task (e.g., Cognitive Absorption scale, \cite{markEffectsIndividualDifferences2018} from \cite{agarwalTimeFliesWhen2000}).
-We suggest, however, that the most directly relevant literature to draw on is existing psychological work on self-regulation and self-control.
-Here, a substantial body of research has investigated how individual differences and environmental influences affect the ability to regulate behaviour towards desired goals in the face of conflicting impulses and distractions.
-HCI research on DSCTs can be viewed as a sub-field in this respect, which investigates self-regulation in relation to digital devices such as smartphones and laptops, and how to support it by design.
+### Combining design patterns to provide the Goldilocks level of support {#combining-patterns}
 
-Thus, early research on 'internet addiction', and later 'smartphone addiction', did derive great value from adapting behavioural addiction research to investigate severe cases of self-regulation breakdown among a minority of users [@chakrabortyInternetAddictionConsensus2010].
-Future HCI research on design remedies for more mundane self-regulation challenges experienced by the great majority of users could benefit from a similar effort to adapt common measurement scales from basic psychological research on self-regulation to the context of digital device use.
-As an example, exploratory work by Mark et al. \cite{markEffectsIndividualDifferences2018} found that blocking tools preferentially benefits individuals low in trait self-control.
-The measure of self-control used was sub-items from the Cognitive Absorption scale, a scale developed for measuring a 'state of deep involvement with software'.
-Follow-up confirmatory work could adapt e.g., the widely used UPPS Impulsivity Behaviour scale [@Whiteside2001] to investigate the robustness of Mark et al.'s finding and how it relates to well-established personality factors underpinning self-regulation ability.
+In this section, we discuss an interesting theme in our data, namely how combining design patterns can provide a Goldilocks level of support.
+This was often suggested in the reviews, where self-tracking in particular tended to be described as insufficient support in itself to change behaviour (see also [@collinsBarriersEngagementPersonal2014]).
+Indeed, DSCTs that combined two or more design patterns tended to receive higher ratings compared to those that focused on only one (see also recent psychological research on self-control strategies in day-to-day life, showing that using multiple strategies is especially effective [@milyavskayaSelfControlDaily2021]).
 
-Similarly, the Brief Self-Control Scale [@tangneyHighSelfcontrolPredicts2004] has been widely used by psychologists as a uni-dimensional measure of ability to self-regulate behaviour in the face of undesired impulses, and found to predict positive life outcomes in various domains \cite{Ferrari2009}.
-This scale has also been used as a measure assessing change in self-control ability over time \cite{Duckworth2010}, and might fruitfully be adapted as an effectiveness metric for interventions in DSCTs.
+In the dual systems lens, a DSCT's interventions need to in aggregate provide sufficient influence on the cognitive system to tip people's digital device use down a path that is aligned with their longer-term goals.
+This may be achieved via a single design pattern, or by combining multiple patterns that intervene on different psychological mechanisms.
+The DSCTs and user reviews provide plenty of examples: Browser extensions can place to-do lists on new tabs to remind the user of their tasks, and also provide pomodoro-style blocking of distracting websites.
+The immensely popular app *Forest* [@seekrtechForestStayFocused2018] combines a countdown-timer --- which reminds people that they are currently wanting to not use their phone --- with an animated growing tree that dies if they fail to adhere to the non-use goal.
+Reviewers suggested that distracting recommender feeds could be replaced with content they would like their better selves to spend more time engaging with.
 
-Adapting such measures to the domain of digital device use could provide metrics that more directly measure impact on the ability to self-regulate behaviour in the face of digital distractions than the alternatives commonly used by HCI researchers.
-Therefore, they might have a closer construct fit with what users commonly want DSCTs to achieve, as evidenced by the user review data and existing research, and also more readily allow for knowledge transfer with fundamental psychology research relevant to DSCTs.
+Each design pattern's 'Goldilocks' level of support should interact with the other patterns included in a tool: we suggest that by combining patterns, each pattern may need to provide only a lower level support to be effective, which can make the tool less likely to be experienced as too coercive.
+To illustrate, consider a woman who finds it challenging to do what she intends on Facebook (e.g., post an update in the local Taekwondo group, and message a family member) without being side-tracked by irrelevant information on the site.
+A DSCT might support her via a pop-up that every few minutes reminds her of her intention (goal-advancement).
+She experiments with the frequency of reminders until she finds the 'Goldilocks level': a balance point where her conscious intentions are re-activated sufficiently often for her to stay on task amidst distractions that trigger irrelevant System 1 habits and/or crowds out her intentions from System 2 awareness.
+Imagine the DSCT now also hides her Facebook newsfeed until she actively chooses to see it (block/removal).
+The level of potentially distracting on-site information --- and the related triggering of irrelevant behaviours --- is now lower.
+Therefore, her Goldilocks frequency for the reminders is also likely to change: she is now able to stay on task (and prefers) less frequent reminders.
 
-## Using psychological theory to explore interactions between design patterns and framings of 'distraction'
+We do note that the higher ratings we observed for DSCTs that combine multiple design patterns have alternative explanations than such tools being more likely to provide a Goldilocks level of support: For example, DSCTs that combine multiple design patterns may simply have received more developer attention, and therefore be better designed on average.
+We encourage future research to explicitly investigate how different design patterns interact in DSCTs, for example via studies with factorial designs similar to those called for in digital behaviour change research [@klasnjaUsableEvidenceOptimizing2017].
 
-Existing work has sometimes implied a single spectrum of 'friction' on which the intervention intensity of a given DSCT may be increased or decreased to find an optimal level \cite{kimGoalKeeperExploringInteraction2019, kimLocknTypeLockoutTask2019, kovacsConservationProcrastinationProductivity2019}.
-However, our observations that reviewers often wanted different types of design patterns to be combined, and that a poor fit between the user's and tool's framing of 'good' and 'bad' device use might make people rebel against a tool, suggest that designers need to consider how several psychological dimensions interact.
+## Designing to reduce psychological reactance
 
-One simple framework for this purpose is the dual-systems model presented by Lyngs et al.
-[@lyngsSelfControlCyberspaceApplying2019], which summarises main mechanisms of self-regulation as viewed by current cognitive psychology.
-According to this model, the effects of DSCTs on self-regulation can be considered in terms of three factors: (i) automatic habits and reflexes (System 1 control), (ii) conscious goals and self-monitoring (System 2 control), and (iii) the expected value of control (mediator of System 2 control).
+Another important challenge arising from the user reviews is to design tools that help people navigate negative emotional reactions that DSCTs can cause in the moment of enforcement.
+The data point to two factors in this respect:
 
-At any given point, these factors activate potential behaviours and thoughts, and the action that receives the highest activation controls our behaviour.
-For example, if you take out your smartphone to check Instagram, it can happen via automatic System 1 control in response to a boring situation, because you have a habit of doing so in the presence of certain cues.
-It might also happen via conscious System 2 control if you remember that you need to message a friend, and so this conscious goal triggers behaviour [@lyngsSelfControlCyberspaceApplying2019].
+### Precisely targeting interventions to users' notion of distracting use {#targeting-interventions}
 
-Self-control challenges occur when actions triggered by System 1 and System 2 are in conflict.
-For example, to meet a goal of not checking Instagram at family dinners, you might need to suppress an impulse to check.
-If you fail, it can be because you had forgotten this goal, or because the 'expected value of control', which mediates the strength of System 2 control, was too low: The more value a conscious goal has (whether because its rewards are greater or more immediate, or because you are more confident in your ability to achieve it), the more powerfully it triggers action.
-In other words, you might fail to suppress your impulse to stay off your phone, because your valuation of staying off was not high enough to compete with the urge to check.
+Future work should focus on how DSCTs' interventions can be precisely tailored to aspects of users' digital environments that cause self-regulation difficulties.
+The more DSCTs' interventions are off the mark in this respect, the more they may cause reactance.
 
-We find this simple partitioning of the main psychological dimensions of self-regulation helpful for considering the purpose and design space of DSCTs.
-Thus, from this perspective, the purpose of DSCTs is to bias the internal competition between potential actions such that the ones that win out are in line with the user's enduringly valued conscious goals [@duckworthStitchTimeStrategic2016].
-DSCTs can achieve this via interventions that target a limited number of factors --- blocking tools can help prevent unwanted System 1 responses from being triggered or acted on, goal advancement and self-tracking tools can enable System 2 control by activating certain goals in working memory, and reward/punishment tools can make System 2 control more likely to succeed by increasing the expected value of control [@lyngsSelfControlCyberspaceApplying2019].
+Most research on DSCTs has focused on innovating 'external mechanisms' that help people manage or monitor their device use, i.e., design patterns like lockout timers or productivity dashboards that apply similarly across many different apps [@lukoffHowDesignYouTube2021].
+However, this approach is often too crude, as evidenced by user reviews expressing a desire for control over, for example, what specific website features are blocked by DSCTs, or what apps count toward 'screen time'.
+Indeed, numerous Chrome Web Store tools included in the present paper serve to redesign only specific design features experienced as problematic (e.g.
+*Inbox When Ready for Gmail* hides the inbox on gmail.com until a button is pressed; *Click to Remove Element* lets users remove website elements by clicking on them).
 
-We can use this model to consider users' suggestions of combining distinct types of design patterns: it is possible that by combining design patterns which target distinct psychological mechanisms, each can be implemented at a more gentle level of intensity while still in aggregate provide enough input to the internal competition to bring about an intended behaviour change.
-By contrast, tools that focus on a single design pattern targeting a single psychological mechanism may need to ramp up its intensity more highly before sufficient impetus for behaviour change is provided --- and in so doing be at greater risk of being perceived as overly annoying.
-For example, \emph{Forest} combines growing of virtual trees with a countdown timer, which influences both the expected reward for control and our sensitivity to delays [@lyngsSelfControlCyberspaceApplying2019].
-Based on this tool's popularity, with over 10 million users on Android alone, this approach seems highly effective for many.
-An alternative version which simply grew virtual trees without a countdown timer might need to correspondingly increase the magnitude of reward to remain useful, but in so doing be more likely to be perceived by users as too forceful or manipulative [@Lee2011Mining].
-[^4]
+In research so far, less effort has been invested in developing and evaluating such tools that empower users to change 'internal mechanisms' within apps and services.
+However, recent studies of the effects or directly modifying design elements of Facebook or the mobile Twitter client (e.g., by hiding specific UI elements) provides strong support for the potential of this approach [@zhangMonitoringScreenTime2022; @lyngsJustWantHack2020].
+Similarly, new research on the *GreaseVision* platform aims to provide researchers and end users with accessible tools for modifying UI elements in mobile apps that they find distracting or harmful [@dattaGreaseVisionRewritingRules2022; @kollnigGreaseDroid2021; @greaseTerminator].
+This holds great promise for future research on how DSCTs can support self-regulation in a precisely targeted fashion to minimise negative emotional reactions to their interventions.
 
-[^4]: We do note that the higher ratings given to tools combining two or more distinct design patterns has an alternative explanation: tools that combine design patterns may simply have received more developer attention, and so higher ratings could simply be due to such tools being better designed.
+### Developing interventions that support internalisation of motivation for DSCT use
 
-Similarly, we may reach to the psychological literature for design guidance related to the observation that alignment between the user and tool in the framing of distraction is important for the user in managing potential negative emotions.
-In Lyngs et al.'s dual-systems model, this might be considered a mediator of the expected value of control, but the model is too high level to provide useful advice in this respect.
-Instead, we propose that future research draw on Self-Determination Theory (SDT), which has been frequently used and well-validated in HCI research but rarely applied in research on DSCTs \cite{ballouSelfDeterminationTheoryInHCI2022, ryan2017}.
-SDT research on the factors that determine whether people identify and integrate with a source of extrinsic motivation, or whether they experience it as a less authentic external influence, could prove highly useful for understanding how to frame DSCTs' interventions so that they are more likely to be accepted by the user as a desirable extension of their personal will.
+Another research and design opportunity concerns the ways in which interventions are framed at the moment of enforcement.
+Moving beyond the dual-systems model, self-determination theory (SDT), an evidence-based theory of human motivation, offers a promising explanation for why DSCTs can trigger frustration and even hostility ([@ryanSelfdeterminationTheoryBasic2017]; recall that reviewers stated that they simultaneously 'loved and hated' their distraction blockers --- even though DSCTs were valuable, they were also 'pissed off' by them).
+According to SDT, people have a basic need for autonomy.
+When a DSCT applies enforcement mechanisms that hold users accountable to their goals, it can be perceived by users as a form of external control that threatens autonomy [@brehm2013Reactance].
+This can cause 'reactance', an unpleasant motivational arousal in the user, which may result in behavioural backlash where the users tries to circumvent or intentionally contradict the tool [@gavanReactance2004].
 
-We encourage future research to explicitly investigate interactions between interventions types and framings of 'distraction' using these psychological theories as guide.
+However, DSCTs' interventions need not be perceived in this way.
+In SDT, 'extrinsic motivation' refers to activity done to obtain some outcome other than the enjoyment of that activity itself.
+Usage of DSCTs is typically extrinsically motivated, because people use them to obtain some change in their use of digital devices.
+Extrinsic motivation can be experienced as more or less autonomous (see Figure \@ref(fig:sdt-figure)): at one extreme, use of a DSCTs can be experienced as fully controlled by external rewards or punishments (e.g., "I have to use the DSCT"), and therefore be likely to lead to reactance.
+At the other, their use might be experienced as fully autonomous and directed by the user's personal goals (e.g.
+"I use the DSCT because it helps me achieve my goals and values"), which should cause less reactance.
+
+(ref:sdt-figure-caption) Self-determination theory distinguishes between different types of motivation, including extrinsic (doing an activity to obtain some separable outcome) and intrinsic motivation (doing an activity because it is inherently interesting or enjoyable).
+DSCT use tends to be extrinsically motivated, because it serves the separable outcome of changing one's patterns of device use.
+DSCTs' interventions can be experienced as externally controlled (“I have to use the DSCT”) and trigger reactance.
+Future research should explore how DSCTs might be designed to consistently be experienced as supporting the user's autonomy (“I use the DSCT because it helps me achieve my goals and values”).
+Illustration adapted from [@lukoffDesigningSupportAutonomy2022], who adapted it from  [@peters2018DesigningForMotivation] and [@ryan2000self].
+
+```{r sdt-figure, out.width='100%', fig.cap="(ref:sdt-figure-caption)"}
+knitr::include_graphics("figures/SDT_reactance.png")
+```
+
+DSCTs that provide distraction blocking currently tend to simply let users set target apps or websites and then provide generic "You've reached your limit" lock screens at the time of enforcement.
+Whereas some of the DSCTs we analysed provided creative variations (e.g., Go Fucking Work shouts obscenities at its users on its lock screens, and has a dedicated niche following; RescueTime's lock screen shows an animated dog with a wagging tail sending a shameful look to the user), there is much room for researchers and designers to explore ways to create interventions that support the experience of DSCT use as being directed by the users's personal goals and values.
+For example, future research could explore how to support autonomy at the time when the user delegates their goals to the DSCT.
+Here, DSCTs might borrow from motivational interviewing techniques to elicit and affirm the reasons behind the user's goals [@Miller2009-ir].
+Future research should also explore how to support autonomy at the time when the DSCT enforces the user's goal.
+Here, DSCTs might, e.g., remind the user of the 'why' behind their use of the tool, using their own words (see [@xuTypeOutLeveragingJustinTime2022] for a recent example), or increase the salience of rewards they value by displaying pictures of loved ones they could spend time with after finishing their tasks [@lukoffDesigningSupportAutonomy2022].
 
@@
 
-Looking towards the future, we note that the underlying technical infrastructures and platforms create very different conditions for the emergence of ecosystems of DSCT in the wild.
-The largest degree of flexibility exists on the web, where browser extensions make it possible for developers --- and tech-savvy end-users --- to easily customise the look and functionality of any website via injected stylesheets and scripts.
-Ecosystems of browser extensions today exist for all major browsers, which all share the cross-browser \href{https://developer.mozilla.org/en-US/docs/Mozilla/Add-ons/WebExtensions/API}{Web Extensions API}.
+The experience of reviewers depended heavily on the underlying technical infrastructures, which create different conditions for the emergence of ecosystems of DSCT in the wild.
+The largest degree of flexibility exists on the web, where browser extensions make it possible for developers --- and tech-savvy end-users --- to customise the look and functionality of any website via injected stylesheets and scripts.
+Ecosystems of browser extensions today exist for all major browsers, which all share the cross-browser [Web Extensions API](https://developer.mozilla.org/en-US/docs/Mozilla/Add-ons/WebExtensions/API).
 
-The situation is very different if a user wishes to experiment with DSCTs on a smartphone or tablet.
-Unless users jailbreak their devices, no options exist for users to customise the look and function of mobile apps beyond the options that app developers chooses to provide.
-Moreover, the options available for self-tracking or blocking tools are dramatically different between Android and iOS: Because of Google's less restrictive approach to developer permissions, a diverse ecosystem exists on Android for apps that provide bespoke usage visualisation or customised blocking schedules, in addition to what Google provides themselves in their 'Digital Wellbeing' app.
-On iOS, by contrast, Apple's Screen Time tool was, at the time data for the present paper was scraped, the only option with which users could block functionality and see their usage of different apps[^5].
-As we observed in the user reviews, workarounds by iOS developers, such as asking users to routinely take screenshots of the battery page on their phone settings, seemed to be too effort-demanding on the part of users to be workable.
+The situation is very different if a user wishes to experiment with DSCTs on their smartphone or tablet.
+Unless users jailbreak their devices, no options exist for users to customise the look and function of mobile apps beyond the options that app developers choose to provide.
+Moreover, the options available for self-tracking or blocking tools differ dramatically between Android and iOS: because of Google's less restrictive approach to developer permissions, a diverse ecosystem exists on Android for apps that provide bespoke usage visualisation or customised app blocking schedules, in addition to what Google provides in their 'Digital Wellbeing' app.
+On iOS, by contrast, Apple's Screen Time tool was, at the time data for the present paper was scraped, the only option with which users could properly block apps or see their usage[^5].
+As mentioned in the results section, DSCTs on iOS sometimes resorted to creative workarounds such as 'blocking' apps by asking users to set a new, random password for their social media accounts and then log out, after which they would only see the password again on a set date, or tracking usage by getting users to take screenshots of the battery page in the smartphone settings.
+Unsurprisingly, such workarounds tended to be seemed by users as too effort-demanding to be workable.
 
-[^5]: A Screen Time API, introduced in the summer of 2021 after long campaigning by developers [@digitalwellnesswarriorsAppleLetDevelopers2018], now allows creation of blocking tools for parental control purposes, but does not enable apps for tracking and visualising screen time.
+We do not take a strong stance on whether Apple or Google are 'right' in terms of the level of permissions they make available to developers, as greater operating system access also entails greater security risks.
+However, we do note that such risks are already present in browser extensions, which both Apple and Google, alongside other companies, already allow in their web browsers.
+In relation to the web, such extensions already give users the power to locally block access to specific websites, or provide detailed modifications to their appearance and functionality.
 
-We make no judgment of whether Apple or Google are 'right' in terms of the level of permissions to make available to developers, as greater operating system access also entails greater security risks.
-We do note, however, that such risks are already present in browser extensions, where both Apple and Google, and other companies, already allow extensions that in many respects have greater powers than Android apps.
-Here we simply wish to highlight the distinct roads ahead, in relation to providing the interventions that users seek to stay in control:
-
+Here, we mainly wish to highlight how Google and Apple's approaches create distinct roads ahead in relation to providing the support that users seek to stay in control.
 Google's less restrictive approach has meant that a diverse range of potent digital self-control tools are available on Android.
-Therefore, Google may face less pressure to 'get it right' with their own Digital Wellbeing tools, because less common user needs can be addressed by tools within an ecosystem of DSCTs on their platform.
+Therefore, Google may face less pressure to 'get it right' with their own Digital Wellbeing tools, because less common user needs can be addressed by third-party tools within an ecosystem of DSCTs on their platform.
 Apple's approach, however, implies that their own Screen Time app needs to accommodate a very wide range of user needs.
 This will be challenging, as there can be a trade-off between customisability and effort of use, where providing sufficiently fine-grained controls to account for all users' needs can make it too complicated to navigate.
+Researcher-led initiatives such as *GreaseVision* (see section [4.2.1](#targeting-interventions){reference-type="ref" reference="targeting-interventions"}), might point the way towards a more inclusive future where end-users have the power to modify mobile user interfaces to suit their individual needs [@dattaGreaseVisionRewritingRules2022].
 
@@
 First, the data used in the present paper was scraped in March 2019.
-Hence, the findings presented may deviate from what would by learned by re-scraping the app and browser extension stores in 2022.
-However, all of our data and materials are openly available, and so our process should be readily repeatable.
-As such, we believe the present paper provides a useful snapshot of DSCTs across three online stores, which can serve as point of comparison for future research tracking how the landscape of these tools have developed.
+Hence, the findings presented may deviate from what would be learned by re-scraping the app and browser extension stores in 2022.
+However, all of our data and materials are openly available, so our process should be readily repeatable.
+As such, we believe the present paper provides a useful snapshot of DSCTs across three online stores, which can serve as point of comparison for future research tracking how the landscape of these tools develop.
 
@@
 That is, for a user to be sufficiently motivated to bother reviewing a tool in the first place, their level of excitement about a tool might exceed the excitement of the average user [@panichellaHowCanImprove2015].
-Neverthelesss, the fact that our findings from reflexive thematic analysis resonated well with findings from studies using other methods such as surveys or interviews [@koNUGUGroupbasedIntervention2015; @tranModelingEngagementDisengagementCycle2019] suggests that user reviews of DSCTs can provide valid and useful insights.
+Neverthelesss, the fact that our findings from reflexive thematic analysis resonate well with findings from studies using other methods such as surveys or interviews [@koNUGUGroupbasedIntervention2015; @tranModelingEngagementDisengagementCycle2019] suggests that user reviews of DSCTs can provide valid and useful insights.
 
@@
 for researchers specifically investigating design parameters relevant for lockout mechanisms [@kimGoalKeeperExploringInteraction2019].
-We encourage future research on DSCTs to make their data sets and analysis procedures open for others to re-use, to encourage synergies between the efforts of different research groups in the field.
+We encourage future research on DSCTs to make their data sets and analysis procedures open for others to re-use, to enable synergies between the efforts of different research groups in the field.
 
@@
 In this paper, we investigated what user reviews reveal about digital self-control tools' contexts of use and design challenges, and how design features relate to ratings and installation numbers, using data from 334 tools on the Google Play, Chrome Web, and Apple App stores.
-Compared to existing research, our work followed a more transparent procedure for sampling user reviews for thematic analysis from the full data set, with reproducible, programmatic sampling and open data and materials.
+Compared to existing research, our work followed a more transparent procedure for sampling user reviews for thematic analysis from the full data set, with reproducible, programmatic sampling, and open data and materials.
 
-As most daily tasks come to be mediated by digital devices, it is important that users are given the tools they need to effectively manage time and attention in their digital environments.
-As such, discovering effective interventions for support self-regulated digital device use is a practical and urgent research challenge.
+As most daily tasks come to be mediated by digital devices, it is important that users have the tools they need to effectively manage time and attention in their digital environments.
+As such, discovering effective interventions to support self-regulated digital device use is a practical and urgent research challenge.
 In this respect, we encourage future work to make further use of the landscape of digital self-control tools on app and browser extension stores as a resource for investigating effective design solutions.
@@
 We thank Claudine Tinsman for proof-reading a late version of the manuscript.
+
+[^1]: We use the term 'design pattern' to refer to specific functions intended to support people in self-regulating digital device use (e.g., hiding newsfeeds, providing goal reminders).
+We use the term 'digital self-control tool' (DSCT) to refer to apps and browser extensions that implement a given design pattern (e.g., a browser extension for hiding Facebook's newsfeed).
+We occasionally use the term 'intervention' to refer to the implementation of a design pattern, in the context of how it changes a user's usual digital environment, or when it is being evaluated in a research study.
+
+[^2]: After the fairy tale 'Goldilocks and the Three Bears', where a young girl tastes three different bowls of porridge and prefers the one that is neither too hot nor too cold, but 'just right'.
+This concept of the 'just right amount' is often referred to as the 'Goldilocks principle'.
+As an example, a widely cited paper on the relationship between screen time and mental well-being found that a balance point in between "too much" and "too little" tech use seemed optimal, which the authors referred to as the 'digital Goldilocks hypothesis' [@Przybylski2017]
+
+[^3]: See Lyngs et al. [@lyngsSelfControlCyberspaceApplying2019] and their supplementary materials on [osf.io/zyj4h/](https://osf.io/zyj4h/) for further details.
+
+[^5]: A Screen Time API, introduced in the summer of '21 after long campaigning by developers [@digitalwellnesswarriorsAppleLetDevelopers2018], now allows creation of blocking tools for parental control purposes, but does not enable apps for tracking and visualising screen time.

### 2022-06-10 paper.Rmd
add missing nums

@@
 
-`r nrow(got_reviews_from)` of these tools had received user reviews (`r num_tools_w_reviews_by_store %>% filter(store == "chrome") %>% pull(n)` from Chrome Web, `r num_tools_w_reviews_by_store %>% filter(store == "google_play") %>% pull(n)` from Google Play, and `r num_tools_w_reviews_by_store %>% filter(store == "apple") %>% pull(n)` from the Apple App store).
-From these, we collected a total of `r format(nrow(reviews_w_main_cat_data), big.mark = ",")` reviews (`r num_reviews_by_store %>% filter(store == "chrome") %>% pull(n) %>% format(big.mark = ",")` from Chrome Web, `r num_reviews_by_store %>% filter(store == "google_play") %>% pull(n) %>% format(big.mark = ",")` from Google Play, and `r num_reviews_by_store %>% filter(store == "apple") %>% pull(n) %>% format(big.mark = ",")` from the Apple App store).
+`r nrow(got_reviews_from)` of these tools had received user reviews (`r num_tools_w_reviews_by_store %>% filter(store == "chrome") %>% pull(n)` from Chrome Web, `r num_tools_w_reviews_by_store %>% filter(store == "play") %>% pull(n)` from Google Play, and `r num_tools_w_reviews_by_store %>% filter(store == "apple") %>% pull(n)` from the Apple App store).
+From these, we collected a total of `r format(nrow(reviews_w_main_cat_data), big.mark = ",")` reviews (`r num_reviews_by_store %>% filter(store == "chrome") %>% pull(n) %>% format(big.mark = ",")` from Chrome Web, `r num_reviews_by_store %>% filter(store == "play") %>% pull(n) %>% format(big.mark = ",")` from Google Play, and `r num_reviews_by_store %>% filter(store == "apple") %>% pull(n) %>% format(big.mark = ",")` from the Apple App store).
 
@@
 For Google Play tools, the median 'minimum number of installs' was `r format(play_user_stats$median_min_installs, big.mark = ",")` (min = `r format(play_user_stats$min_min_installs, big.mark = ",")`, max = `r format(play_user_stats$max_min_installs, big.mark = ",")`, IQR = `r format(play_user_stats$q25_min_installs, big.mark = ",")` to `r format(play_user_stats$q75_min_installs, big.mark = ",")`).
-For Chrome Web tools, the median number of users was `r format(chrome_user_stats$media_installs, big.mark=",")` (min = `r format(chrome_user_stats$min_installs, big.mark=",")`, max = `r format(chrome_user_stats$max_installs, big.mark=",")`, IQR = `r format(chrome_user_stats$q25_installs, big.mark=",")` to `r format(chrome_user_stats$q75_installs, big.mark=",")`).
+For Chrome Web tools, the median number of users was `r format(chrome_user_stats$median_installs, big.mark=",")` (min = `r format(chrome_user_stats$min_installs, big.mark=",")`, max = `r format(chrome_user_stats$max_installs, big.mark=",")`, IQR = `r format(chrome_user_stats$q25_installs, big.mark=",")` to `r format(round(chrome_user_stats$q75_installs), big.mark=",")`).
 The Apple App store did not display information about number of users (the median number of ratings for DSCTs from the Apple App store was `r format(num_ratings_apple$median, big.mark=",")`; max = `r format(num_ratings_apple$max, big.mark=",")`, IQR = `r format(num_ratings_apple$q25, big.mark=",")` to `r format(num_ratings_apple$q75, big.mark=",")`).
-The median average rating across all tools was `r format(ratings_overall$median_rating, big.mark=",")` (min = `r format(ratings_overall$min_rating, big.mark=",")`, max = `r format(ratings_overall$max_rating, big.mark=",")`, IQR = `r format(ratings_overall$q25_rating, big.mark=",")` to `r format(ratings_overall$q75_rating, big.mark=",")`; ratings given on a scale from 1 to 5 stars on all stores).
+The median average rating across all tools was `r format(ratings_overall$median_rating, big.mark=",")` (min = `r format(ratings_overall$min_rating, big.mark=",")`, max = `r format(ratings_overall$max_rating, big.mark=",")`, IQR = `r format(round(ratings_overall$q25_rating), nsmall=1)` to `r format(round(ratings_overall$q75_rating), nsmall=1)`; ratings given on a scale from 1 to 5 stars on all stores).
 

### 2022-06-10 figures/figures.Rmd
correct figure plot code

@@
 library(cowplot)
-library(ggtextures)
+library(ggtextures) #remotes::install_github("clauswilke/ggtextures") 
 ```
@@
 ```{r}
-### try and draw the apple one again with color
+# set up the mapping between categories and colors
 tools_w_categories_and_image <- meta_data %>% 
@@
 # Plot of tools by user numbers
-generate apple plot
+## generate apple plot
 ```{r}
@@
 
-save_plot(here::here("figures", "apple_ratings_and_features.png"), test_apple,
+save_plot(here::here("figures", "apple_ratings_and_features.png"), apple_combined,
           # each individual subplot should have an aspect ratio of 1.3
@@
 
-generate google play plot
+## generate google play plot
 
@@
 
-(play_combined <- plot_grid(play_install_bar, play_feature_plot, rel_widths = c(4,1), labels = c("Google Play", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1))
+play_combined <- plot_grid(play_install_bar, play_feature_plot, rel_widths = c(4,1), labels = c("Google Play", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1)
 
@@
 
-generate chrome web plot
+## generate chrome web store plot
 
@@
 
-put them all together
+## Combine them
 
@@
   mutate(categories_alphabetically = factor(categories_alphabetically, levels = rev(ave_ratings_levels), ordered = TRUE)) %>%
+  rename(Store = store) %>% 
   ggplot(aes(x = categories_alphabetically, y = average_rating)) +
     geom_boxplot(outlier.shape = NA) +
-    geom_jitter(width = 0.3, height = 0.1, size = 1.1, alpha = 0.7) +
+    geom_jitter(width = 0.34, height = 0.06, size = 1.1, alpha = 0.8, aes(fill = Store, shape = Store)) +
     theme_minimal() +
@@
     coord_flip() +
-    theme(axis.text.y=element_blank()) +
-    theme(plot.margin = margin(1.5, 0, 0, 0, "cm")) +
-    scale_y_continuous(limits = c(1,5))
+    theme(axis.text.y=element_blank(), plot.margin = margin(1.5, 0, 0, 0, "cm"), legend.position="left", legend.box.margin=margin(0,-0.5,0,0, "cm")) +
+    scale_y_continuous(limits = c(1,5)) +
+    scale_fill_manual(values = c("darkgrey", "#f0027f", "white")) +
+    scale_shape_manual(values = c(22, 21, 25))
 
@@
 ggsave(here::here("figures", "ratings_by_type_annotated.pdf"), height = 4.3, width = 7)
+
 ```
